### 2012-01-23 Paper/draft1_MCMCsmoother.tex
First commit: MCMC smoother, new state smoother and papers drafts.

\documentclass[journal]{IEEEtran}

%% Packages
\usepackage{cite}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{{../pdf/}{../jpeg/}}
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  \usepackage[dvips]{graphicx}
  \graphicspath{{../eps/}}
  \DeclareGraphicsExtensions{.eps}
\fi
\usepackage[cmex10]{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{stfloats}
\usepackage{url}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Faster particle smoothing using Markov chain Monte Carlo sampling}

\author{Pete~Bunch,~\IEEEmembership{}
        Simon~Godsill,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
\thanks{P. Bunch and S. Godsill are with the Department
of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}% <-this % stops a space
\thanks{Manuscript received January 01, 1901; revised January 02, 1901.}}

% The paper headers
\markboth{IEEE Transaction in Signal Processing,~Vol.~1, No.~1, January~1901}%
{Shell \MakeLowercase{\textit{et al.}}: Particle smoothing using Markov chain Monte Carlo sampling for backwards simulation}

% make the title area
\maketitle

\begin{abstract}
The abstract goes here.
\end{abstract}

\begin{IEEEkeywords}

\end{IEEEkeywords}

\section{Introduction}

\IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quanitity from incomplete or inaccurate observations. This is commonly achieved through the use of probabilistic models for the evolution of the latent state and the measurement process.

\begin{IEEEeqnarray}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \label{eq:state_proc}\\
y_{k} & \sim & p(y_{k}|x_{k})   \label{eq:observ_proc}
\end{IEEEeqnarray}

Here, the random variable $x_k$ denotes the value of the latent state at the $k^{th}$ instant, and $y_k$ the value of the observation. A set of random variables will be written as $x_{1:k} = \{x_1, x_2, \dots, x_k \}$. The state evolution is assumed to be Markovian, i.e. $x_k$ depends only on the previous value, $x_{k-1}$.

Commonly of interest are the problems of filtering and smoothing. Filtering is the inference of $p(x_k|y_{1:k})$, the distribution of the current state given all previous observations. Smoothing consists of two related tasks, the inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps), the joint distribution of the entire state sequence given all the observations, and that of $p(x_{k}|y_{1:K})$, the marginal distribution of each state given all the observations. In this paper, we are concerned with the estimation of the joint smoothing distribution.

Several problems are of interest; first, the inference of the \emph{filtering} distributions, $p(x_k|y_{1:k})$, an estimate of the current state given all previous observations; second, that of the \emph{smoothing} distribution, $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps), an estimate of the entire state sequence given all the observations; and third, that of the \emph{marginal smoothing} distributions, $p(x_{k}|y_{1:K})$ for each $k$. The estimation of these marginal smoothing distributions is not considered in this paper.

In the case where the state and observation processes of equations~\ref{eq:state_proc}~and~\ref{eq:observ_proc} are linear and Gaussian, the filtering problem may be solved analytically using the Kalman filter \cite{Kalman1960}. For nonlinear or non-Gaussian models, tractable closed-form solutions are rare. In such cases, numerical approximations are often employed, including the particle filter algorithm, first introduced by \cite{Gordon1993}. (See \cite{Cappe2007,Doucet2009} for a thorough introduction.)

In a particle filter, the filtering distribution is approximated by a set of weighted samples (or ``particles'') drawn from that distribution.

\begin{IEEEeqnarray}{rCl}
\hat{p}(x_{k}|y_{1:k}) & = & \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)
\end{IEEEeqnarray}

Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$. $\hat{p}$ means an approximation to $p$.

A similar story applies to the problem of smoothing. In the linear Gaussian case, the entire state sequence, $x_{1:K}$, may be estimated in closed-form using the a Rauch-Tung-Striebel (RTS) smoother \cite{Rauch1965}. This works by modifying the ordinary Kalman filter results in a backward processing pass through the observations. For nonlinear or non-Gaussian models, it is possible again to resort to numerical methods. Now, the particles of the approximation are drawn from the smoothing distribution.

\begin{IEEEeqnarray}{rCl}
\hat{p}(x_{1:K}|y_{1:K}) & = & \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})
\end{IEEEeqnarray}

Each particle is an entire realisation of the state process, with $K$ values corresponding to the state at each time step. The conventional method for generating these samples was described in \cite{Godsill2004}. In this paper a new method is presented for drawing samples from the smoothing distribution, using Markov chain Monte Carlo (MCMC). This novel procedure has computational advantages and greater flexibility over the previous methods.

In section~\ref{}, we review briefly review the basic particle filter and smoother algorithms. The new MCMC-based smoothing method is presented in section~\ref{}, with supporting simulations in section~\ref{}.

\section{Particle Filtering and Smoothing}

\subsection{The Particle Filter}

The particle filter is a recursive numerical algorithm for estimation of the filtering distribution. At the $k^{th}$ processing step, a particle estimate of the joint distribution over $x_{k-1}$ and $x_k$ is made by drawing samples from a proposal distribution.

\begin{IEEEeqnarray}{rCl}
\{ x_{k-1}^{(i)}, x_k^{(i)} \} & \sim & q(x_{x}|x_{k-1}) q(x_{k-1})
\end{IEEEeqnarray}

The $k-1$ state proposal distribution, $q(x_{k-1})$, is constructed from the particles of the filtering distribution from the previous time step. The choice of weights and sampling procedure here determines what from of ``resampling'' is to be used. Here we use arbitrary weights for generality.

\begin{IEEEeqnarray}{rCl}
q(x_{k-1}) & = & \sum_i v_{k-1}^{(i)} \delta_{x_{k-1}^{(i)}}(x_{k-1})
\end{IEEEeqnarray}

Note that the choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999}. For the least computational expense, the proposal distribution with weights $v_{k-1}^{(i)} = 1/N_P$ (where $N_p$ is the number of particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling. (See \cite{Cappe2007,Doucet2009} for further discussion of resampling.)

Next, the particles are assigned an importance weight according to the ratio of the targeted joint distribution and the proposal.

\begin{IEEEeqnarray}{rCl}
w_{k}^{(i)} & =       & \frac{ p(x_{k-1}^{(i)}, x_k^{(i)}|y_{1:k}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) q(x_{k-1}^{(i)}) } \nonumber \\
            & \propto & \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) p(x_{k-1}^{(i)}|y_{1:k-1}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) q(x_{k-1}^{(i)}) } \nonumber \\
            & \approx & \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }
\end{IEEEeqnarray}

Normalisation is enforced by scaling the weights so that they sum to 0. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the values from each particles.

\begin{algorithm}
  \begin{algorithmic}
  	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
  	\FOR{$k = 1 \dots K$}
  		\FOR{$i = 1 \dots N_P$}
  			\STATE Sample $\{ x_{k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
  			\STATE Weight $w_{k}^{(i)} \propto \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }$.
  			\STATE Discard $x_{k-1}^{(i)}$.
  		\ENDFOR
  	  \STATE Scale weights so that $\sum_i w_{k}^{(i)} = 1$.
  	\ENDFOR
  \end{algorithmic}
  \caption{Particle filter algorithm}
  \label{alg:PF}
\end{algorithm}

\subsection{Existing Algorithms for Particle Smoothing }

The particle filter generates a numerical approximation to each filtering distribution, $p(x_k|y_{1:k})$ by simulating a set of samples from each. A similar sampling-based method may be used to estimate the smoothing distribution.

The most basic particle smoother was presented in \cite{Kitigawa1996}, and is a trivial extension of the particle filter. Rather than marginalising the previous states from each particle, these past values are stored as well. Thus, at each step the algorithm generates an approximation to $p(x_{1:k}|y_{1:k})$. The output from the final processing step is an approximation to the desired smoothing distribution.

The weakness of the Kitigawa smoother (KS) is degeneracy. Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, while those with high weights appear multiple times. The result is that many, or even all, particles will have the same values at early time steps. This effect is illustrated in figure~\ref{} in section~\ref{}. If the objective is only to produce a single point estimate then this might be sufficient, but to characterise the complete smoothing distribution the approximation must be rejeuvenated.

\begin{algorithm}
  \begin{algorithmic}
  	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
  	\FOR{$k = 1 \dots K$}
  		\FOR{$i = 1 \dots N_P$}
  			\STATE Sample $\{ x_{1:k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
  			\STATE Weight $w_{k}^{(i)} \propto \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }$.
  		\ENDFOR
  	  \STATE Scale weights so that $\sum_i w_{k}^{(i)} = 1$.
  	\ENDFOR
  \end{algorithmic}
  \caption{Kitigawa smoother algorithm}
  \label{alg:PF}
\end{algorithm}

An improved particle smoothing algorithm was presented in \cite{Godsill2004}.

\newpage

\begin{itemize}
	\item Kitigawa smoother
	\item Doucet re-weighting smoother
	\item Godsill re-sampling smoother
	\item Complexities
	\item Algorithms
\end{itemize}

\section{New MCMC Particle Smoother}

\begin{itemize}
	\item Motivate - high complexity, low speed of other smoothers
	\item MCMC avoids need to calculate all weights at each step
	\item maths
	\item algorithm
	\item complexity
\end{itemize}

\subsection{Performance Control}

\begin{itemize}
	\item The effect of varying M, the number of MH steps
	\item Situations where the new smoother works well/poorly.
	\item Backward sampling weight ESS.
\end{itemize}

\section{Simulations}

%%%% Template bits

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
\bibliography{D:/pb404/Dropbox/PhD/OTbib}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Pete Bunch}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{Simon Godsill}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2012-01-23 Paper/draft2_MCMCsmoother.tex
First commit: MCMC smoother, new state smoother and papers drafts.

--- 2012-01-23 Paper/draft1_MCMCsmoother.tex
@@
 
-\section{Introduction}
+\section{Introduction} \label{sec:intro}
 
@@
 
-Several problems are of interest; first, the inference of the \emph{filtering} distributions, $p(x_k|y_{1:k})$, an estimate of the current state given all previous observations; second, that of the \emph{smoothing} distribution, $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps), an estimate of the entire state sequence given all the observations; and third, that of the \emph{marginal smoothing} distributions, $p(x_{k}|y_{1:K})$ for each $k$.
-The estimation of these marginal smoothing distributions is not considered in this paper.
-
 In the case where the state and observation processes of equations~\ref{eq:state_proc}~and~\ref{eq:observ_proc} are linear and Gaussian, the filtering problem may be solved analytically using the Kalman filter \cite{Kalman1960}.
@@
 For nonlinear or non-Gaussian models, it is possible again to resort to numerical methods.
-Now, the particles of the approximation are drawn from the smoothing distribution.
+Now, the particles of the approximation are drawn from the joint smoothing distribution.
 
@@
 Each particle is an entire realisation of the state process, with $K$ values corresponding to the state at each time step.
-The conventional method for generating these samples was described in \cite{Godsill2004}.
-In this paper a new method is presented for drawing samples from the smoothing distribution, using Markov chain Monte Carlo (MCMC).
-This novel procedure has computational advantages and greater flexibility over the previous methods.
+Methods for generating these samples have been described in \cite{Kitagawa1996,Godsill2004,Briers2010}.
+In this paper new methods are presented for drawing samples from the smoothing distribution, using Markov chain Monte Carlo (MCMC).
+These novel procedures have computational advantages and greater flexibility over previous methods.
 
-In section~\ref{}, we review briefly review the basic particle filter and smoother algorithms.
-The new MCMC-based smoothing method is presented in section~\ref{}, with supporting simulations in section~\ref{}.
+In section~\ref{sec:basics}, we review briefly review the basic particle filter and existing joint smoother algorithms.
+The new MCMC-based smoothing methods are presented in sections~\ref{sec:mcmc_smoother} and~\ref{sec:new_state_smoother} with supporting simulations in section~\ref{sec:simulations}.
 
-\section{Particle Filtering and Smoothing}
+\section{Particle Filtering and Smoothing} \label{sec:basics}
 
@@
 Note that the choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999}.
-For the least computational expense, the proposal distribution with weights $v_{k-1}^{(i)} = 1/N_P$ (where $N_p$ is the number of particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling.
+For the least computational expense, the proposal distribution with weights $v_{k-1}^{(i)} = 1/N_F$ (where $N_F$ is the number of filter particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling.
 (See \cite{Cappe2007,Doucet2009} for further discussion of resampling.)
@@
 
+The particle filter is summarised in algorithm~\ref{alg:PF}.
+
 \begin{algorithm}
@@
   	\FOR{$k = 1 \dots K$}
-  		\FOR{$i = 1 \dots N_P$}
+  		\FOR{$i = 1 \dots N_F$}
   			\STATE Sample $\{ x_{k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
@@
 
-The most basic particle smoother was presented in \cite{Kitigawa1996}, and is a trivial extension of the particle filter.
+The most basic particle smoother was presented in \cite{Kitagawa1996}, and is a trivial extension of the particle filter.
 Rather than marginalising the previous states from each particle, these past values are stored as well.
@@
 
-The weakness of the Kitigawa smoother (KS) is degeneracy.
+The weakness of this simple, filter-smoother (FS) is particle degeneracy.
 Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, while those with high weights appear multiple times.
@@
 
+The filter-smoother is summarised in algorithm~\ref{alg:FS}.
+
 \begin{algorithm}
@@
   	\FOR{$k = 1 \dots K$}
-  		\FOR{$i = 1 \dots N_P$}
+  		\FOR{$i = 1 \dots N_F$}
   			\STATE Sample $\{ x_{1:k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
@@
   \end{algorithmic}
-  \caption{Kitigawa smoother algorithm}
-  \label{alg:PF}
+  \caption{Filter-smoother algorithm}
+  \label{alg:FS}
 \end{algorithm}
@@
 An improved particle smoothing algorithm was presented in \cite{Godsill2004}.
+This works by constructing new particles from the smoothing distribution by sampling from the filtering distributions.
+This has the limitation that the particle support for the smoother is limited to that of the filter.
+The sampling procedure exploits the following factorisation of the joint distribution.
 
-\newpage
+\begin{IEEEeqnarray}{rCl}
+p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K})
+\end{IEEEeqnarray}
 
-\begin{itemize}
-	\item Kitigawa smoother
-	\item Doucet re-weighting smoother
-	\item Godsill re-sampling smoother
-	\item Complexities
-	\item Algorithms
-\end{itemize}
+Particles are generated by sampling backwards in time from the factors of this expansion.
+At the $k^{th}$ step, a sample $\tilde{x}_{k+1:K} \sim p(x_{k+1:K}|y_{1:K})$ will already have been drawn.
+By sampling $x_k \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$ and appending $x_k$ to $\tilde{x}_{k+1:K}$, the state sequence is sequentially extended back in time.
+The recursion is initialised with $x_K \sim p(x_K|y_{1:K})$, i.e.
+a sample from the final filtering approximation.
 
-\section{New MCMC Particle Smoother}
+The backwards conditional distributions may be expanded with Bayes rule.
+
+\begin{IEEEeqnarray}{rCl}
+p(x_k|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_k|\tilde{x}_{k+1}, y_{1:k}) \nonumber \\
+                                & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|y_{1:k})
+\end{IEEEeqnarray}
+
+Substituting the filtering approximation into this expression yields a particle distribution which may be sampled easily.
+
+\begin{IEEEeqnarray}{rCl}
+\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k}) \\
+\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:DBRS_weights}
+\end{IEEEeqnarray}
+
+The procedure may be repeated to produce as many samples from the smoother distribution as required. This backward-resampling smoother is summarised in algorithm~\ref{alg:DBRS}.
+
+\begin{algorithm}
+  \begin{algorithmic}
+  	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$.
+		\FOR{$i = 1 \dots N_S$}
+			\STATE Sample $\tilde{x}_{K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{K}^{(j)}}(x_{K})$.
+			\FOR{$k = K-1 \dots 1$}
+				\FOR{$j = 1 \dots N_F$}
+					\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
+				\ENDFOR
+				\STATE Sample $\tilde{x}_{k}^{(i)} \sim \sum_j \tilde{w}_k^{(j)} \delta_{x_{k}^{(j)}}(x_{k})$.
+			\ENDFOR
+		\ENDFOR
+  \end{algorithmic}
+  \caption{Direct backward-resampling smoother algorithm}
+  \label{alg:DBRS}
+\end{algorithm}
+
+A third method for drawing samples from the joint smoothing distribution, $p(x_{1:K}|y_{1:K})$, is included in \cite{Briers2010}.
+The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K}$, by combination of the results from one forward and one backward particle filter.
+Particles from the joint distribution may then be generated by resampling from either the forward or backward filters in the manner of algorithm~\ref{alg:DBRS}.
+
+The complexity of both the backward-resampling smoother of \cite{Godsill2004} and the two-filter variant of \cite{Briers2010} is $\mathcal{O}(N_F \times N_S \times K)$, where $N_F$ is the number of filter particles, $N_S$ the number of smoother particles, and $K$ the number of time steps.
+Methods for approximation of such algorithms with lower complexity are discussed in \cite{Klaas2006}.
+However, computational cost is often prohibitive, requiring either $N_S$ or $N_F$ to be held very low.
+
+In the algorithms described so far, no new states are sampled during the smoothing stage; there is only a resampling of the filtering distributions.
+This limits the support of the smoother to the states that appear in the filtering particles.
+In \cite{Fearnhead2010}, an algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal, allowing full support.
+In addition, this algorithm achieves $\mathcal{O}( N_S \times K)$ complexity, linear in the number of particles!
+
+In section~\ref{sec:mcmc_smoother}, we introduce an MCMC approach for drawing samples from the joint smoothing distribution with an improved computational complexity over the direct backward-resampling smoother.
+In section~\ref{sec:new_state_smoother}, we combine this procedure with the new-proposal method used in \cite{Fearnhead2010} to generate joint smoothing distribution particles with full support.
+
+\section{New MCMC Particle Smoother} \label{sec:mcmc_smoother}
+
+The bottleneck of the backward-resampling smoother (algorithm~\ref{alg:DBRS}) is the calculation of the sampling weights (equation~\ref{eq:DBRS_weights}).
+For each smoother particle and for each time step, a weight must be calculated for each of the $N_F$ filter particles, in order to draw a sample from the conditional distribution $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
+The innovation here is to use Markov chain Monte Carlo (MCMC) to produce these samples instead of sampling directly.
+
+MCMC algorithms are a class of numerical procedures which use a Markov chain to draw dependent samples from a probability distribution.
+A thorough introduction can be found in \cite{Gilks1996}.
+The basic Metropolis-Hastings (MH) \cite{Metropolis1953} sampler uses a series of steps in which a new state is drawn from a proposal distribution and accepted with a probability determined by the ratios of the target and proposal densities.
+
+For the new MCMC smoother, the target distribution will be the conditional $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
+If the current state of the chain is $x_k^{(m)}$, and a new sample is drawn from the proposal $x_k^{*} \sim q(x_k|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability.
+
+\begin{IEEEeqnarray}{rCl}
+\alpha &=& \frac{ \hat{p}(x_k^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_k^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_k^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_k^{*}|\tilde{x}_{k+1}, y_{1:K}) }
+\end{IEEEeqnarray}
+
+A valid proposal distribution may be constructed using the particles of the filter approximation with any arbitrary weights.
+
+\begin{IEEEeqnarray}{rCl}
+q(x_k|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_j \tilde{v}_k \delta_{x_{k}^{(j)}}(x_{k})
+\end{IEEEeqnarray}
+
+The resulting acceptance probability simplifies to the following:
+
+\begin{IEEEeqnarray}{rCl}
+\alpha &=& \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }
+\end{IEEEeqnarray}
+
+Note that $w_k^{*}$ indicates the particle weight of the proposed new state, $x_k^{*}$.
+
+The final state of each chain is used as a sample from the required conditional distribution.
+The complete MCMC backward-resampling smoother is summarised in algorithm~\ref{alg:MCMC-BRS}.
+Here we target the entire history conditional distribution, $\hat{p}(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, which simpifies the initialisation of chains.
+This makes no practical difference to the algorithm.
+
+\begin{algorithm}
+  \begin{algorithmic}
+  	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
+		\FOR{$i = 1 \dots N_S$}
+			\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
+			\FOR{$k = K-1 \dots 1$}
+				\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
+				\FOR{$m = 1 \dots M$}
+					\STATE Propose a new history, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
+					\STATE With probability $\alpha = \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }$, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$.
+Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
+				\ENDFOR
+				\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
+			\ENDFOR
+		\ENDFOR
+  \end{algorithmic}
+  \caption{MCMC backward-resampling smoother algorithm}
+  \label{alg:MCMC-BRS}
+\end{algorithm}
+
+The computational complexity of the MCMC-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used at each time step and smoothing particle.
+The advantage of the new procedure is that $M$ can often be much smaller $N_F$ and yet achieve the same error performance.
 
@@
 	\item complexity
+	\item See direct simulation smoother as a single step of Gibbs Sampler
 \end{itemize}
+
+\newpage
 
@@
 
-\section{Simulations}
+\section{Improving the smoother support} \label{sec:new_state_smoother}
+
+\begin{itemize}
+	\item Outline Fearnhead smoother
+	\item Discuss extension to joint estimation (possibility of IS)
+	\item MCMC method
+\end{itemize}
+
+\section{Simulations} \label{sec:simulations}
 
@@
 
-\section{Conclusion}
+\section{Conclusion} \label{sec:conclusions}
 The conclusion goes here.
@@
 \appendices
-\section{Proof of the First Zonklar Equation}
+\section{Appendix Title}
 Appendix one text goes here.

### 2012-01-23 MATLAB/ESS.m

@@ -1,0 +1,6 @@
+function ess = ESS( weight )
+%CALCESS Calculates effective sample size from array of (logarithmic) weights
+weight = exp(weight);
+ess = 1/( sum( weight.^2 ) );
+end
+

### 2012-01-23 MATLAB/RMSE.m

@@ -1,0 +1,21 @@
+function [ rmse ] = RMSE( x, pts )
+%RMSE Calculate the root-mean-squared error between the mean particle
+%estimate and the true values
+
+[d, K] = size(x);
+
+rmse.pos = zeros(1, K);
+rmse.vel = zeros(1, K);
+
+pts_x = cell2mat(permute(pts, [3,2,1]));
+est_x = mean(pts_x, 3);
+error = bsxfun(@minus, est_x, x);
+
+rmse.pos = sqrt(sum(error(1:2,:).^2,1));
+rmse.vel = sqrt(sum(error(3:4,:).^2,1));
+
+rmse.mean_pos = sqrt(mean(rmse.pos.^2));
+rmse.mean_vel = sqrt(mean(rmse.vel.^2));
+
+end
+

### 2012-01-23 MATLAB/backard_sampling_smoother.m

@@ -1,0 +1,61 @@
+function [ smooth_pts ] = backard_sampling_smoother( S, times, pts_array, wts_array, h_trans )
+%BACKARD_SAMPLING_SMOOTHER Normal Godsill2004-type backward sampling
+%smoother.
+%
+%
+%
+
+fprintf(1, '\n\n*** Running direct backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Get weights
+        wts = wts_array{kk};
+        
+        % Loop through filter particles
+        for jj = 1:length(pts_array{kk})
+            
+            % Calculate weight
+            [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{jj}(:,kk), smooth_pts{ii}(:,kk+1) );
+            wts(jj) = wts(jj) + trans_prb;
+            
+        end
+        
+        % Normalise weights
+        wts = wts - max(wts);
+        lin_wts = exp(wts); lin_wts = lin_wts/sum(lin_wts);
+        wts = log(lin_wts);
+        
+        % Sample
+%         fprintf(1, '     ESS in frame %u: %f\n', kk, ESS(wts));
+        ind = randsample(length(wts), 1, true, exp(wts));
+        
+        % Update trajectory
+        smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+        
+    end
+    
+end
+
+end
+

### 2012-01-23 MATLAB/batch_analysis.m

@@ -1,0 +1,30 @@
+%% Preliminaries
+
+% Set test flag
+test_flag = 2;
+
+% Load data
+load(['smoother_test' num2str(test_flag) '.mat']);
+
+% 
+t = params.dt:params.dt:params.dt*params.N;
+
+%% Output results
+
+figure(3), clf, hold on
+plot(t, results.kiti.unique_pts, 'r'), plot(t, results.bs.unique_pts, 'b')
+plot(t, results.mcmc1.unique_pts, 'g'), plot(t, results.mcmc3.unique_pts, 'g'), plot(t, results.mcmc10.unique_pts, 'g'), plot(t, results.mcmc30.unique_pts, 'g'), plot(t, results.mcmc100.unique_pts, 'g')
+legend('Kitigawa smoother', 'direct smoother', 'MCMC smoother');
+xlabel('time'), ylabel('num. of unique particles')
+
+fprintf(1, '\n');
+fprintf(1, 'Mean num of unique pts: |  KitiSm  |  DireSm  |  MCMCS1  |  MCMCS3  |  MCMCS10 |  MCMCS30 | MCMCS100 |\n');
+fprintf(1, '                           %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |\n', mean(results.kiti.unique_pts), mean(results.bs.unique_pts), mean(results.mcmc1.unique_pts), mean(results.mcmc3.unique_pts), mean(results.mcmc10.unique_pts), mean(results.mcmc30.unique_pts), mean(results.mcmc100.unique_pts));
+fprintf(1, '\n');
+fprintf(1, 'RMSEs:  |  Filter  |  KitiSm  |  DireSm  |  MCMCS1  |  MCMCS3  |  MCMCS10 |  MCMCS30 | MCMCS100 |\n');
+fprintf(1, 'Position:  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |\n', results.filt.mean_pos_rmse, results.kiti.mean_pos_rmse, results.bs.mean_pos_rmse, results.mcmc1.mean_pos_rmse, results.mcmc3.mean_pos_rmse, results.mcmc10.mean_pos_rmse, results.mcmc30.mean_pos_rmse, results.mcmc100.mean_pos_rmse);
[…]

### 2012-01-23 MATLAB/batch_smoother_testing.m

@@ -1,0 +1,198 @@
+clup
+dbstop if error
+
+% Get test flag
+% test_flag = str2double(getenv('SGE_TASK_ID'));
+test_flag = 2;
+
+% Batch test parameters
+num_seed = 10;
+set_parameters;
+
+% Results arrays
+bs_times = zeros(num_seed,1);
+mcmc1_times = zeros(num_seed,1);
+mcmc3_times = zeros(num_seed,1);
+mcmc10_times = zeros(num_seed,1);
+mcmc30_times = zeros(num_seed,1);
+mcmc100_times = zeros(num_seed,1);
+
+kiti_unique_pts = zeros(num_seed,params.N);
+bs_unique_pts = zeros(num_seed,params.N);
+mcmc1_unique_pts = zeros(num_seed,params.N);
+mcmc3_unique_pts = zeros(num_seed,params.N);
+mcmc10_unique_pts = zeros(num_seed,params.N);
+mcmc30_unique_pts = zeros(num_seed,params.N);
+mcmc100_unique_pts = zeros(num_seed,params.N);
+
+filt_mean_pos_rmse = zeros(num_seed,1);
+kiti_mean_pos_rmse = zeros(num_seed,1);
+bs_mean_pos_rmse = zeros(num_seed,1);
+mcmc1_mean_pos_rmse = zeros(num_seed,1);
+mcmc3_mean_pos_rmse = zeros(num_seed,1);
+mcmc10_mean_pos_rmse = zeros(num_seed,1);
+mcmc30_mean_pos_rmse = zeros(num_seed,1);
+mcmc100_mean_pos_rmse = zeros(num_seed,1);
+
+filt_mean_vel_rmse = zeros(num_seed,1);
+kiti_mean_vel_rmse = zeros(num_seed,1);
+bs_mean_vel_rmse = zeros(num_seed,1);
+mcmc1_mean_vel_rmse = zeros(num_seed,1);
+mcmc3_mean_vel_rmse = zeros(num_seed,1);
+mcmc10_mean_vel_rmse = zeros(num_seed,1);
+mcmc30_mean_vel_rmse = zeros(num_seed,1);
+mcmc100_mean_vel_rmse = zeros(num_seed,1);
+
+
+% Loop through random seend
+for rs = 1:num_seed;
+    
+    % Set random seed
+    s = RandStream('mt19937ar', 'seed', rs);
+    RandStream.setDefaultStream(s);
+    
+    % Set parameters
[…]

### 2012-01-23 MATLAB/build_AQ.m

@@ -1,0 +1,18 @@
+function [ A, Q ] = build_AQ( T, proc_var )
+%BUILD_AQ Construct transition matrixes A and Q
+
+d = 0.1;
+
+A = [1 0 T           0          ;
+     0 1 0           T          ;
+     0 0 exp(-d*T) 0          ;
+     0 0 0           exp(-d*T)];
+ 
+Q = proc_var * ...
+    [T^3/3  0      T^2/2  0    ;
+     0      T^3/3  0      T^2/2;
+     T^2/2  0      T      0    ;
+     0      T^2/2  0      T    ];
+
+end
+

### 2012-01-23 MATLAB/mcmc_newstate_smoother.m

@@ -1,0 +1,86 @@
+function [ smooth_pts ] = mcmc_newstate_smoother( S, M, times, pts_array, wts_array, observs, h_trans, h_obs, h_bidirec_ppsl )
+%MCMC_NEWSTATE_SMOOTHER 
+%
+%
+%
+
+fprintf(1, '\n\n*** Running MCMC-based backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+Np = size(pts_array{K},1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:2
+        
+        dt_next = times(kk+1) - times(kk);
+        dt_prev = times(kk) - times(kk-1);
+        
+        old_x = smooth_pts{ii}(:,kk);
+        
+        % Initialise probability
+        [~, old_trans_prb_prev] = feval(h_trans, dt_prev, smooth_pts{ii}(:,kk-1), old_x );
+        [~, old_trans_prb_next] = feval(h_trans, dt_next, old_x, smooth_pts{ii}(:,kk+1) );
+        [~, old_obs_prb] = feval(h_obs, smooth_pts{ii}(:,kk), observs(:,kk));
+        [~, old_ppsl_prb] = feval(h_bidirec_ppsl, smooth_pts{ii}(:,kk-1), smooth_pts{ii}(:,kk+1), observs(:,kk), dt_prev, dt_next, h_trans, h_obs, smooth_pts{ii}(:,kk));
+        
+%         ind = 1;
+%         while smooth_pts{ii}(1,kk-1)~=pts_array{kk-1}{ind}(1,kk-1)
+%             ind = ind + 1;
+%         end
+%         old_weight = wts_array{kk-1}(ind);
+        
+        for jj = 1:M
+            
+            % Propose a new k-1 particle
+%             ind = randsample(Np, 1);
+            ind = randsample(Np, 1, true, exp(wts_array{kk-1}));
+            
+            % Propose a new k state
[…]

### 2012-01-23 MATLAB/mcmc_smoother.m

@@ -1,0 +1,68 @@
+function [ smooth_pts ] = mcmc_smoother( S, M, times, pts_array, wts_array, h_trans )
+%MCMC_SMOOTHER New MCMC-based backward sampling smoother
+%
+%
+%
+
+fprintf(1, '\n\n*** Running MCMC-based backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+Np = size(pts_array{K},1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Initialise probability
+        [~, old_trans_prb] = feval(h_trans, dt, smooth_pts{ii}(:,kk), smooth_pts{ii}(:,kk+1) );
+%         ind = 1;
+%         while smooth_pts{ii}(1,kk)~=pts_array{kk}{ind}(1,kk)
+%             ind = ind + 1;
+%         end
+%         old_weight = wts_array{kk}(ind);
+        
+        for jj = 1:M
+            
+            % Propose a new particle
+%             ind = randsample(Np, 1);
+            ind = randsample(Np, 1, true, exp(wts_array{kk}));
+            
+            % Calculate probability
+            [~, new_trans_prb] = feval(h_trans, dt, pts_array{kk}{ind}(:,kk), smooth_pts{ii}(:,kk+1) );
+            new_weight = wts_array{kk}(ind);
+            
+            % Construct acceptance probability
+            ap = (new_trans_prb - old_trans_prb);%...
+%                 +(new_weight - old_weight);
+            
+            if log(rand)<ap
+                % Update trajectory
+                smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+                old_weight = new_weight;
+                old_trans_prb = new_trans_prb;
+            end
+            
+        end
+        
+    end
+    
+end
+
+end
+

### 2012-01-23 MATLAB/particle_filter.m

@@ -1,0 +1,108 @@
+function [ pts_array, wts_array, filter_pts ] = particle_filter( init_pts, times, observs, h_ppsl, h_trans, h_obs, ess_thresh )
+%PARTICLE_FILTER Run a particle filter
+%
+% *** Input ***
+% init_pts      - Cell array of particles defining prior (see below)
+% times         - Row vector of observation times
+% observs       - Matrix of observations, observs(:,k) is the kth observation
+% h_ppsl        - Function handle to the proposal function.
+% h_trans       - Function handle to the transition likelihood function.
+% h_obs         - Function handle to the observation likelihood function.
+%
+% *** Output ***
+% pts_array     - Cell array of particle cell arrays, the filter output at each observation time
+% wts_array     - Cell array of corresponding weight column vectors
+% filter_pts    - Cell array of equal-weight, current filter points (i.e. no Kitigawa smoothing)
+%
+% Particle cell arrays have the following structure: pts{ii}(:,k) is the
+% state of the iith particle at the kth time.
+%
+% All weights and probabilities are log-ed.
+%
+
+fprintf(1, '\n\n*** Running particle filter. ***\n');
+
+% Initialise constants
+Np = length(init_pts);
+K = size(times, 2);
+d = size(init_pts{1}, 1);
+
+% Initialise arrays
+pts_array = cell(K,1);
+wts_array = cell(K,1);
+filter_pts = repmat({zeros(d,K)}, Np, 1);
+wts = log(ones(Np,1)/Np);
+last_pts = init_pts;
+
+% Loop through time
+for kk = 1:K
+    
+    fprintf(1, 'Now processing frame %u.\n', kk);
+    
+    % Create a new particle array
+    pts = repmat({zeros(d,kk)}, Np, 1);
+    
+    % Find time increment
+    t = times(kk);
+    if kk == 1
+        dt = t;
+    else
+        dt = times(kk) - times(kk-1);
+    end
+    
+    % Loop through particles
+    for ii = 1:Np
+        
+        % Copy state trajectory from old particle
+        pts{ii} = last_pts{ii}(:,1:kk-1);
+        
+        % Get previous state
+        if kk == 1
+            last_x = init_pts{ii}(:,1);
+        else
+            last_x = pts{ii}(:,kk-1);
+        end
+        
[…]

### 2012-01-23 MATLAB/run_mcmcsmoother_demo.m

@@ -1,0 +1,117 @@
+clup
+dbstop if error
+
+% DEFINE RANDOM SEED
+rand_seed = 1;
+
+% Set random seed
+s = RandStream('mt19937ar', 'seed', rand_seed);
+RandStream.setDefaultStream(s);
+
+% Set parameters
+set_parameters;
+
+%% Generate some Bearings only tracking data
+[ t, x, y ] = generate_radar_data;
+
+% Plot it
+xmax = max(max(abs(x(1:2,:))))+5;
+figure(1); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b')
+[x_obs, y_obs] = pol2cart(y(1,:), y(2,:)); plot(x_obs, y_obs, 'r');
+figure(2); plot(t, y);
+
+%% Run a PF
+init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);
+[pts_array, wts_array, filter_pts] = particle_filter( init_pts, t, y, @tracking_ppsl, @tracking_trans, @tracking_obs, 0.5 );
+
+% Resample final frame particles
+parents = systematic_resample(wts_array{end});
+kiti_pts = pts_array{end}(parents);
+
+% Plot filter
+figure(3); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), filter_pts)
+
+% Plot Kitigawa smoother
+figure(4); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), kiti_pts)
+
+%% Run an ordinary smoother
+tic;
+bs_smooth_pts = backard_sampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
+bs_time = toc;
+
+% Plot it
+figure(5); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
[…]

### 2012-01-23 MATLAB/set_parameters.m

@@ -1,0 +1,17 @@
+global params;
+
+%Model
+params.dt = 1;
+params.N = 100;
+params.proc_var = 1;
+params.bng_var = (pi/180)^2;    % Try (pi/720)^2; (pi/180)^2;
+params.rng_var = 10;           % Try 0.1;        10;
+params.x0 = [-100; 50; 10; 0];
+
+% Prior
+params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
+
+% Algorithm
+params.Np = 100;
+params.S = 10;
+params.M = 5;                   % Try 5;           50;

### 2012-01-23 MATLAB/tracking_bidirec_ppsl.m

@@ -1,0 +1,26 @@
+function [x, ppsl_prb] = tracking_bidirec_ppsl(x_prev, x_next, y, dt_prev, dt_next, h_trans, h_obs, x)
+%TRACKING_BIDIREC_PPSL Propose a new state for a 2D tracking model when the
+% next state is already known (for smoother).
+
+global params;
+
+[A_prev, Q_prev] = build_AQ(dt_prev, params.proc_var);
+[A_next, Q_next] = build_AQ(dt_next, params.proc_var);
+R = [params.bng_var 0; 0 params.rng_var];
+
+% Interpolate between previous and next x
+[m, P] = kf_update(A_prev*x_prev, Q_prev, x_next, A_next, Q_next);
+
+% Now include y
+H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
+     m(1)/sqrt(m(1)^2+m(2)^2), m(2)/sqrt(m(1)^2+m(2)^2), 0, 0];
+[m, P] = ekf_update1(m, P, y, H, R, h_obs);
+
+% Sample
+if (nargin < 8)||isempty(x)
+    x = mvnrnd(m', P)';
+end
+ppsl_prb = log(mvnpdf(x', m', P));
+
+end
+

### 2012-01-23 MATLAB/tracking_obs.m

@@ -1,0 +1,16 @@
+function [ y_mn, obs_prb ] = tracking_obs( x, y )
+%TRACKING_OBS Observation likelihood for bearing-only 2D radar observations
+
+global params;
+
+R = [params.bng_var 0; 0 params.rng_var];
+
+[b, r] = cart2pol(x(1), x(2));
+y_mn = [b; r];
+
+if ~isempty(y)
+    obs_prb = log(mvnpdf(y', y_mn', R));
+end
+
+end
+

### 2012-01-23 MATLAB/tracking_ppsl.m

@@ -1,0 +1,22 @@
+function [ x, ppsl_prb ] = tracking_ppsl( last_x, y, dt, h_trans, h_obs)
+%TRACKING_PPSL Propose a new state for a 2D tracking model
+
+global params;
+
+[A, Q] = build_AQ(dt, params.proc_var);
+R = [params.bng_var 0; 0 params.rng_var];
+
+% Optimal importance
+[m, P] = kf_predict(last_x, eps*eye(4), A, Q);
+H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
+     m(1)/sqrt(m(1)^2+m(2)^2), m(2)/sqrt(m(1)^2+m(2)^2), 0, 0];
+[m, P] = ekf_update1(m, P, y, H, R, h_obs);
+x = mvnrnd(m', P)';
+ppsl_prb = log(mvnpdf(x', m', P));
+
+% % Bootstrap
+% x = mvnrnd((A*last_x), Q)';
+% ppsl_prb = log(mvnpdf(x', (A*last_x)', Q));
+
+end
+

### 2012-01-23 MATLAB/tracking_trans.m

@@ -1,0 +1,15 @@
+function [ x_mn, trans_prb ] = tracking_trans( dt, last_x, x )
+%TRACKING_TRANS Transition probability for linear 2D tracking dynamics
+
+global params;
+
+[A, Q] = build_AQ(dt, params.proc_var);
+
+x_mn = A*last_x;
+
+if ~isempty(x)
+    trans_prb = log(mvnpdf(x', x_mn', Q));
+end
+
+end
+

### 2012-01-24 Paper/draft3_MCMCsmoother.tex
First paper draft complete except simulations, conclusions, and abstract. Batch testing re-implemented.

--- 2012-01-23 Paper/draft2_MCMCsmoother.tex
@@
 An improved particle smoothing algorithm was presented in \cite{Godsill2004}.
-This works by constructing new particles from the smoothing distribution by sampling from the filtering distributions.
+This works by constructing new particles from the smoothing distribution by resampling from the filtering approximations.
 This has the limitation that the particle support for the smoother is limited to that of the filter.
@@
 \begin{IEEEeqnarray}{rCl}
-p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K})
+p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K}) \label{eq:smoothing_factorisation}
 \end{IEEEeqnarray}
@@
 \begin{IEEEeqnarray}{rCl}
-\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k}) \\
+\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k}) \label{eq:backward_conditional_filter} \\
 \tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:DBRS_weights}
@@
 
-The procedure may be repeated to produce as many samples from the smoother distribution as required. This backward-resampling smoother is summarised in algorithm~\ref{alg:DBRS}.
+The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised in algorithm~\ref{alg:DBRS}.
 
@@
 
-A third method for drawing samples from the joint smoothing distribution, $p(x_{1:K}|y_{1:K})$, is included in \cite{Briers2010}.
-The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K}$, by combination of the results from one forward and one backward particle filter.
+A method for drawing samples from the joint smoothing distribution, $p(x_{1:K}|y_{1:K})$, is included in \cite{Briers2010}.
+The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K})$, by combination of the results from one forward and one backward particle filter.
 Particles from the joint distribution may then be generated by resampling from either the forward or backward filters in the manner of algorithm~\ref{alg:DBRS}.
@@
 
-MCMC algorithms are a class of numerical procedures which use a Markov chain to draw dependent samples from a probability distribution.
+MCMC algorithms are a class of numerical procedures which use a Markov chain to draw dependent samples from a target probability distribution.
 A thorough introduction can be found in \cite{Gilks1996}.
-The basic Metropolis-Hastings (MH) \cite{Metropolis1953} sampler uses a series of steps in which a new state is drawn from a proposal distribution and accepted with a probability determined by the ratios of the target and proposal densities.
+The basic Metropolis-Hastings (MH) \cite{Metropolis1953} sampler uses a series of steps in which a new state is drawn from a proposal distribution and accepted with a probability determined by the ratios of the target and proposal probabilities.
 
-For the new MCMC smoother, the target distribution will be the conditional $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
-If the current state of the chain is $x_k^{(m)}$, and a new sample is drawn from the proposal $x_k^{*} \sim q(x_k|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability.
+For the direct backward sampling smoother, the objective at each time step was to draw a sample from $p(x_{k}|\tilde{x}_{k+1}, y_{1:K})$.
+To simplify the notation and initialisation of chains, the MCMC version will target $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
+Theoretically this means using the particles of the filter-smoother, $\{x_{1:k}^{(i)}\}$, rather than just the filter, $\{x_k^{(j)}\}$, which would require $\mathcal{O}(K^2)$ memory.
+However, as discussed later, this is not required in practice.
+
+As before, the algorithm begins by first running a particle filter-smoother and then resampling states backwards through time to produce particles from the joint smoothing distribution, using the factorisation of equation~\ref{eq:smoothing_factorisation}.
+At the $k^{th}$ time step, a Markov chain is used to sample from $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
+The chain is initialised with the first $k$ states from the previous processing step.
+As this is itself a sample from the target distribution, no burn-in period is required.
+
+Following an equivalent derivation to equation~\ref{eq:backward_conditional_filter}, the target distribution may be approximated by a particle distribution.
 
 \begin{IEEEeqnarray}{rCl}
-\alpha &=& \frac{ \hat{p}(x_k^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_k^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_k^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_k^{*}|\tilde{x}_{k+1}, y_{1:K}) }
+\hat{p}(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{1:k}^{(i)}}(x_{1:k}) \\
+\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:MCMC-BRS_weights}
 \end{IEEEeqnarray}
 
-A valid proposal distribution may be constructed using the particles of the filter approximation with any arbitrary weights.
+A valid proposal distribution may be constructed using the particles of the filter-smoother approximation with any arbitrary weights.
 
 \begin{IEEEeqnarray}{rCl}
-q(x_k|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_j \tilde{v}_k \delta_{x_{k}^{(j)}}(x_{k})
+q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})
 \end{IEEEeqnarray}
 
-The resulting acceptance probability simplifies to the following:
+If the current state of the chain is $x_{1:k}^{(m)}$, and a new sample is drawn from the proposal $x_{1:k}^{*} \sim q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability:
 
 \begin{IEEEeqnarray}{rCl}
-\alpha &=& \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }
+\alpha &=& \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \nonumber \\
+ &=& \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }  .
 \end{IEEEeqnarray}
 
-Note that $w_k^{*}$ indicates the particle weight of the proposed new state, $x_k^{*}$.
+Note that $w_k^{*}$ indicates the particle weight of the proposed new state, $x_{1:k}^{*}$.
 
 The final state of each chain is used as a sample from the required conditional distribution.
+
 The complete MCMC backward-resampling smoother is summarised in algorithm~\ref{alg:MCMC-BRS}.
-Here we target the entire history conditional distribution, $\hat{p}(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, which simpifies the initialisation of chains.
-This makes no practical difference to the algorithm.
 
@@
 				\FOR{$m = 1 \dots M$}
-					\STATE Propose a new history, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
+					\STATE Propose a new state, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
 					\STATE With probability $\alpha = \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }$, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$.
@@
 				\ENDFOR
-				\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
+				\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{k}^{(i)(M)}$
 			\ENDFOR
@@
 
-The computational complexity of the MCMC-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used at each time step and smoothing particle.
+For a practical implementation, it is not necessary to store the complete particles of the filter-smoother approximations, $\hat{p}(x_{1:k}|y_{1:k})$, which would require memory space which scaled quadratically as $K$ increased.
+As the smoother progresses backwards in time, the earlier states in the sequence are repeatedly overwritten, and so it is not necessary ever to have known them.
+Thus, it is only necessary to store the last two states in each filter-smoother particle (i.e.
+$\{x_{k-1}^{(j)}\}$ and $\{x_{k}^{(j)}\}$).
+
+The computational complexity of the MCMC-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used for each time step and smoothing particle.
 The advantage of the new procedure is that $M$ can often be much smaller $N_F$ and yet achieve the same error performance.
 
-\begin{itemize}
-	\item Motivate - high complexity, low speed of other smoothers
-	\item MCMC avoids need to calculate all weights at each step
-	\item maths
-	\item algorithm
-	\item complexity
-	\item See direct simulation smoother as a single step of Gibbs Sampler
-\end{itemize}
-
-\newpage
-
-\subsection{Performance Control}
-
-\begin{itemize}
-	\item The effect of varying M, the number of MH steps
-	\item Situations where the new smoother works well/poorly.
-	\item Backward sampling weight ESS.
-\end{itemize}
+With the direct backward-resampling smoother, the only parameter to be chosen is $N_S$, the number of smoother particles.
+With the MCMC variant, we also have the freedom to set $M$, the number of MH steps in each Markov chain.
+This controls a trade-off between speed and particle diversity.
+If the chains are short, and the acceptance probabilities low, then this scheme will simply reproduce particles from the filter-smoother.
+However, when the weights of the conditional approximation, equation~\ref{eq:MCMC-BRS_weights} are similar, then an independent sample may be produced with only a very short chain.
 
@@
 
+In \cite{Fearnhead2010}, the authors described an algorithm for estimating the marginal smoothing distributions.
+For this method, states were proposed afresh rather than simply resampling from the filtering approximations.
+This yielded improvements in accuracy and particle diversity over other marginal smoothing techniques, and also the advantage of $\mathcal{O}(N_S \times K)$ complexity.
+Here we adapt the method of \cite{Fearnhead2010} for generating particles from the joint smoothing distribution.
+Again, an MCMC procedure is employed.
+
+The method described in \cite{Fearnhead2010} uses a forward and a backward particle filter.
+To develop a a marginal smoothing approximation at time step $k$, particles are resampled from the forward filter at step $k-1$, and from the backward filter and step $k+1$.
+A new state for step $k$ is then sampled from a continuous proposal distribution conditional on the previous and next states, and an importance weight is assigned.
+It is undesirable to use importance sampling for generating particles from the joint sampling distribution, as the required resampling is liable to lead to a loss of particle diversity similar to that suffered by the filter-smoother.
+
+The proposed new method proceeds in a similar manner to algorithm~\ref{alg:MCMC-BRS}.
+However, this time $x_{1:k-1}$, is resampled from the particle filter-smoother approximation from time step $k-1$.
+The current state, $x_k$, is then sampled afresh from a new proposal distribution.
+Thus the factorised proposal may be written as follows:
+
+\begin{IEEEeqnarray}{rCl}
+q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & = & q(x_{1:k-1}|y_{1:k-1}) q(x_{k}|x_{k-1}, \tilde{x}_{k+1}, y_{k}) \nonumber \\
+                                    & = & \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1}) q(x_{k}|x_{k-1}^{(j)}, \tilde{x}_{k+1}, y_{k})  .
+\end{IEEEeqnarray}
+
+The target distribution is also factorised.
+
+\begin{IEEEeqnarray}{rCl}
+p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) \nonumber \\
+                                    & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k}|y_{1:k-1})
+\end{IEEEeqnarray}
+
+Using the filter-smoother approximation, the MH acceptance probabilities may then be calculated.
+
+\begin{IEEEeqnarray}{rCl}
+\alpha &=& \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \nonumber \\
+ &=& \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \nonumber \\
+ \IEEEeqnarraymulticol{3}{l}{\frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } } .
+\end{IEEEeqnarray}
+
+The complete MCMC backward-sampling smoother is summarised in algorithm~\ref{alg:MCMC-BSS}.
+
+\begin{algorithm}
+  \begin{algorithmic}
+  	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
+		\FOR{$i = 1 \dots N_S$}
+			\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
+			\FOR{$k = K-1 \dots 1$}
+				\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
+				\FOR{$m = 1 \dots M$}
+					\STATE Propose a new state hisory, $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
+					\STATE Propose a new state, $x_{1:k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
+					\STATE With probability $\alpha = \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }$, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$.
+Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
+				\ENDFOR
+				\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{k}^{(i)(M)}$
+			\ENDFOR
+		\ENDFOR
+  \end{algorithmic}
+  \caption{MCMC backward-sampling smoother algorithm}
+  \label{alg:MCMC-BSS}
+\end{algorithm}
+
+The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother.
+Because the support of the smoother particles is not restricted to the states which appear in the filter approximation, the error performance is expected to be improved.
+
+\section{Simulations} \label{sec:simulations}
+
+The performance of the new smoothers was tested on a simple 2D tracking model.
+The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model.
+
+\begin{IEEEeqnarray}{rCl}
+\underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \underbrace{\begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix}}_{\mathbf{A}} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
+\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k} \\
+\end{IEEEeqnarray}
+
+The time between successive time steps is $\Delta t$.
+$I_2$ and $0_2$ denote the $2 \times 2$ identity and zero matrices.
+The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ have a zero-mean Gaussian distribution with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively.
+
+\begin{IEEEeqnarray}{rCl}
+\mathbf{Q} &=& \sigma^2_P \begin{bmatrix}\frac{\Delta t^3}{3} I_2 & \frac{\Delta t^2}{2} I_2 \\ \frac{\Delta t^2}{2} I_2 & \Delta t I_2 \end{bmatrix} \\
+\mathbf{R} &=& \begin{bmatrix}\sigma^2_B & 0 \\ 0 & \sigma^2_R\end{bmatrix}
+\end{IEEEeqnarray}
+
+The following algorithms were all implemented for comparison in MATLAB:
+
 \begin{itemize}
-	\item Outline Fearnhead smoother
-	\item Discuss extension to joint estimation (possibility of IS)
-	\item MCMC method
+	\item The simple filter-smoother of \cite{Kitagawa1996}
+	\item The direct backward-resampling smoother of \cite{Godsill2004}
+	\item The new MCMC backward resampling smoother of section~\ref{sec:mcmc_smoother}, with several values of $M$
+	\item The new MCMC backward sampling smoother of section~\ref{sec:new_state_smoother}, with several values of $M$
 \end{itemize}
 
-\section{Simulations} \label{sec:simulations}
+The algorithms were all used to generate XX particles from the joint smoothing distribution, using particle filters with XX particles.
+Where appropriate, proposal weights were chosen to be equal to filtering weights, i.e.
+$v_k^{(j)} = w_k^{(j)}$.
+
+XX realisations were simulated from the model and the algorithms tested on each one with XX different random seeds.
+
+Tables~\ref{} and~\ref{} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms for two different settings of $\mathbf{Q}$ and $\mathbf{R}$, while tables~\ref{} and~\ref{} show the corresponding running times.
+The trade-off between speed and error is shown in figure~\ref{}.
+The mean number of particles over time is shown in figure~\ref{}.
 

### 2012-01-24 MATLAB/batch_analysis.m

@@ -1,30 +1,47 @@
 %% Preliminaries
 
+close all
+
 % Set test flag
-test_flag = 2;
+test_flag = 1;
 
 % Load data
 load(['smoother_test' num2str(test_flag) '.mat']);
 
 % 
+NA = 13;
 t = params.dt:params.dt:params.dt*params.N;
+col = {[1 0 0], [.5 0 0], [0 1 0], [0 .5 0], [0 0 1], [0 0 .5]};
 
 %% Output results
 
-figure(3), clf, hold on
-plot(t, results.kiti.unique_pts, 'r'), plot(t, results.bs.unique_pts, 'b')
-plot(t, results.mcmc1.unique_pts, 'g'), plot(t, results.mcmc3.unique_pts, 'g'), plot(t, results.mcmc10.unique_pts, 'g'), plot(t, results.mcmc30.unique_pts, 'g'), plot(t, results.mcmc100.unique_pts, 'g')
-legend('Kitigawa smoother', 'direct smoother', 'MCMC smoother');
+% Unique particles over time
+figure(1), hold on
+to_plot = [2,3,4,6,8,10];
+for ii = 1:length(to_plot)
+    alg = to_plot(ii);
+    plot(t, results{alg}.unique_pts, 'color', col{ii});
+end
+legend('filter-smoother', 'direct resampling smoother', 'MCMC resampling smoother (M=1)', 'MCMC resampling smoother (M=10)', 'MCMC sampling smoother (M=1)', 'MCMC sampling smoother (M=10)');
 xlabel('time'), ylabel('num. of unique particles')
 
-fprintf(1, '\n');
-fprintf(1, 'Mean num of unique pts: |  KitiSm  |  DireSm  |  MCMCS1  |  MCMCS3  |  MCMCS10 |  MCMCS30 | MCMCS100 |\n');
[…]

### 2012-01-24 MATLAB/batch_smoother_testing.m

@@ -9,40 +9,29 @@ test_flag = 2;
 num_seed = 10;
 set_parameters;
 
+% Set test-specific parameters
+if test_flag == 1
+    params.bng_var = (pi/720)^2;
+    params.rng_var = 0.1;
+elseif test_flag == 2
+    params.bng_var = (pi/180)^2;
+    params.rng_var = 10;
+end
+
+% Number of smoothing algorithms
+NA = 13; % filter, kitigawa, godsill, 5 MCMC-resampling, 5 MCMC-Fearnhead
+
 % Results arrays
-bs_times = zeros(num_seed,1);
-mcmc1_times = zeros(num_seed,1);
-mcmc3_times = zeros(num_seed,1);
-mcmc10_times = zeros(num_seed,1);
-mcmc30_times = zeros(num_seed,1);
-mcmc100_times = zeros(num_seed,1);
-
-kiti_unique_pts = zeros(num_seed,params.N);
-bs_unique_pts = zeros(num_seed,params.N);
-mcmc1_unique_pts = zeros(num_seed,params.N);
-mcmc3_unique_pts = zeros(num_seed,params.N);
-mcmc10_unique_pts = zeros(num_seed,params.N);
-mcmc30_unique_pts = zeros(num_seed,params.N);
-mcmc100_unique_pts = zeros(num_seed,params.N);
-
-filt_mean_pos_rmse = zeros(num_seed,1);
-kiti_mean_pos_rmse = zeros(num_seed,1);
-bs_mean_pos_rmse = zeros(num_seed,1);
-mcmc1_mean_pos_rmse = zeros(num_seed,1);
-mcmc3_mean_pos_rmse = zeros(num_seed,1);
-mcmc10_mean_pos_rmse = zeros(num_seed,1);
-mcmc30_mean_pos_rmse = zeros(num_seed,1);
-mcmc100_mean_pos_rmse = zeros(num_seed,1);
-
-filt_mean_vel_rmse = zeros(num_seed,1);
-kiti_mean_vel_rmse = zeros(num_seed,1);
-bs_mean_vel_rmse = zeros(num_seed,1);
-mcmc1_mean_vel_rmse = zeros(num_seed,1);
-mcmc3_mean_vel_rmse = zeros(num_seed,1);
-mcmc10_mean_vel_rmse = zeros(num_seed,1);
-mcmc30_mean_vel_rmse = zeros(num_seed,1);
[…]

### 2012-01-24 MATLAB/mcmc_newstate_smoother.m

@@ -4,7 +4,7 @@ function [ smooth_pts ] = mcmc_newstate_smoother( S, M, times, pts_array, wts_ar
 %
 %
 
-fprintf(1, '\n\n*** Running MCMC-based backward sampling smoother. ***\n');
+fprintf(1, '\n\n*** Running MCMC-based backward new-state sampling smoother. ***\n');
 
 % Initialise variables
 K = size(pts_array,1);

### 2012-01-24 MATLAB/mcmc_smoother.m

@@ -4,7 +4,7 @@ function [ smooth_pts ] = mcmc_smoother( S, M, times, pts_array, wts_array, h_tr
 %
 %
 
-fprintf(1, '\n\n*** Running MCMC-based backward sampling smoother. ***\n');
+fprintf(1, '\n\n*** Running MCMC-based backward resampling smoother. ***\n');
 
 % Initialise variables
 K = size(pts_array,1);

### 2012-01-24 MATLAB/set_parameters.m

@@ -4,8 +4,8 @@ global params;
 params.dt = 1;
 params.N = 100;
 params.proc_var = 1;
-params.bng_var = (pi/180)^2;    % Try (pi/720)^2; (pi/180)^2;
-params.rng_var = 10;           % Try 0.1;        10;
+params.bng_var = (pi/720)^2;    % Try (pi/720)^2; (pi/180)^2;
+params.rng_var = 0.1;           % Try 0.1;        10;
 params.x0 = [-100; 50; 10; 0];
 
 % Prior
@@ -13,5 +13,5 @@ params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
 
 % Algorithm
 params.Np = 100;
-params.S = 10;
+params.S = 100;
 params.M = 5;                   % Try 5;           50;

### 2012-01-25 Paper/draft3_MCMCsmoother.tex
Complete first draft. Minor changes to plotting.

@@
 
-\title{Faster particle smoothing using Markov chain Monte Carlo sampling}
+\title{Improved Particle Approximations \\ to the Joint Smoothing Distribution \\ Using Markov Chain Monte Carlo}
 
@@
 \begin{abstract}
-The abstract goes here.
+Particle filtering and smoothing algorithms approximate posterior state distributions with a set of samples drawn from those distributions.
+Conventionally, samples from the joint smoothing distribution are generated by sequentially resampling from the particle filter results.
+If the number of filtering particles is high, this process is limited by computational complexity.
+In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation.
+In this paper, a Markov chain Monte Carlo sampling procedure is used to improve the efficiency of the particle smoother.
+In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-
+state space model, particle filter, smoothing, MCMC, Bayesian inference
 \end{IEEEkeywords}
@@
 Methods for generating these samples have been described in \cite{Kitagawa1996,Godsill2004,Briers2010}.
-In this paper new methods are presented for drawing samples from the smoothing distribution, using Markov chain Monte Carlo (MCMC).
+In this paper new methods are presented for drawing samples from the joint smoothing distribution, using Markov chain Monte Carlo (MCMC).
 These novel procedures have computational advantages and greater flexibility over previous methods.
 
-In section~\ref{sec:basics}, we review briefly review the basic particle filter and existing joint smoother algorithms.
+In section~\ref{sec:basics}, we briefly review the basic particle filter and existing joint smoother algorithms.
 The new MCMC-based smoothing methods are presented in sections~\ref{sec:mcmc_smoother} and~\ref{sec:new_state_smoother} with supporting simulations in section~\ref{sec:simulations}.
@@
 
-The particle filter is summarised in algorithm~\ref{alg:PF}.
+The particle filter is summarised below.%in algorithm~\ref{alg:PF}.
 
-\begin{algorithm}
+%\begin{algorithm}
 \begin{algorithmic}
@@
 \end{algorithmic}
-  \caption{Particle filter algorithm}
-  \label{alg:PF}
-\end{algorithm}
+%\caption{Particle filter algorithm}
+%\label{alg:PF}
+%\end{algorithm}
 
@@
 
-The filter-smoother is summarised in algorithm~\ref{alg:FS}.
+The filter-smoother is summarised below.%in algorithm~\ref{alg:FS}.
 
-\begin{algorithm}
+%\begin{algorithm}
 \begin{algorithmic}
@@
  \end{algorithmic}
-  \caption{Filter-smoother algorithm}
-  \label{alg:FS}
-\end{algorithm}
+%\caption{Filter-smoother algorithm}
+%\label{alg:FS}
+%\end{algorithm}
 
@@
 This works by constructing new particles from the smoothing distribution by resampling from the filtering approximations.
-This has the limitation that the particle support for the smoother is limited to that of the filter.
 The sampling procedure exploits the following factorisation of the joint distribution.
@@
 The recursion is initialised with $x_K \sim p(x_K|y_{1:K})$, i.e.
-a sample from the final filtering approximation.
+a sample from the final filtering distribution.
 
@@
 
-The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised in algorithm~\ref{alg:DBRS}.
+The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised below.% in algorithm~\ref{alg:DBRS}.
 
-\begin{algorithm}
+%\begin{algorithm}
 \begin{algorithmic}
@@
 \end{algorithmic}
-  \caption{Direct backward-resampling smoother algorithm}
-  \label{alg:DBRS}
-\end{algorithm}
+%\caption{Direct backward-resampling smoother algorithm}
+%\label{alg:DBRS}
+%\end{algorithm}
 
@@
 The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K})$, by combination of the results from one forward and one backward particle filter.
-Particles from the joint distribution may then be generated by resampling from either the forward or backward filters in the manner of algorithm~\ref{alg:DBRS}.
+Particles from the joint distribution may then be generated by resampling from either the forward or backward filters.% in the manner of algorithm~\ref{alg:DBRS}.
 
@@
 
-In section~\ref{sec:mcmc_smoother}, we introduce an MCMC approach for drawing samples from the joint smoothing distribution with an improved computational complexity over the direct backward-resampling smoother.
+%In section~\ref{sec:mcmc_smoother}, we introduce an MCMC approach for drawing samples from the joint smoothing distribution with an improved computational complexity over the direct backward-resampling smoother.
 In section~\ref{sec:new_state_smoother}, we combine this procedure with the new-proposal method used in \cite{Fearnhead2010} to generate joint smoothing distribution particles with full support.
@@
 
-The bottleneck of the backward-resampling smoother (algorithm~\ref{alg:DBRS}) is the calculation of the sampling weights (equation~\ref{eq:DBRS_weights}).
+The bottleneck of the backward-resampling smoother is the calculation of the sampling weights (equation~\ref{eq:DBRS_weights}).
 For each smoother particle and for each time step, a weight must be calculated for each of the $N_F$ filter particles, in order to draw a sample from the conditional distribution $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
@@
 
-The complete MCMC backward-resampling smoother is summarised in algorithm~\ref{alg:MCMC-BRS}.
+The complete MCMC backward-resampling smoother is summarised below.% in algorithm~\ref{alg:MCMC-BRS}.
 
-\begin{algorithm}
+%\begin{algorithm}
 \begin{algorithmic}
@@
 \end{algorithmic}
-  \caption{MCMC backward-resampling smoother algorithm}
-  \label{alg:MCMC-BRS}
-\end{algorithm}
+%\caption{MCMC backward-resampling smoother algorithm}
+%\label{alg:MCMC-BRS}
+%\end{algorithm}
 
@@
 The computational complexity of the MCMC-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used for each time step and smoothing particle.
-The advantage of the new procedure is that $M$ can often be much smaller $N_F$ and yet achieve the same error performance.
+The advantage of the new procedure is that $M$ can often be much smaller than $N_F$ and yet achieve almost equal error performance.
 
@@
 If the chains are short, and the acceptance probabilities low, then this scheme will simply reproduce particles from the filter-smoother.
-However, when the weights of the conditional approximation, equation~\ref{eq:MCMC-BRS_weights} are similar, then an independent sample may be produced with only a very short chain.
+However, when the weights of the conditional distribution approximation (equation~\ref{eq:MCMC-BRS_weights}) are similar, then an independent sample may be produced with only a very short chain.
 
@@
 The method described in \cite{Fearnhead2010} uses a forward and a backward particle filter.
-To develop a a marginal smoothing approximation at time step $k$, particles are resampled from the forward filter at step $k-1$, and from the backward filter and step $k+1$.
+To develop a marginal smoothing approximation at time step $k$, particles are resampled from the forward filter at step $k-1$, and from the backward filter and step $k+1$.
 A new state for step $k$ is then sampled from a continuous proposal distribution conditional on the previous and next states, and an importance weight is assigned.
@@
 
-The proposed new method proceeds in a similar manner to algorithm~\ref{alg:MCMC-BRS}.
+The proposed new method proceeds in a similar manner to the MCMC backward-resampling algorithm.
 However, this time $x_{1:k-1}$, is resampled from the particle filter-smoother approximation from time step $k-1$.
@@
 
-The complete MCMC backward-sampling smoother is summarised in algorithm~\ref{alg:MCMC-BSS}.
+The complete MCMC backward-sampling smoother is summarised below.% in algorithm~\ref{alg:MCMC-BSS}.
 
-\begin{algorithm}
+%\begin{algorithm}
  \begin{algorithmic}
@@
 \end{algorithmic}
-  \caption{MCMC backward-sampling smoother algorithm}
-  \label{alg:MCMC-BSS}
-\end{algorithm}
+%\caption{MCMC backward-sampling smoother algorithm}
+%\label{alg:MCMC-BSS}
+%\end{algorithm}
 
@@
 \begin{itemize}
-	\item The simple filter-smoother of \cite{Kitagawa1996}
-	\item The direct backward-resampling smoother of \cite{Godsill2004}
-	\item The new MCMC backward resampling smoother of section~\ref{sec:mcmc_smoother}, with several values of $M$
-	\item The new MCMC backward sampling smoother of section~\ref{sec:new_state_smoother}, with several values of $M$
+	\item The simple filter-smoother (FS) of \cite{Kitagawa1996}
+	\item The direct backward-resampling smoother (DBRS) of \cite{Godsill2004}
+	\item The new MCMC backward resampling smoother (MCMC-BRS) of section~\ref{sec:mcmc_smoother}, with several values of $M$
+	\item The new MCMC backward sampling smoother (MCMC-BSS) of section~\ref{sec:new_state_smoother}, with several values of $M$
 \end{itemize}
 
-The algorithms were all used to generate XX particles from the joint smoothing distribution, using particle filters with XX particles.
+The algorithms were all used to generate 100 particles from the joint smoothing distribution, using particle filters with 100 particles.
 Where appropriate, proposal weights were chosen to be equal to filtering weights, i.e.
 $v_k^{(j)} = w_k^{(j)}$.
+The particle filter and MCMC-BSS both use ``optimal importance distributions'' (see e.g.
+\cite{Cappe2007}) with linearisations of the observation model where appropriate.
 
-XX realisations were simulated from the model and the algorithms tested on each one with XX different random seeds.
+10 realisations of 500 time steps were simulated from the model and the algorithms tested on each one.
+Results are shown here for high and low observation noise cases.
+For case 1, the observation variances were set to $\sigma_B = \pi/720$ and $\sigma_R = 0.1$.
+For case 2, $\sigma_B = \pi/36$ and $\sigma_R = 100$.
+Other model settings are $\Delta t = 1$, $\sigma_P = 1$, and $x_0 = [-100, 50, 10, 0]^T$.
+Example tracjectories and observations from the two cases are shown in figure~\ref{fig:example_trajectories}.
 
-Tables~\ref{} and~\ref{} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms for two different settings of $\mathbf{Q}$ and $\mathbf{R}$, while tables~\ref{} and~\ref{} show the corresponding running times.
-The trade-off between speed and error is shown in figure~\ref{}.
-The mean number of particles over time is shown in figure~\ref{}.
+\begin{figure}
+\centerline{\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory-crop.pdf}%
+\label{fig_first_case}}
+\hfil
+\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory-crop.pdf}%
+\label{fig_second_case}}}
+\caption{Example trajectory with observations from the two parameter cases.}
+\label{fig:example_trajectories}
+\end{figure}
+
+Tables~\ref{tab:case1_performance} and~\ref{tab:case1_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms with the two parameter cases, along with running times (excluding the time for filtering) and the mean number of unique particles at each time instant.
+The trade-off between speed and error is shown in figures~\ref{fig:case1_rmse_vs_time} and~\ref{fig:case2_rmse_vs_time}.
+
+\begin{table}%
+\begin{tabular}{|c||c|c|c|c|}
+\hline
+Algorithm & \begin{minipage}[c]{1cm} Postion RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
+\hline
+FS 						& 0.578 & 0.968 & 0 & 2.25 \\
+DBRS					& 0.475 & 0.762 & 122 & 20.3 \\
+\hline
+MCMC-BRS (1)	& 0.509 & 0.823 & 3 & 13.7 \\
+MCMC-BRS (3)	& 0.493 & 0.790 & 7 & 17.4 \\
+MCMC-BRS (10)	& 0.482 & 0.768 & 21 & 19.6 \\
+MCMC-BRS (30)	& 0.480 & 0.761 & 62 & 20.2 \\
+MCMC-BRS (100)& 0.476 & 0.764 & 202 & 20.4 \\
+\hline
+MCMC-BSS (1)	& 0.473 & 0.758 & 14 & 44.3 \\
+MCMC-BSS (3)	& 0.461 & 0.734 & 29 & 69.7 \\
+MCMC-BSS (10)	& 0.451 & 0.720 & 82 & 89.8 \\
+MCMC-BSS (30)	& 0.444 & 0.713 & 232 & 96.7 \\
+MCMC-BSS (100)& 0.443 & 0.709 & 755 & 98.6 \\
+\hline
+\end{tabular}
+\caption{Smoother performance measures for parameter case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
+\label{tab:case1_performance}
+\end{table}
+
+\begin{table}%
+\begin{tabular}{|c||c|c|c|c|}
+\hline
+Algorithm & \begin{minipage}[c]{1cm} Postion RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
+\hline
+FS 						& 0. & 0. & 0 &  \\
+DBRS					& 0. & 0. &  &  \\
+\hline
+MCMC-BRS (1)	& 0. & 0. &  &  \\
+MCMC-BRS (3)	& 0. & 0. &  &  \\
+MCMC-BRS (10)	& 0. & 0. &  &  \\
+MCMC-BRS (30)	& 0. & 0. &  &  \\
+MCMC-BRS (100)& 0. & 0. &  &  \\
+\hline
+MCMC-BSS (1)	& 0. & 0. &  &  \\
+MCMC-BSS (3)	& 0. & 0. &  &  \\
+MCMC-BSS (10)	& 0. & 0. &  &  \\
+MCMC-BSS (30)	& 0. & 0. &  &  \\
+MCMC-BSS (100)& 0. & 0. &  &  \\
+\hline
+\end{tabular}
+\caption{Smoother performance measures for parameter case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
+\label{tab:case2_performance}
+\end{table}
+
+\begin{figure}%
+\includegraphics[width=\columnwidth]{case1_smoother_comparison_posRMSE_time-crop.pdf}%
+\caption{RMSEs and running times of various smoothing algorithms}%
+\label{fig:case1_rmse_vs_time}%
+\end{figure}
+
+\begin{figure}%
+%\includegraphics[width=\columnwidth]{case2_smoother_comparison_posRMSE_time-crop.pdf}%
+\caption{RMSEs and running times of various smoothing algorithms}%
+\label{fig:case2_rmse_vs_time}%
+\end{figure}
+
+The MCMC-BRS smoother error approaches that of the DBRS as the number of MCMC steps, $M$, increases.
+However, for small $M$, the error is only very slightly larger than that of the DBRS for very low running times.
+For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed.
+
+The MCMC-BSS outperforms all the other smoothers, achieveing a lower RMSE even with $M=1$. As $M$ is increased further, the accuracy increases.
+
+\section{Conclusion} \label{sec:conclusions}
+Two new algorithms have been presented for drawing particles from the joint smoothing distribution using MCMC.
+The MCMC schemes have a degree of flexibility in the choice of chain length, which controls a trade-off between accuracy and running time.
+
+The first algorithm resamples particles from the filtering distributions to produce particles from the smoothing distribution.
+The error performance of this scheme approaches that of the standard resampling scheme of \cite{Godsill2004} as the chain length increases.
+However, comparable error performance may be achieved with very short chains with a significant saving in processing time.
+
+The second algorithm samples new values of the state from a proposal distribution.
+Although slower than the resampling version for a fixed length of Markov chain, the error performance is reduced relative to the other smoothers.
 
@@
 % stfloats package.
-
-\section{Conclusion} \label{sec:conclusions}
-The conclusion goes here.
 

### 2012-01-25 MATLAB/backard_sampling_smoother.m

@@ -48,7 +48,7 @@ for ii = 1:S
         
         % Sample
 %         fprintf(1, '     ESS in frame %u: %f\n', kk, ESS(wts));
-        ind = randsample(length(wts), 1, true, exp(wts));
+        ind = sample_weights(exp(wts));
         
         % Update trajectory
         smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];

### 2012-01-25 MATLAB/batch_analysis.m

@@ -34,6 +34,18 @@ bar( cellfun(@(x) x.mean_pos_rmse, results) );
 figure(4), hold on
 bar( cellfun(@(x) x.mean_vel_rmse, results) );
 
+figure(5), hold on
+bar( cellfun(@(x) x.times, results) );
+
+figure(6), hold on
+inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*k' )
+inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*r' )
+inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-b' )
+inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-g' )
+legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
+xlabel('Running Time'); ylabel('Position RMSE');
+% saveas(6, 'smoother_comparison_posRMSE_time.pdf', 'pdf');
+
 % fprintf(1, '\n');
 % fprintf(1, 'Mean num of unique pts: |  KitiSm  |  DireSm  |  MCMCS1  |  MCMCS3  |  MCMCS10 |  MCMCS30 | MCMCS100 |\n');
 % fprintf(1, '                           %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |  %5.2f   |\n', mean(results.kiti.unique_pts), mean(results.bs.unique_pts), mean(results.mcmc1.unique_pts), mean(results.mcmc3.unique_pts), mean(results.mcmc10.unique_pts), mean(results.mcmc30.unique_pts), mean(results.mcmc100.unique_pts));

### 2012-01-25 MATLAB/batch_smoother_testing.m

@@ -3,7 +3,7 @@ dbstop if error
 
 % Get test flag
 % test_flag = str2double(getenv('SGE_TASK_ID'));
-test_flag = 2;
+test_flag = 3;
 
 % Batch test parameters
 num_seed = 10;
@@ -16,6 +16,9 @@ if test_flag == 1
 elseif test_flag == 2
     params.bng_var = (pi/180)^2;
     params.rng_var = 10;
+elseif test_flag == 3
+    params.bng_var = (pi/36)^2;
+    params.rng_var = 100;
 end
 
 % Number of smoothing algorithms
@@ -35,6 +38,8 @@ end
 
 % Loop through random seend
 for rs = 1:num_seed;
+    
+    rs
     
     % Set random seed
     s = RandStream('mt19937ar', 'seed', rs);

### 2012-01-25 MATLAB/mcmc_newstate_smoother.m

@@ -48,7 +48,7 @@ for ii = 1:S
             
             % Propose a new k-1 particle
 %             ind = randsample(Np, 1);
-            ind = randsample(Np, 1, true, exp(wts_array{kk-1}));
+            ind = sample_weights(exp(wts_array{kk-1}));
             
             % Propose a new k state
             [new_x, new_ppsl_prb] = feval(h_bidirec_ppsl, pts_array{kk-1}{ind}(:,kk-1), smooth_pts{ii}(:,kk+1), observs(:,kk), dt_prev, dt_next, h_trans, h_obs);

### 2012-01-25 MATLAB/mcmc_smoother.m

@@ -41,7 +41,7 @@ for ii = 1:S
             
             % Propose a new particle
 %             ind = randsample(Np, 1);
-            ind = randsample(Np, 1, true, exp(wts_array{kk}));
+            ind = sample_weights(exp(wts_array{kk}));
             
             % Calculate probability
             [~, new_trans_prb] = feval(h_trans, dt, pts_array{kk}{ind}(:,kk), smooth_pts{ii}(:,kk+1) );

### 2012-01-25 MATLAB/run_mcmcsmoother_demo.m

@@ -17,9 +17,11 @@ set_parameters;
 % Plot it
 xmax = max(max(abs(x(1:2,:))))+5;
 figure(1); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
-plot(x(1,:), x(2,:), 'b')
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3);
 [x_obs, y_obs] = pol2cart(y(1,:), y(2,:)); plot(x_obs, y_obs, 'r');
-figure(2); plot(t, y);
+plot(0, 0, 'xk', 'markersize', 10)
+legend('Target Position', 'Observations')
+% figure(2); plot(t, y);
 
 %% Run a PF
 init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);

### 2012-01-25 MATLAB/set_parameters.m

@@ -14,4 +14,4 @@ params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
 % Algorithm
 params.Np = 100;
 params.S = 100;
-params.M = 5;                   % Try 5;           50;
+params.M = 10;                   % Try 5;           50;

### 2012-01-25 MATLAB/tracking_bidirec_ppsl.m

@@ -20,7 +20,7 @@ H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
 if (nargin < 8)||isempty(x)
     x = mvnrnd(m', P)';
 end
-ppsl_prb = log(mvnpdf(x', m', P));
+ppsl_prb = fast_log_mvnpdf(x', m', P);
 
 end
 

### 2012-01-25 MATLAB/tracking_obs.m

@@ -9,7 +9,7 @@ R = [params.bng_var 0; 0 params.rng_var];
 y_mn = [b; r];
 
 if ~isempty(y)
-    obs_prb = log(mvnpdf(y', y_mn', R));
+    obs_prb = fast_log_mvnpdf(y', y_mn', R);
 end
 
 end

### 2012-01-25 MATLAB/tracking_ppsl.m

@@ -12,7 +12,7 @@ H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
      m(1)/sqrt(m(1)^2+m(2)^2), m(2)/sqrt(m(1)^2+m(2)^2), 0, 0];
 [m, P] = ekf_update1(m, P, y, H, R, h_obs);
 x = mvnrnd(m', P)';
-ppsl_prb = log(mvnpdf(x', m', P));
+ppsl_prb = fast_log_mvnpdf(x', m', P);
 
 % % Bootstrap
 % x = mvnrnd((A*last_x), Q)';

### 2012-01-25 MATLAB/tracking_trans.m

@@ -8,7 +8,7 @@ global params;
 x_mn = A*last_x;
 
 if ~isempty(x)
-    trans_prb = log(mvnpdf(x', x_mn', Q));
+    trans_prb = fast_log_mvnpdf(x', x_mn', Q);
 end
 
 end

### 2012-02-15 Paper/draft4_MCMCsmoother.tex
Rewriting draft.

--- 2012-01-25 Paper/draft3_MCMCsmoother.tex
@@
 \usepackage{eqparbox}
-\usepackage[font=footnotesize]{subfig}
+\usepackage[caption=false,font=footnotesize]{subfig}
 \usepackage{fixltx2e}
-\usepackage{stfloats}
+%\usepackage{stfloats}
 \usepackage{url}
@@
 \markboth{IEEE Transaction in Signal Processing,~Vol.~1, No.~1, January~1901}%
-{Shell \MakeLowercase{\textit{et al.}}: Particle smoothing using Markov chain Monte Carlo sampling for backwards simulation}
+{Bunch \& Godsill \MakeLowercase{\textit{et al.}}: Improved Particle Approximations to the Joint Smoothing Distribution Using Markov Chain Monte Carlo}
 
@@
 p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) \nonumber \\
-                                    & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k}|y_{1:k-1})
+                                    & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k-1}|y_{1:k-1})
 \end{IEEEeqnarray}
@@
 \underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \underbrace{\begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix}}_{\mathbf{A}} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
-\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k} \\
+\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}
 \end{IEEEeqnarray}
@@
 
-\begin{figure}
+\begin{figure}[!t]
 \centerline{\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory-crop.pdf}%
@@
 
-\begin{table}%
+\begin{table}[!t]%
+\renewcommand{\arraystretch}{1.3}
+\caption{Smoother performance measures for parameter case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
+\label{tab:case1_performance}
+\centering
 \begin{tabular}{|c||c|c|c|c|}
@@
 \end{tabular}
-\caption{Smoother performance measures for parameter case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
-\label{tab:case1_performance}
 \end{table}
 
-\begin{table}%
+\begin{table}[!t]%
+\renewcommand{\arraystretch}{1.3}
+\caption{Smoother performance measures for parameter case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
+\label{tab:case2_performance}
+\centering
 \begin{tabular}{|c||c|c|c|c|}
@@
 \hline
-FS 						& 0. & 0. & 0 &  \\
-DBRS					& 0. & 0. &  &  \\
+FS 						& 6.17 & 2.03 & 0 & 3.96 \\
+DBRS					& 5.79 & 1.89 & 63 & 7.34 \\
 \hline
-MCMC-BRS (1)	& 0. & 0. &  &  \\
-MCMC-BRS (3)	& 0. & 0. &  &  \\
-MCMC-BRS (10)	& 0. & 0. &  &  \\
-MCMC-BRS (30)	& 0. & 0. &  &  \\
-MCMC-BRS (100)& 0. & 0. &  &  \\
+MCMC-BRS (1)	& 6.13 & 2.01 & 2 & 4.77 \\
+MCMC-BRS (3)	& 6.07 & 1.99 & 4 & 5.29 \\
+MCMC-BRS (10)	& 5.94 & 1.94 & 10 & 6.1 \\
+MCMC-BRS (30)	& 5.82 & 1.91 & 29 & 6.73 \\
+MCMC-BRS (100)& 5.80 & 1.89 & 95 & 7.38 \\
 \hline
-MCMC-BSS (1)	& 0. & 0. &  &  \\
-MCMC-BSS (3)	& 0. & 0. &  &  \\
-MCMC-BSS (10)	& 0. & 0. &  &  \\
-MCMC-BSS (30)	& 0. & 0. &  &  \\
-MCMC-BSS (100)& 0. & 0. &  &  \\
+MCMC-BSS (1)	& 5.90 & 1.91 & 9 & 12.6 \\
+MCMC-BSS (3)	& 5.67 & 1.81 & 18 & 22.7 \\
+MCMC-BSS (10)	& 5.48 & 1.73 & 50 & 44.1 \\
+MCMC-BSS (30)	& 5.43 & 1.69 & 144 & 70.9 \\
+MCMC-BSS (100)& 5.34 & 1.68 & 471 & 91.8 \\
 \hline
 \end{tabular}
-\caption{Smoother performance measures for parameter case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
-\label{tab:case2_performance}
 \end{table}
 
-\begin{figure}%
-\includegraphics[width=\columnwidth]{case1_smoother_comparison_posRMSE_time-crop.pdf}%
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.8\columnwidth]{case1_smoother_comparison_posRMSE_time-crop.pdf}%
 \caption{RMSEs and running times of various smoothing algorithms}%
@@
 
-\begin{figure}%
-%\includegraphics[width=\columnwidth]{case2_smoother_comparison_posRMSE_time-crop.pdf}%
+\begin{figure}[!t]
+\centering
+\includegraphics[width=0.8\columnwidth]{case2_smoother_comparison_posRMSE_time-crop.pdf}%
 \caption{RMSEs and running times of various smoothing algorithms}%
@@
 
-\appendices
-\section{Appendix Title}
-Appendix one text goes here.
+%\appendices
+%\section{Appendix Title}
+%Appendix one text goes here.
 
@@
 % if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
+%\section{}
+%Appendix two text goes here.
 
 % use section* for acknowledgement
-\section*{Acknowledgment}
+%\section*{Acknowledgment}
 
-The authors would like to thank...
+%The authors would like to thank...
 
@@
 
-\begin{IEEEbiography}{Pete Bunch}
+\begin{IEEEbiographynophoto}{Pete Bunch}
 Biography text here.
-\end{IEEEbiography}
+\end{IEEEbiographynophoto}
 

### 2012-02-15 Paper/draft5_MCMCsmoother.tex
Rewriting draft.

--- 2012-02-15 Paper/draft4_MCMCsmoother.tex
@@
 \IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quanitity from incomplete or inaccurate observations.
-This is commonly achieved through the use of probabilistic models for the evolution of the latent state and the measurement process.
+This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities.
 
@@
 
-Here, the random variable $x_k$ denotes the value of the latent state at the $k^{th}$ instant, and $y_k$ the value of the observation.
+Here, the random variable $x_k$ denotes the value of the latent state at the $k$th instant, and $y_k$ the value of the observation.
 A set of random variables will be written as $x_{1:k} = \{x_1, x_2, \dots, x_k \}$.
@@
 
-In the case where the state and observation processes of equations~\ref{eq:state_proc}~and~\ref{eq:observ_proc} are linear and Gaussian, the filtering problem may be solved analytically using the Kalman filter \cite{Kalman1960}.
+In the case where the state and observation processes of (\ref{eq:state_proc}) and~(\ref{eq:observ_proc}) are linear and Gaussian, the filtering problem may be solved analytically using the Kalman filter \cite{Kalman1960}.
 For nonlinear or non-Gaussian models, tractable closed-form solutions are rare.
-In such cases, numerical approximations are often employed, including the particle filter algorithm, first introduced by \cite{Gordon1993}.
+Instead, numerical approximations are often employed, including the particle filter algorithm, first introduced by \cite{Gordon1993}.
 (See \cite{Cappe2007,Doucet2009} for a thorough introduction.)
@@
 
-Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$. $\hat{p}$ means an approximation to $p$.
+Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$.% $\hat{p}$ means an approximation to $p$.
 
@@
 In the linear Gaussian case, the entire state sequence, $x_{1:K}$, may be estimated in closed-form using the a Rauch-Tung-Striebel (RTS) smoother \cite{Rauch1965}.
-This works by modifying the ordinary Kalman filter results in a backward processing pass through the observations.
+This works by modifying the ordinary Kalman filter results in a backward processing pass through the state sequence.
 For nonlinear or non-Gaussian models, it is possible again to resort to numerical methods.
-Now, the particles of the approximation are drawn from the joint smoothing distribution.
+Now, particles are used to approximate the joint smoothing distribution.
 
@@
 Methods for generating these samples have been described in \cite{Kitagawa1996,Godsill2004,Briers2010}.
-In this paper new methods are presented for drawing samples from the joint smoothing distribution, using Markov chain Monte Carlo (MCMC).
+In this paper new algorithms are presented for drawing samples from the joint smoothing distribution, using Markov chain Monte Carlo (MCMC).
 These novel procedures have computational advantages and greater flexibility over previous methods.
@@
 The particle filter is a recursive numerical algorithm for estimation of the filtering distribution.
-At the $k^{th}$ processing step, a particle estimate of the joint distribution over $x_{k-1}$ and $x_k$ is made by drawing samples from a proposal distribution.
+At the $k$th processing step, a particle estimate of the joint density over $x_{k-1}$ and $x_k$ is made using importance sampling (IS).
+A particle is first sampled from the factorised proposal density,
 
 \begin{IEEEeqnarray}{rCl}
-\{ x_{k-1}^{(i)}, x_k^{(i)} \} & \sim & q(x_{x}|x_{k-1}) q(x_{k-1})
+\{ x_{k-1}^{(i)}, x_k^{(i)} \} & \sim & q(x_{x}|x_{k-1}) q(x_{k-1}).
 \end{IEEEeqnarray}
 
-The $k-1$ state proposal distribution, $q(x_{k-1})$, is constructed from the particles of the filtering distribution from the previous time step.
-The choice of weights and sampling procedure here determines what from of ``resampling'' is to be used.
+The proposal density for $x_k$, $q(x_{x}|x_{k-1})$, is often called the ``importance density''.
+Common choices include the model transition density (\ref{eq:state_proc}), and the ``optimal importance distribution'' \cite{Doucet2000a}.
+The proposal density for $x_{k-1}$, $q(x_{k-1})$, is constructed from the particles of the filtering approximation from the previous time step.
+The choice of weights and sampling procedure for $x_{k-1}$ determines what from of ``resampling'' is to be used.
 Here we use arbitrary weights for generality.
@@
 
-Note that the choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999}.
-For the least computational expense, the proposal distribution with weights $v_{k-1}^{(i)} = 1/N_F$ (where $N_F$ is the number of filter particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling.
-(See \cite{Cappe2007,Doucet2009} for further discussion of resampling.)
+the choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999}.
+For the least computational expense, the proposal density with weights $v_{k-1}^{(i)} = 1/N_F$ (where $N_F$ is the number of filter particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling.
+(See \cite{Douc2005} for further discussion of resampling.)
 
-Next, the particles are assigned an importance weight according to the ratio of the targeted joint distribution and the proposal.
+Next, the particles are assigned an importance weight according to the ratio of the target and the proposal densities.
 
@@
 
-Normalisation is enforced by scaling the weights so that they sum to 0. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the values from each particles.
+Normalisation is enforced by scaling the weights so that they sum to 0. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the values from each particle.
 
@@
 
-The weakness of this simple, filter-smoother (FS) is particle degeneracy.
+The weakness of this simple, filter-smoother (FS) is degeneracy of the particles in the path-space.
 Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, while those with high weights appear multiple times.
-The result is that many, or even all, particles will have the same values at early time steps.
-This effect is illustrated in figure~\ref{} in section~\ref{}.
-If the objective is only to produce a single point estimate then this might be sufficient, but to characterise the complete smoothing distribution the approximation must be rejeuvenated.
+The result is that many, or even all, particles will share a common ancestor, before which they all have the same set of state values.
+If the objective is only to produce a single point estimate then this might be sufficient, but to characterise the complete smoothing distribution, the path-space diversity of the approximation must be improved.
 
@@
 Particles are generated by sampling backwards in time from the factors of this expansion.
-At the $k^{th}$ step, a sample $\tilde{x}_{k+1:K} \sim p(x_{k+1:K}|y_{1:K})$ will already have been drawn.
+At the $k$th step, a sample $\tilde{x}_{k+1:K} \sim p(x_{k+1:K}|y_{1:K})$ will already have been drawn.
 By sampling $x_k \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$ and appending $x_k$ to $\tilde{x}_{k+1:K}$, the state sequence is sequentially extended back in time.
 The recursion is initialised with $x_K \sim p(x_K|y_{1:K})$, i.e.
-a sample from the final filtering distribution.
+a sample from the final filtering approximation.
 
-The backwards conditional distributions may be expanded with Bayes rule.
+The backwards conditional density may be expanded with Bayes rule.
 
@@
 The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K})$, by combination of the results from one forward and one backward particle filter.
-Particles from the joint distribution may then be generated by resampling from either the forward or backward filters.% in the manner of algorithm~\ref{alg:DBRS}.
+Particles from the joint distribution may then be generated by resampling from either the forward or backward filters (or some combination) in the manner of the backward-resampling smoother outlined above.%algorithm~\ref{alg:DBRS}.
 
@@
 This limits the support of the smoother to the states that appear in the filtering particles.
-In \cite{Fearnhead2010}, an algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal, allowing full support.
-In addition, this algorithm achieves $\mathcal{O}( N_S \times K)$ complexity, linear in the number of particles!
+In \cite{Fearnhead2010}, an IS-based algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal, allowing full support.
+In addition, this algorithm achieves $\mathcal{O}( N_S \times K)$ complexity, linear in the number of particles.
 
@@
 
-The bottleneck of the backward-resampling smoother is the calculation of the sampling weights (equation~\ref{eq:DBRS_weights}).
+The bottleneck of the backward-resampling smoother is the calculation of the sampling weights, specified by (\ref{eq:DBRS_weights}).
 For each smoother particle and for each time step, a weight must be calculated for each of the $N_F$ filter particles, in order to draw a sample from the conditional distribution $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
@@
 A thorough introduction can be found in \cite{Gilks1996}.
-The basic Metropolis-Hastings (MH) \cite{Metropolis1953} sampler uses a series of steps in which a new state is drawn from a proposal distribution and accepted with a probability determined by the ratios of the target and proposal probabilities.
+The basic Metropolis-Hastings (MH) \cite{Hastings1970} sampler uses a series of steps in which a new state is drawn from a proposal distribution and accepted with a probability determined by the ratios of the target and proposal densities.
 
-For the direct backward sampling smoother, the objective at each time step was to draw a sample from $p(x_{k}|\tilde{x}_{k+1}, y_{1:K})$.
+For the direct backward-resampling smoother, the objective at each time step was to draw a sample from $p(x_{k}|\tilde{x}_{k+1}, y_{1:K})$.
 To simplify the notation and initialisation of chains, the MCMC version will target $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
@@
 
-As before, the algorithm begins by first running a particle filter-smoother and then resampling states backwards through time to produce particles from the joint smoothing distribution, using the factorisation of equation~\ref{eq:smoothing_factorisation}.
-At the $k^{th}$ time step, a Markov chain is used to sample from $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
-The chain is initialised with the first $k$ states from the previous processing step.
+As before, the algorithm begins by first running a particle filter-smoother and then resampling states backwards through time to produce particles from the joint smoothing distribution, using the factorisation of (\ref{eq:smoothing_factorisation}).
+At the $k$th time step, a Markov chain is used to sample from $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
+The chain is initialised with the output from the previous processing step.
 As this is itself a sample from the target distribution, no burn-in period is required.
 
-Following an equivalent derivation to equation~\ref{eq:backward_conditional_filter}, the target distribution may be approximated by a particle distribution.
+Following an equivalent derivation to that for (\ref{eq:backward_conditional_filter}), the target distribution may be approximated by a particle distribution.
 
@@
 
-A valid proposal distribution may be constructed using the particles of the filter-smoother approximation with any arbitrary weights.
+A valid proposal may be constructed using the particles of the filter-smoother approximation with any arbitrary weights.
 
@@
 \begin{IEEEeqnarray}{rCl}
-\alpha &=& \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \nonumber \\
- &=& \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }  .
+\alpha &=& \min \Bigg[ 1, \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \Bigg] \nonumber \\
+       &=& \min \Bigg[1, \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }  \Bigg]  . \label{eq:MCMC-BRS_ap}
 \end{IEEEeqnarray}
@@
 				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
-				\STATE With probability $\alpha = \frac{ w_k^{*} \tilde{v}_k^{(m)} }{  w_k^{(m)} \tilde{v}_k^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) }$, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$.
-Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
+				\STATE With probability $\alpha$ specified by (\ref{eq:MCMC-BRS_ap}),\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
 			\ENDFOR
-			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{k}^{(i)(M)}$
+			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
 		\ENDFOR
@@
 
-For a practical implementation, it is not necessary to store the complete particles of the filter-smoother approximations, $\hat{p}(x_{1:k}|y_{1:k})$, which would require memory space which scaled quadratically as $K$ increased.
+For a practical implementation, it is not necessary to store the complete particles of the filter-smoother approximations, $\hat{p}(x_{1:k}|y_{1:k})$, which would require memory space which scaled quadratically with $K$.
 As the smoother progresses backwards in time, the earlier states in the sequence are repeatedly overwritten, and so it is not necessary ever to have known them.
@@
 With the MCMC variant, we also have the freedom to set $M$, the number of MH steps in each Markov chain.
-This controls a trade-off between speed and particle diversity.
+This controls a trade-off between speed and accuracy/particle diversity.
 If the chains are short, and the acceptance probabilities low, then this scheme will simply reproduce particles from the filter-smoother.
-However, when the weights of the conditional distribution approximation (equation~\ref{eq:MCMC-BRS_weights}) are similar, then an independent sample may be produced with only a very short chain.
+However, when the weights of the conditional distribution approximation (\ref{eq:MCMC-BRS_weights}) are similar, then an independent sample may be produced with only a very short chain.
 
-\section{Improving the smoother support} \label{sec:new_state_smoother}
+\section{An MCMC particle smoother with improved support} \label{sec:new_state_smoother}
 
@@
 This yielded improvements in accuracy and particle diversity over other marginal smoothing techniques, and also the advantage of $\mathcal{O}(N_S \times K)$ complexity.
-Here we adapt the method of \cite{Fearnhead2010} for generating particles from the joint smoothing distribution.
+Here we adapt this method for generating particles from the joint smoothing distribution.
 Again, an MCMC procedure is employed.
 
-The method described in \cite{Fearnhead2010} uses a forward and a backward particle filter.
+%The method described in \cite{Fearnhead2010} uses a forward and a backward particle filter.
 To develop a marginal smoothing approximation at time step $k$, particles are resampled from the forward filter at step $k-1$, and from the backward filter and step $k+1$.
@@
 It is undesirable to use importance sampling for generating particles from the joint sampling distribution, as the required resampling is liable to lead to a loss of particle diversity similar to that suffered by the filter-smoother.
+
+The smoother described in \cite{Fearnhead2010} uses importance sampling to approximate the marginal smoothing distributions.
+This method could be extended to estimation of the joint smoothing distribution by applying it sequentially.
+However, the resampling required to limit weight degeneracy would lead to a loss of path-space diversity similar to that suffered by the filter-smoother.
 

### 2012-02-15 MATLAB/batch_analysis.m

@@ -3,7 +3,7 @@
 close all
 
 % Set test flag
-test_flag = 1;
+test_flag = 3;
 
 % Load data
 load(['smoother_test' num2str(test_flag) '.mat']);
@@ -11,7 +11,7 @@ load(['smoother_test' num2str(test_flag) '.mat']);
 % 
 NA = 13;
 t = params.dt:params.dt:params.dt*params.N;
-col = {[1 0 0], [.5 0 0], [0 1 0], [0 .5 0], [0 0 1], [0 0 .5]};
+col = {[1 0 0], [1 .5 .5], [0 1 0], [.5 1 .5], [0 0 1], [.5 .5 1]};
 
 %% Output results
 

### 2012-02-16 Paper/draft5_MCMCsmoother.tex
Final draft for Simon to proofread.

@@
 \usepackage[cmex10]{amsmath}
-\usepackage{algorithm}
+%\usepackage{algorithm}
 \usepackage{algorithmic}
@@
 
-The particle filter is summarised below.%in algorithm~\ref{alg:PF}.
+The particle filter is summarised in Fig.~\ref{alg:PF}.
 
 %\begin{algorithm}
-\begin{algorithmic}
+\begin{figure}
+\fbox{\parbox{\columnwidth}{
+\begin{algorithmic}[1]
 	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
-	\FOR{$k = 1 \dots K$}
- 		\FOR{$i = 1 \dots N_F$}
+	\FOR{$k = 1$ \TO $K$}
+ 		\FOR{$i = 1$ \TO $N_F$}
  			\STATE Sample $\{ x_{k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
@@
 \end{algorithmic}
-%\caption{Particle filter algorithm}
-%\label{alg:PF}
+}}
+\caption{Particle filter algorithm}
+\label{alg:PF}
+\end{figure}
 %\end{algorithm}
@@
 
-The filter-smoother is summarised below.%in algorithm~\ref{alg:FS}.
+The filter-smoother is summarised in Fig.~\ref{alg:FS}.
 
 %\begin{algorithm}
-\begin{algorithmic}
+\begin{figure}
+\fbox{\parbox{\columnwidth}{
+\begin{algorithmic}[1]
  	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
- 	\FOR{$k = 1 \dots K$}
- 		\FOR{$i = 1 \dots N_F$}
+ 	\FOR{$k = 1$ \TO $K$}
+ 		\FOR{$i = 1$ \TO $N_F$}
  			\STATE Sample $\{ x_{1:k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
@@
 \end{algorithmic}
-%\caption{Filter-smoother algorithm}
-%\label{alg:FS}
+}}
+\caption{Filter-smoother algorithm}
+\label{alg:FS}
+\end{figure}
 %\end{algorithm}
@@
 
-The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised below.% in algorithm~\ref{alg:DBRS}.
+The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised in Fig.~\ref{alg:DBRS}.
 
 %\begin{algorithm}
-\begin{algorithmic}
+\begin{figure}
+\fbox{\parbox{\columnwidth}{
+\begin{algorithmic}[1]
  	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$.
-	\FOR{$i = 1 \dots N_S$}
+	\FOR{$i = 1$ \TO $N_S$}
 		\STATE Sample $\tilde{x}_{K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{K}^{(j)}}(x_{K})$.
-		\FOR{$k = K-1 \dots 1$}
-			\FOR{$j = 1 \dots N_F$}
+		\FOR{$k = K-1$ \TO $1$}
+			\FOR{$j = 1$ \TO $N_F$}
 				\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
@@
 \end{algorithmic}
-%\caption{Direct backward-resampling smoother algorithm}
-%\label{alg:DBRS}
+}}
+\caption{Direct backward-resampling smoother algorithm}
+\label{alg:DBRS}
+\end{figure}
 %\end{algorithm}
@@
 
-The complete MCMC backward-resampling smoother is summarised below.% in algorithm~\ref{alg:MCMC-BRS}.
+The complete MCMC backward-resampling smoother is summarised in Fig.~\ref{alg:MCMC-BRS}.
 
 %\begin{algorithm}
-\begin{algorithmic}
+\begin{figure}
+\fbox{\parbox{\columnwidth}{
+\begin{algorithmic}[1]
  	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
-	\FOR{$i = 1 \dots N_S$}
+	\FOR{$i = 1$ \TO $N_S$}
 		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
-		\FOR{$k = K-1 \dots 1$}
+		\FOR{$k = K-1$ \TO $1$}
 			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
-			\FOR{$m = 1 \dots M$}
+			\FOR{$m = 1$ \TO $M$}
 				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
@@
 \end{algorithmic}
-%\caption{MCMC backward-resampling smoother algorithm}
-%\label{alg:MCMC-BRS}
+}}
+\caption{MCMC backward-resampling smoother algorithm}
+\label{alg:MCMC-BRS}
+\end{figure}
 %\end{algorithm}
@@
 
+Smoothers based on backward-resampling, both the direct sampler of \cite{Godsill2004} and the MCMC version outlined in the previous section, are limited by the fact that states are only selected from those which appear in the filtering approximation.
+If there is a significant difference between the filtering and smoothing densities, then the the particles are unlikely to be in the right locations to represent the smoothing distribution well, and the resulting approximation will be poor.
+
 In \cite{Fearnhead2010}, the authors described an algorithm for estimating the marginal smoothing distributions.
@@
 The proposed new method proceeds in a similar manner to the MCMC backward-resampling algorithm.
-However, this time $x_{1:k-1}$, is resampled from the particle filter-smoother approximation from time step $k-1$.
+However, this time only $x_{1:k-1}$ is resampled from the particle filter-smoother approximation from time step $k-1$.
 The current state, $x_k$, is then sampled afresh from a new proposal distribution.
@@
 \begin{IEEEeqnarray}{rCl}
-q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & = & q(x_{1:k-1}|y_{1:k-1}) q(x_{k}|x_{k-1}, \tilde{x}_{k+1}, y_{k}) \nonumber \\
+\IEEEeqnarraymulticol{3}{l}{ q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = q(x_{1:k-1}|y_{1:k-1}) q(x_{k}|x_{k-1}, \tilde{x}_{k+1}, y_{k}) } \nonumber \\
                                     & = & \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1}) q(x_{k}|x_{k-1}^{(j)}, \tilde{x}_{k+1}, y_{k})  .
@@
 \begin{IEEEeqnarray}{rCl}
-p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) \nonumber \\
+\IEEEeqnarraymulticol{3}{l}{p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) } \nonumber \\
                                     & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k-1}|y_{1:k-1})
@@
 \begin{IEEEeqnarray}{rCl}
-\alpha &=& \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \nonumber \\
- &=& \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \nonumber \\
- \IEEEeqnarraymulticol{3}{l}{\frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } } .
+\alpha &=& \min \Bigg[ 1, \frac{ \hat{p}(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) }{ \hat{p}(x_{1:k}^{(m)}|\tilde{x}_{k+1}, y_{1:K}) q(x_{1:k}^{*}|\tilde{x}_{k+1}, y_{1:K}) } \Bigg] \nonumber \\
+ &=& \min \Bigg[ 1, \frac{ w_{k-1}^{*} p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) }{ w_{k-1}^{(m)} p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) } \nonumber \\
+ & & \times \frac{ \tilde{v}_{k-1}^{(m)} q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ \tilde{v}_{k-1}^{*} q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } \Bigg] \label{eq:MCMC-BSS_ap}
+% \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \nonumber \\
+% \IEEEeqnarraymulticol{3}{l}{\frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } } .
 \end{IEEEeqnarray}
 
-The complete MCMC backward-sampling smoother is summarised below.% in algorithm~\ref{alg:MCMC-BSS}.
+As noted in \cite{Fearnhead2010}, the optimal choice of proposal is the conditional posterior state distribution,
+
+\begin{IEEEeqnarray}{rCl}
+q(x_k|x_{k-1},x_{k+1},y_k) & = & p(x_k|x_{k-1},x_{k+1},y_k)  .
+\end{IEEEeqnarray}
+
+The complete MCMC backward-sampling smoother is summarised in Fig.~\ref{alg:MCMC-BSS}.
 
 %\begin{algorithm}
- \begin{algorithmic}
+ \begin{figure}
+\fbox{\parbox{\columnwidth}{
+\begin{algorithmic}[1]
  	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
-	\FOR{$i = 1 \dots N_S$}
+	\FOR{$i = 1$ \TO $N_S$}
 		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
-		\FOR{$k = K-1 \dots 1$}
+		\FOR{$k = K-1$ \TO $1$}
 			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
-			\FOR{$m = 1 \dots M$}
+			\FOR{$m = 1$ \TO $M$}
 				\STATE Propose a new state hisory, $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
 				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
-				\STATE With probability $\alpha = \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }$, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$.
-Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
+				\STATE With probability $\alpha$ specified by (\ref{eq:MCMC-BSS_ap}),\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
 			\ENDFOR
-			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{k}^{(i)(M)}$
+			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
 		\ENDFOR
@@
 \end{algorithmic}
-%\caption{MCMC backward-sampling smoother algorithm}
-%\label{alg:MCMC-BSS}
+}}
+\caption{MCMC backward-sampling smoother algorithm}
+\label{alg:MCMC-BSS}
+\end{figure}
 %\end{algorithm}
@@
 The performance of the new smoothers was tested on a simple 2D tracking model.
-The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model.
+The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model (see \cite{Bar-Shalom2002} for details).
+
+%\begin{IEEEeqnarray}{rCl}
+%\underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \underbrace{\begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix}}_{\mathbf{A}} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
+%\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}
+%\end{IEEEeqnarray}
 
 \begin{IEEEeqnarray}{rCl}
-\underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \underbrace{\begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix}}_{\mathbf{A}} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
+\underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
 \underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}
@@
 $I_2$ and $0_2$ denote the $2 \times 2$ identity and zero matrices.
-The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ have a zero-mean Gaussian distribution with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively.
+The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ are independent and are zero-mean Gaussian distributed with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively.
 
@@
 $v_k^{(j)} = w_k^{(j)}$.
-The particle filter and MCMC-BSS both use ``optimal importance distributions'' (see e.g.
-\cite{Cappe2007}) with linearisations of the observation model where appropriate.
+The particle filter and MCMC-BSS both use ``optimal importance distributions'' with linearisations of the observation model where appropriate.
 
@@
 Other model settings are $\Delta t = 1$, $\sigma_P = 1$, and $x_0 = [-100, 50, 10, 0]^T$.
-Example tracjectories and observations from the two cases are shown in figure~\ref{fig:example_trajectories}.
+Example trajectories and observations from the two cases are shown in figure~\ref{fig:example_trajectories}.
 
@@
 
-Tables~\ref{tab:case1_performance} and~\ref{tab:case1_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms with the two parameter cases, along with running times (excluding the time for filtering) and the mean number of unique particles at each time instant.
+Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms with the two parameter cases, along with running times (excluding the time for filtering) and the mean number of unique particles at each time instant.
 The trade-off between speed and error is shown in figures~\ref{fig:case1_rmse_vs_time} and~\ref{fig:case2_rmse_vs_time}.
@@
 The MCMC-BRS smoother error approaches that of the DBRS as the number of MCMC steps, $M$, increases.
-However, for small $M$, the error is only very slightly larger than that of the DBRS for very low running times.
+However, for small $M$ the error is only very slightly larger than that of the DBRS but with significantly less running time.
 For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed.
@@
 Two new algorithms have been presented for drawing particles from the joint smoothing distribution using MCMC.
-The MCMC schemes have a degree of flexibility in the choice of chain length, which controls a trade-off between accuracy and running time.
+These MCMC schemes have a degree of flexibility in the choice of chain length, which controls a trade-off between accuracy and running time.
 
@@
 
-The second algorithm samples new values of the state from a proposal distribution.
-Although slower than the resampling version for a fixed length of Markov chain, the error performance is reduced relative to the other smoothers.
+The second algorithm samples new values of the state from a continuous proposal density.
+Although slower than the resampling version for a given length of Markov chain, the error performance is reduced relative to the other smoothers.
 

### 2012-02-22 Paper/complete/finaldraft_MCMCsmoother.tex
final draft with Simon's changes, nice images, nice references

--- 2012-02-16 Paper/draft5_MCMCsmoother.tex
@@
 \usepackage{cite}
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-  \graphicspath{{../pdf/}{../jpeg/}}
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  \usepackage[dvips]{graphicx}
-  \graphicspath{{../eps/}}
-  \DeclareGraphicsExtensions{.eps}
-\fi
+\usepackage[pdftex]{graphicx} \graphicspath{{images/}}
 \usepackage[cmex10]{amsmath}
@@
 
-\author{Pete~Bunch,~\IEEEmembership{}
-        Simon~Godsill,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space
+\author{Pete~Bunch,~\IEEEmembership{Member,~IEEE,} and
+        Simon~Godsill,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
 \thanks{P. Bunch and S. Godsill are with the Department
@@
 
-\IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quanitity from incomplete or inaccurate observations.
-This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities.
-
+\IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quantity from incomplete or inaccurate observations.
+This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities,
+%
 \begin{IEEEeqnarray}{rCl}
 x_{k} & \sim & p(x_{k}|x_{k-1}) \label{eq:state_proc}\\
-y_{k} & \sim & p(y_{k}|x_{k})   \label{eq:observ_proc}
+y_{k} & \sim & p(y_{k}|x_{k})   \label{eq:observ_proc}   .
 \end{IEEEeqnarray}
@@
 Filtering is the inference of $p(x_k|y_{1:k})$, the distribution of the current state given all previous observations.
-Smoothing consists of two related tasks, the inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps), the joint distribution of the entire state sequence given all the observations, and that of $p(x_{k}|y_{1:K})$, the marginal distribution of each state given all the observations.
+Smoothing comprises two related tasks, the inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps), the joint distribution of the entire state sequence given all the observations, and that of $p(x_{k}|y_{1:K})$, the marginal distribution of each state given all the observations.
 In this paper, we are concerned with the estimation of the joint smoothing distribution.
@@
 In a particle filter, the filtering distribution is approximated by a set of weighted samples (or ``particles'') drawn from that distribution.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$.% $\hat{p}$ means an approximation to $p$.
+Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$, and $w_k{(i)}$ is the weight of the $i$th particle.
+Weights are non-negative and sum to 1.% $\hat{p}$ means an approximation to $p$.
 
-A similar story applies to the problem of smoothing.
+Similar ideas apply to the problem of smoothing.
 In the linear Gaussian case, the entire state sequence, $x_{1:K}$, may be estimated in closed-form using the a Rauch-Tung-Striebel (RTS) smoother \cite{Rauch1965}.
@@
 Now, particles are used to approximate the joint smoothing distribution.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 A particle is first sampled from the factorised proposal density,
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 Here we use arbitrary weights for generality.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-the choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999}.
+The choice $v_{k-1}^{(i)} = w_{k-1}^{(i)}$ corresponds to ordinary resampling, and other choices of $v_{k-1}^{(i)}$ represent auxiliary sampling schemes \cite{Pitt1999,Cappe2007}.
 For the least computational expense, the proposal density with weights $v_{k-1}^{(i)} = 1/N_F$ (where $N_F$ is the number of filter particles) can be ``sampled'' by simply keeping the set of particles generated in the last step, with no resampling.
-(See \cite{Douc2005} for further discussion of resampling.)
+This, however, leads to degeneracy of the particle weights over time.
+(See \cite{Douc2005,Cappe2007,Doucet2009} for further discussion of resampling.)
 
 Next, the particles are assigned an importance weight according to the ratio of the target and the proposal densities.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-Normalisation is enforced by scaling the weights so that they sum to 0. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the values from each particle.
+Normalisation is enforced by scaling the weights so that they sum to 1. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the $x_{k-1}^{(i)}$ values from each particle.
 
@@
 \begin{algorithmic}[1]
+\linespread{1.7} \selectfont
 	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
@@
  		\FOR{$i = 1$ \TO $N_F$}
- 			\STATE Sample $\{ x_{k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
+ 		  \STATE Resample last state $x_{k-1}^{(i)} \sim \sum_j v_{k-1}^{(j)} \delta_{x_{k-1}^{(j)}}(x_{k-1})$.
+ 			\STATE Sample new state $x_k^{(i)} \sim q(x_k|x_{k-1}^{(i)})$.
  			\STATE Weight $w_{k}^{(i)} \propto \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }$.
@@
 
-\subsection{Existing Algorithms for Particle Smoothing }
+\subsection{Existing Algorithms for Particle Smoothing } \label{sec:existing_smoothers}
 
@@
 
-The most basic particle smoother was presented in \cite{Kitagawa1996}, and is a trivial extension of the particle filter.
+The most basic particle smoother was presented in \cite{Kitagawa1996}, and is a straightforward extension of the particle filter.
 Rather than marginalising the previous states from each particle, these past values are stored as well.
@@
 
-The weakness of this simple, filter-smoother (FS) is degeneracy of the particles in the path-space.
-Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, while those with high weights appear multiple times.
+The weakness of this simple, filter-smoother is degeneracy of the particles in the path-space.
+Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, whereas those with high weights appear multiple times.
 The result is that many, or even all, particles will share a common ancestor, before which they all have the same set of state values.
@@
 \begin{algorithmic}[1]
+\linespread{1.7} \selectfont
  	\STATE Initialise particles from prior, $x_{0}^{(i)} \sim p(x_{0})$.
@@
  		\FOR{$i = 1$ \TO $N_F$}
- 			\STATE Sample $\{ x_{1:k-1}^{(i)}, x_k^{(i)} \} \sim \sum_j v_{k-1}^{(j)} q(x_k|x_{k-1}^{(j)})$.
+ 			\STATE Resample history $x_{1:k-1}^{(i)} \sim \sum_j v_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
+ 			\STATE Sample new state $x_k^{(i)} \sim q(x_k|x_{k-1}^{(i)})$.
  			\STATE Weight $w_{k}^{(i)} \propto \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }$.
@@
 The sampling procedure exploits the following factorisation of the joint distribution.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-The backwards conditional density may be expanded with Bayes rule.
-
+The backwards conditional density may be expanded with Bayes' rule,
+%
 \begin{IEEEeqnarray}{rCl}
 p(x_k|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_k|\tilde{x}_{k+1}, y_{1:k}) \nonumber \\
-                                & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|y_{1:k})
+                                & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|y_{1:k})   .
 \end{IEEEeqnarray}
@@
 Substituting the filtering approximation into this expression yields a particle distribution which may be sampled easily.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 \begin{algorithmic}[1]
- 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$.
+\linespread{1.7} \selectfont
+ 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$ and store the resulting particles $\{x_k^{(i)}, w_k^{(i)}\}$.
 	\FOR{$i = 1$ \TO $N_S$}
@@
 
-Following an equivalent derivation to that for (\ref{eq:backward_conditional_filter}), the target distribution may be approximated by a particle distribution.
-
+Following an equivalent derivation to that for (\ref{eq:backward_conditional_filter}), the target distribution may be approximated by a particle distribution,
+%
 \begin{IEEEeqnarray}{rCl}
 \hat{p}(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{1:k}^{(i)}}(x_{1:k}) \\
-\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:MCMC-BRS_weights}
+\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:MCMC-BRS_weights}   .
 \end{IEEEeqnarray}
@@
 A valid proposal may be constructed using the particles of the filter-smoother approximation with any arbitrary weights.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 If the current state of the chain is $x_{1:k}^{(m)}$, and a new sample is drawn from the proposal $x_{1:k}^{*} \sim q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability:
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-Note that $w_k^{*}$ indicates the particle weight of the proposed new state, $x_{1:k}^{*}$.
+Note that $w_k^{*}$ indicates the filtering particle weight of the proposed new state, $x_{1:k}^{*}$.
 
@@
 \begin{algorithmic}[1]
- 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
+\linespread{1.7} \selectfont
+ 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$ and store the resulting particles $\{x_{1:k}^{(i)}, w_k^{(i)}\}$.
 	\FOR{$i = 1$ \TO $N_S$}
@@
 Smoothers based on backward-resampling, both the direct sampler of \cite{Godsill2004} and the MCMC version outlined in the previous section, are limited by the fact that states are only selected from those which appear in the filtering approximation.
-If there is a significant difference between the filtering and smoothing densities, then the the particles are unlikely to be in the right locations to represent the smoothing distribution well, and the resulting approximation will be poor.
+If there is a significant difference between the filtering and smoothing densities, then the particles are unlikely to be in the right locations to represent the smoothing distribution well, and the resulting approximation will be poor.
 
@@
 Thus the factorised proposal may be written as follows:
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 The target distribution is also factorised.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-Using the filter-smoother approximation, the MH acceptance probabilities may then be calculated.
-
+Using the filter-smoother approximation for $p(x_{1:k-1}|y_{1:k-1})$, the MH acceptance probabilities may then be calculated.
+%
 \begin{IEEEeqnarray}{rCl}
@@
 
-As noted in \cite{Fearnhead2010}, the optimal choice of proposal is the conditional posterior state distribution,
+%As noted in \cite{Fearnhead2010}, the optimal choice of proposal is the conditional posterior state distribution,
 
-\begin{IEEEeqnarray}{rCl}
-q(x_k|x_{k-1},x_{k+1},y_k) & = & p(x_k|x_{k-1},x_{k+1},y_k)  .
-\end{IEEEeqnarray}
+%\begin{IEEEeqnarray}{rCl}
+%q(x_k|x_{k-1},x_{k+1},y_k) & = & p(x_k|x_{k-1},x_{k+1},y_k)  .
+%\end{IEEEeqnarray}
 
@@
 \begin{algorithmic}[1]
+\linespread{1.7} \selectfont
  	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
@@
 			\FOR{$m = 1$ \TO $M$}
-				\STATE Propose a new state hisory, $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
+				\STATE Propose a new state history,\\ $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
 				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
@@
 
-The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother.
+The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
 Because the support of the smoother particles is not restricted to the states which appear in the filter approximation, the error performance is expected to be improved.
+
+\subsection{Proposal Densities}
+
+When using MCMC, if samples can be proposed directly from the target distribution then the acceptance probability is always 1 (In combination with a block-sampling scheme, this becomes the Gibbs sampler of \cite{geman1984}).
+This is desirable as it maximises the acceptance rate.
+Clearly, for the basic MCMC backward-resampling smoother of section~\ref{sec:mcmc_smoother}, such a ``target density proposal'' simply yields the direct backward-resampling smoother of \cite{Godsill2004}, as described in section~\ref{sec:existing_smoothers}.
+For the MCMC backward-sampling smoother, a target density proposal can be made by sequentially sampling $x_{1:k-1}$ followed by $x_k$ using the following factorisation,
+%
+\begin{IEEEeqnarray}{rCl}
+\IEEEeqnarraymulticol{3}{l}{p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) } \nonumber \\
+                                    & \propto & p(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k) p(x_{1:k-1}|y_{1:k}, \tilde{x}_{k+1})   .
+\end{IEEEeqnarray}
+
+A good choice for the $x_k$ proposal is thus given by,
+%
+\begin{IEEEeqnarray}{rCl}
+q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k) & = & p(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)   .
+\end{IEEEeqnarray}
+
+The matching $x_{k-1}$ proposal can be approximated using the filter-smoother particles,
+%
+\begin{IEEEeqnarray}{rCl}
+q(x_{1:k-1}|y_{1:k}, \tilde{x}_{k+1}) & =       & p(x_{1:k-1}|y_{1:k}, \tilde{x}_{k+1}) \nonumber \\
+                                      & \propto & p(y_{k}, \tilde{x}_{k+1}| x_{1:k-1}) p(x_{1:k-1}|y_{1:k-1}) \nonumber \\
+                                      & \approx & \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})   .
+\end{IEEEeqnarray}
+
+\begin{IEEEeqnarray}{rCl}
+\tilde{v}_{k-1}^{(j)} & = & w_{k-1}^{(j)} p(y_{k}, \tilde{x}_{k+1}| x_{1:k-1}^{(j)})  .
+\end{IEEEeqnarray}
+
+The $x_k$ proposal density may be approximated with linear-Gaussian dynamics, e.g.
+using an extended \cite{Bar-Shalom2002} or unscented Kalman filter, \cite{Julier2004}.
+The weights for the $x_{k-1}$ proposal may be approximated using similar methods, but this will often be undesirable.
+If it is necessary to calculate proposal weights for all of the filter-smoother particles, then the computational advantages of using MCMC are lost.
+With this in mind, an appropriate choice for the proposal weights might be simply $\tilde{v}_{k-1}^{(j)} = w_{k-1}^{(j)}$.
 
@@
 The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model (see \cite{Bar-Shalom2002} for details).
-
+%
 %\begin{IEEEeqnarray}{rCl}
@@
 %\end{IEEEeqnarray}
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 The time between successive time steps is $\Delta t$.
-$I_2$ and $0_2$ denote the $2 \times 2$ identity and zero matrices.
+$I_2$ and $0_2$ denote the $2 \times 2$ identity and zero matrices respectively.
 The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ are independent and are zero-mean Gaussian distributed with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively.
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 $v_k^{(j)} = w_k^{(j)}$.
-The particle filter and MCMC-BSS both use ``optimal importance distributions'' with linearisations of the observation model where appropriate.
+The particle filter uses the ``optimal importance distributions'' and the MCMC-BSS a ``target density proposal'' with linearisations of the observation model where appropriate.
 
@@
 Other model settings are $\Delta t = 1$, $\sigma_P = 1$, and $x_0 = [-100, 50, 10, 0]^T$.
-Example trajectories and observations from the two cases are shown in figure~\ref{fig:example_trajectories}.
+Example trajectories and observations from the two cases are shown in Fig.~\ref{fig:example_trajectories}.
 
 \begin{figure}[!t]
-\centerline{\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory-crop.pdf}%
+\centering
+\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_trajectory.pdf}%
 \label{fig_first_case}}
-\hfil
-\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory-crop.pdf}%
-\label{fig_second_case}}}
-\caption{Example trajectory with observations from the two parameter cases.}
+\\
+\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_trajectory.pdf}%
+\label{fig_second_case}}
+\caption{Example trajectory with observations from the two noise covariance cases.}
 \label{fig:example_trajectories}
@@
 
-Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms with the two parameter cases, along with running times (excluding the time for filtering) and the mean number of unique particles at each time instant.
-The trade-off between speed and error is shown in figures~\ref{fig:case1_rmse_vs_time} and~\ref{fig:case2_rmse_vs_time}.
+Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms in the two noise scenarios, along with running times (excluding the time for filtering) and the mean number of unique smoother particles at each time instant.
+The trade-off between speed and error is shown in Figs.~\ref{fig:case1_rmse_vs_time} and~\ref{fig:case2_rmse_vs_time}.
 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Smoother performance measures for parameter case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
+\caption{Smoother performance measures for noise covariance case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
 \label{tab:case1_performance}
@@
 \hline
-Algorithm & \begin{minipage}[c]{1cm} Postion RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
+Algorithm & \begin{minipage}[c]{1cm} Position RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
 \hline
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Smoother performance measures for parameter case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
+\caption{Smoother performance measures for noise covariance case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
 \label{tab:case2_performance}
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{case1_smoother_comparison_posRMSE_time-crop.pdf}%
+\includegraphics[width=0.9\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}%
 \caption{RMSEs and running times of various smoothing algorithms}%
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{case2_smoother_comparison_posRMSE_time-crop.pdf}%
+\includegraphics[width=0.9\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}%
 \caption{RMSEs and running times of various smoothing algorithms}%
@@
 The MCMC-BRS smoother error approaches that of the DBRS as the number of MCMC steps, $M$, increases.
-However, for small $M$ the error is only very slightly larger than that of the DBRS but with significantly less running time.
-For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed.
+Similarly, the mean number of unique particles in the MCMC-BRS approximation approaches that of the DBRS.
+This behaviour is expected as a longer Markov chain is more likely to deliver an independent sample from the target distribution, whereas independent samples are guaranteed in the DBRS approximation due to the direct sampling procedure.
 
-The MCMC-BSS outperforms all the other smoothers, achieveing a lower RMSE even with $M=1$. As $M$ is increased further, the accuracy increases.
+When the number of MCMC steps, $M$, is small, the error of the MCMC-BRS is only very slightly larger than that of the DBRS but with a significantly shorter running time.
+For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed of computation.
+Thus, the MCMC-BRS provides increased flexibility when compared to the DBRS scheme.
+
+The MCMC-BSS outperforms all of the other smoothers, achieving the lowest RMSE for any given running time.
+Furthermore, because the support is not limited to using states from the filter distributions, the smoothing particle approximation has many more unique particles, giving a less degenerate representation.
 
@@
 The error performance of this scheme approaches that of the standard resampling scheme of \cite{Godsill2004} as the chain length increases.
-However, comparable error performance may be achieved with very short chains with a significant saving in processing time.
+However, comparable error performance may be achieved with very short chains, leading to a significant saving in processing time.
 
@@
 
-%%%% Template bits
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals use top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
-% stfloats package.
-
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-%\appendices
 %\section{Appendix Title}
@@
 
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-%\section{}
-%Appendix two text goes here.
-
-% use section* for acknowledgement
 %\section*{Acknowledgment}
-
 %The authors would like to thank...
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
+\bibliographystyle{IEEEtran}
+\bibliography{MCMCsmootherbib}
+
 \newpage
-\fi
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Bunch-bw.jpg}}]{Pete Bunch} (M'11) has an MEng.
+degree from the University of Cambridge, UK, and is currently cogitating towards a PhD degree in signal processing at the Cambridge University Engineering Department.
 
-% references section
+He is an enthusiastic Bayesian, with research interests in sequential inference, numerical approximations and target tracking.
+\end{IEEEbiography}
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-\bibliography{D:/pb404/Dropbox/PhD/OTbib}
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Godsill.jpg}}]{Simon Godsill} is a Professor of Statistical Signal Processing with the Engineering Department, Cambridge University, Cambridge, U.K.
+He is also a Professorial Fellow and tutor with Corpus Christi College Cambridge.
 
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
+He coordinates an active research group in Signal Inference and its Applications within the Signal Processing and Communications Laboratory at Cambridge University, specializing in Bayesian computational methodology, multiple object tracking, audio and music processing, and financial time-series modeling.
+A particular methodological theme over recent years has been the development of novel techniques for optimal Bayesian filtering, using Sequential Monte Carlo or Particle Filtering methods.
+He has published extensively in journals, books, and international conference proceedings.
 
-\begin{IEEEbiographynophoto}{Pete Bunch}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Simon Godsill}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
+Prof.
+Godsill has given a number of high profile invited and plenary addresses at conferences such as the Valencia conference on Bayesian Statistics and the IEEE Statistical Signal Processing Workshop.
+He was Technical Chair of the successful IEEE NSSPW workshop in 2006 on sequential and nonlinear filtering methods, and has been on the conference panel for numerous other conferences/workshops.
+He has served as an Associate Editor for the IEEE Transactions on Signal Processing and the journal Bayesian Analysis.
+He was Theme Leader in Tracking and Reasoning over Time for the UK’s Data and Information Fusion Defence Technology Centre (DIF-DTC) and Principal Investigator on many grants funded by the EU, EPSRC, QinetiQ, MOD, Microsoft UK, Citibank, and Mastercard.
+\end{IEEEbiography}
 

### 2012-02-22 MATLAB/plot_graphs_for_paper.m

@@ -1,0 +1,76 @@
+clup
+
+% Set test flag
+for test_flag = [1,3];
+    
+    test_case = test_flag;
+    if test_flag == 3
+        test_case = 2;
+    end
+    
+    % Load data
+    load(['smoother_test' num2str(test_flag) '.mat']);
+    
+    fig = figure; hold on
+    inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'ok', 'markersize', 7 )
+    inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'r^', 'markersize', 7 )
+    inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-b', 'markersize', 7 )
+    inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'x-.g', 'markersize', 7 )
+    
+    legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
+    xlabel('Running Time (s)'); ylabel('Position RMSE');
+    xlimits = get(gca, 'XLim');
+    ylimits = get(gca, 'YLim');
+    xlim([-50, xlimits(2)]);
+    plot([0 0], ylimits, ':k');
+    ylim(ylimits);
+    
+    wid = 4; hei = 3;
+    format_graph_for_pdf;
+    print(fig, '-dpdf', ['case' num2str(test_case) '_smoother_comparison_posRMSE_time.pdf']);
+    
+    %%% Trajectories
+    
+    % DEFINE RANDOM SEED
+    rand_seed = 1;
+    
+    % Set random seed
+    s = RandStream('mt19937ar', 'seed', rand_seed);
+    RandStream.setDefaultStream(s);
+    
+    % Set parameters
+    set_parameters;
+    
+    if test_flag == 1
+        params.bng_var = (pi/720)^2;
+        params.rng_var = 0.1;
+    elseif test_flag == 2
+        params.bng_var = (pi/180)^2;
+        params.rng_var = 10;
+    elseif test_flag == 3
+        params.bng_var = (pi/36)^2;
+        params.rng_var = 100;
+    end
+    
[…]

### 2012-02-23 Paper/complete/finaldraft_MCMCsmoother.tex
small corrections

@@
 
-The matching $x_{k-1}$ proposal can be approximated using the filter-smoother particles,
+The matching $x_{1:k-1}$ proposal can be approximated using the filter-smoother particles,
 %
@@
 The $x_k$ proposal density may be approximated with linear-Gaussian dynamics, e.g.
-using an extended \cite{Bar-Shalom2002} or unscented Kalman filter, \cite{Julier2004}.
-The weights for the $x_{k-1}$ proposal may be approximated using similar methods, but this will often be undesirable.
+using an extended \cite{Bar-Shalom2002} or unscented \cite{Julier2004} Kalman filter.
+The weights for the $x_{1:k-1}$ proposal may be approximated using similar methods, but this will often be undesirable.
 If it is necessary to calculate proposal weights for all of the filter-smoother particles, then the computational advantages of using MCMC are lost.
@@
 Results are shown here for high and low observation noise cases.
-For case 1, the observation variances were set to $\sigma_B = \pi/720$ and $\sigma_R = 0.1$.
-For case 2, $\sigma_B = \pi/36$ and $\sigma_R = 100$.
+For case 1, the observation variances were set to $\sigma_B^2 = (\pi/720)^2$ and $\sigma_R^2 = 0.1$.
+For case 2, $\sigma_B = (\pi/36)^2$ and $\sigma_R^2 = 100$.
 Other model settings are $\Delta t = 1$, $\sigma_P = 1$, and $x_0 = [-100, 50, 10, 0]^T$.

### 2012-02-28 Paper/complete/finaldraft_MCMCsmoother.tex
Added rejection sampling to code. Added references to Douc2009 to paper.

@@
 In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation.
-In this paper, a Markov chain Monte Carlo sampling procedure is used to improve the efficiency of the particle smoother.
-In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented.
+In this paper, a Markov chain Monte Carlo sampling procedure is used to improve the efficiency of the particle smoother, achieving comparable error performance but with a lower execution time.
+In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error.
+These algorithms also provide a greater degree of flexibility over existing methods, allowing a trade-off between execution time and error, controlled by the length of the Markov chains.
 \end{abstract}
@@
 The complexity of both the backward-resampling smoother of \cite{Godsill2004} and the two-filter variant of \cite{Briers2010} is $\mathcal{O}(N_F \times N_S \times K)$, where $N_F$ is the number of filter particles, $N_S$ the number of smoother particles, and $K$ the number of time steps.
+In \cite{Douc2009}, a theoretically linear-cost scheme using rejection sampling is outlined, but in practice this appears to be slower than the standard algorithms.
 Methods for approximation of such algorithms with lower complexity are discussed in \cite{Klaas2006}.
-However, computational cost is often prohibitive, requiring either $N_S$ or $N_F$ to be held very low.
+However, computational cost is often still prohibitive, requiring either $N_S$ or $N_F$ to be held very low.
 
@@
 
+The new scheme targets the same empirical approximation to the smoothing distribution as used in the direct backward-resampling method.
+However, rather than drawing independent samples from this approximation, MCMC is used, resulting in dependent samples.
+In the limit as the number of MH steps, $M$, tends to infinity, the Markov chain will produce an independent sample from the target distribution, as the initial state of the chain will have been forgotten.
+At this point, the performance of the MCMC method is expected to match that of the direct method in terms of error and particle diversity.
+At the other extreme, with $M=0$, the algorithm simply outputs particles from the filter-smoother approximation, which is expected to give worse performance.
+With intermediate values of $M$, the quality of the approximation is expected to fall between those of the filter-smoother and direct backward-resampling smoother.
+
+This new scheme bears a resemblance to the ``linear cost'' simulation scheme of \cite{Douc2009} which uses rejection sampling instead of MCMC.
+Although theoretically elegant, the latter algorithm has been found to suffer from such high rejection rates as to render it consistently slower than the direct sampling implementation on a practical problem (see section~\ref{sec:simulations}).
+
 \section{An MCMC particle smoother with improved support} \label{sec:new_state_smoother}
@@
 The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
-Because the support of the smoother particles is not restricted to the states which appear in the filter approximation, the error performance is expected to be improved.
+In contrast with the MCMC resampling method, this smoother targets the exact posterior distribution for the current state at time $t_n$, rather than a particle approximation.
+Consequently, the performance is expected to be improved relative to the resampling-based methods in terms of both error and particle diversity.
 
@@
 If it is necessary to calculate proposal weights for all of the filter-smoother particles, then the computational advantages of using MCMC are lost.
-With this in mind, an appropriate choice for the proposal weights might be simply $\tilde{v}_{k-1}^{(j)} = w_{k-1}^{(j)}$.
+With this in mind, an appropriate choice for the proposal weights might be simply $\tilde{v}_{k-1}^{(j)} = w_{k-1}^{(j)}$, which avoids calculation of $p(y_{k}, \tilde{x}_{k+1}| x_{1:k-1}^{(j)})$ for all $j$.
 
@@
 
+The rejection-sampling version of the DBRS described by \cite{Douc2009} was also implemented, but high rejection rates meant that it was impractically slow so it was omitted from further testing.
+
 The algorithms were all used to generate 100 particles from the joint smoothing distribution, using particle filters with 100 particles.
@@
 Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms in the two noise scenarios, along with running times (excluding the time for filtering) and the mean number of unique smoother particles at each time instant.
-The trade-off between speed and error is shown in Figs.~\ref{fig:case1_rmse_vs_time} and~\ref{fig:case2_rmse_vs_time}.
+The trade-off between speed and error is shown in Fig.~\ref{fig:rmse_vs_time}.
 
@@
 MCMC-BRS (30)   & 0.480 & 0.761 & 62 & 20.2 \\
-MCMC-BRS (100)& 0.476 & 0.764 & 202 & 20.4 \\
+MCMC-BRS (100)  & 0.476 & 0.764 & 202 & 20.4  \\
 \hline
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}%
-\caption{RMSEs and running times of various smoothing algorithms}%
-\label{fig:case1_rmse_vs_time}%
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=0.9\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}%
-\caption{RMSEs and running times of various smoothing algorithms}%
-\label{fig:case2_rmse_vs_time}%
+\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
+\\
+\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%
+\label{fig:rmse_vs_time}%
 \end{figure}

### 2012-02-28 MATLAB/backward_rejectionsampling_smoother.m

@@ -1,0 +1,65 @@
+function [ smooth_pts ] = backward_rejectionsampling_smoother( S, times, pts_array, wts_array, h_trans )
+%BACKARD_SAMPLING_SMOOTHER Godsill2004-type backward sampling smoother
+%using rejection sampling
+%
+%
+%
+global params; %UGLY
+
+fprintf(1, '\n\n*** Running direct backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Get weights
+        wts = wts_array{kk};
+        
+        % Rejection sample
+        [~, Q] = build_AQ(dt, params.proc_var);  % This is ugly and should be encapsulated.
+        max_trans_prb = log(det(2*pi*Q));
+        u = inf; thresh = 1;
+        count = 0;
+        while u > thresh
+            
+            % Sample an index and a uniform variable
+            ind = sample_weights(exp(wts));
+            u = log(rand);
+            
+            % Calculate the acceptance threshold
+            [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{ind}(:,kk), smooth_pts{ii}(:,kk+1) );
+            thresh = trans_prb - max_trans_prb;
+            
+            count = count + 1;
+            
+        end
+        
+        count
+        
+        % Update trajectory
+        smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+        
+    end
+    
+end
+
+end
+

### 2012-02-28 MATLAB/backward_sampling_smoother.m

@@ -1,0 +1,61 @@
+function [ smooth_pts ] = backard_sampling_smoother( S, times, pts_array, wts_array, h_trans )
+%BACKARD_SAMPLING_SMOOTHER Normal Godsill2004-type backward sampling
+%smoother.
+%
+%
+%
+
+fprintf(1, '\n\n*** Running direct backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Get weights
+        wts = wts_array{kk};
+        
+        % Loop through filter particles
+        for jj = 1:length(pts_array{kk})
+            
+            % Calculate weight
+            [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{jj}(:,kk), smooth_pts{ii}(:,kk+1) );
+            wts(jj) = wts(jj) + trans_prb;
+            
+        end
+        
+        % Normalise weights
+        wts = wts - max(wts);
+        lin_wts = exp(wts); lin_wts = lin_wts/sum(lin_wts);
+        wts = log(lin_wts);
+        
+        % Sample
+%         fprintf(1, '     ESS in frame %u: %f\n', kk, ESS(wts));
+        ind = sample_weights(exp(wts));
+        
+        % Update trajectory
+        smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+        
+    end
+    
+end
+
+end
+

### 2012-02-28 MATLAB/run_mcmcsmoother_demo.m

@@ -43,7 +43,7 @@ cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), kiti_pts)
 
 %% Run an ordinary smoother
 tic;
-bs_smooth_pts = backard_sampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
+bs_smooth_pts = backward_sampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
 bs_time = toc;
 
 % Plot it
@@ -51,13 +51,23 @@ figure(5); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
 plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
 cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), bs_smooth_pts)
 
+%% Run a rejection-sampling smoother
+tic;
+brs_smooth_pts = backward_rejectionsampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
+brs_time = toc;
+
+% Plot it
+figure(6); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), brs_smooth_pts)
+
 %% Run the MCMC smoother
 tic;
 mcmc_smooth_pts = mcmc_smoother( params.S, params.M, t, pts_array, wts_array, @tracking_trans );
 mcmc_time = toc;
 
 % Plot it
-figure(6); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+figure(7); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
 plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
 cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), mcmc_smooth_pts)
 
@@ -67,7 +77,7 @@ mcmc_newstate_smooth_pts = mcmc_newstate_smoother( params.S, params.M, t, pts_ar
 mcmc_ns_time = toc;
 
 % Plot it
-figure(7); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
[…]

### 2012-04-03 Paper/complete/submit_double_finaldraft_MCMCsmoother.tex
Submitted paper version

--- 2012-02-28 Paper/complete/finaldraft_MCMCsmoother.tex
@@
-\documentclass[journal]{IEEEtran}
+\documentclass[10pt,twocolumn,twoside]{IEEEtran}
 
@@
 
-\author{Pete~Bunch,~\IEEEmembership{Member,~IEEE,} and
+\author{Pete~Bunch*,~\IEEEmembership{Member,~IEEE,} and
         Simon~Godsill,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
@@
 of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}% <-this % stops a space
-\thanks{Manuscript received January 01, 1901; revised January 02, 1901.}}
+\thanks{Manuscript received April 03, 2012.}}
 
-% The paper headers
-\markboth{IEEE Transaction in Signal Processing,~Vol.~1, No.~1, January~1901}%
-{Bunch \& Godsill \MakeLowercase{\textit{et al.}}: Improved Particle Approximations to the Joint Smoothing Distribution Using Markov Chain Monte Carlo}
+%% The paper headers
+%\markboth{IEEE Transactions on Signal Processing,~Vol.~AA, No.~BB, January~YYYY}%
+%{MCMC Particle Smoothing Approximations}
 
@@
 
-Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$, and $w_k{(i)}$ is the weight of the $i$th particle.
+Here, $\delta_{a}(x)$ indicates a discrete probability mass at the point $x = a$, and $w_k^{(i)}$ is the weight of the $i$th particle.
 Weights are non-negative and sum to 1.% $\hat{p}$ means an approximation to $p$.
@@
 
-The proposal density for $x_k$, $q(x_{x}|x_{k-1})$, is often called the ``importance density''.
-Common choices include the model transition density (\ref{eq:state_proc}), and the ``optimal importance distribution'' \cite{Doucet2000a}.
+The proposal for $x_k$, $q(x_{x}|x_{k-1})$, is often called the ``importance density''.
+Common choices for this include the model transition density (\ref{eq:state_proc}), and the ``optimal importance density'' \cite{Doucet2000a}.
 The proposal density for $x_{k-1}$, $q(x_{k-1})$, is constructed from the particles of the filtering approximation from the previous time step.
@@
 The output from the final processing step is an approximation to the desired smoothing distribution.
+This simple filter-smoother is summarised in Fig.~\ref{alg:FS}.
 
-The weakness of this simple, filter-smoother is degeneracy of the particles in the path-space.
+The weakness of the filter-smoother is degeneracy of the particles in the path-space.
 Every time the particles are resampled, those with low weights do not appear in the approximation at the next step, whereas those with high weights appear multiple times.
 The result is that many, or even all, particles will share a common ancestor, before which they all have the same set of state values.
-If the objective is only to produce a single point estimate then this might be sufficient, but to characterise the complete smoothing distribution, the path-space diversity of the approximation must be improved.
-
-The filter-smoother is summarised in Fig.~\ref{alg:FS}.
+If the objective is only to produce a single point estimate then this might be sufficient, but to characterise the complete smoothing distribution, the path-space diversity of the approximation must be increased.
 
@@
 				\STATE Propose a new state history,\\ $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
-				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
+				\STATE Propose a new state, $x_{k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
 				\STATE With probability $\alpha$ specified by (\ref{eq:MCMC-BSS_ap}),\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
@@
 The $x_k$ proposal density may be approximated with linear-Gaussian dynamics, e.g.
-using an extended \cite{Bar-Shalom2002} or unscented \cite{Julier2004} Kalman filter.
+using an extended \cite{Grewal2002} or unscented \cite{Julier2004} Kalman filter.
 The weights for the $x_{1:k-1}$ proposal may be approximated using similar methods, but this will often be undesirable.
@@
 $v_k^{(j)} = w_k^{(j)}$.
-The particle filter uses the ``optimal importance distributions'' and the MCMC-BSS a ``target density proposal'' with linearisations of the observation model where appropriate.
+The particle filter uses the ``optimal importance density'' and the MCMC-BSS a ``target density proposal'' with linearisations of the observation model where appropriate.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{MCMCsmootherbib}
+\bibliography{D:/pb404/Dropbox/PhD/Cleanbib}
 
-\newpage
+%\newpage
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Bunch-bw.jpg}}]{Pete Bunch} (M'11) has an MEng.
-degree from the University of Cambridge, UK, and is currently cogitating towards a PhD degree in signal processing at the Cambridge University Engineering Department.
+%\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Bunch.tex}
+%\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Godsill.tex}
 
-He is an enthusiastic Bayesian, with research interests in sequential inference, numerical approximations and target tracking.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Godsill.jpg}}]{Simon Godsill} is a Professor of Statistical Signal Processing with the Engineering Department, Cambridge University, Cambridge, U.K.
-He is also a Professorial Fellow and tutor with Corpus Christi College Cambridge.
-
-He coordinates an active research group in Signal Inference and its Applications within the Signal Processing and Communications Laboratory at Cambridge University, specializing in Bayesian computational methodology, multiple object tracking, audio and music processing, and financial time-series modeling.
-A particular methodological theme over recent years has been the development of novel techniques for optimal Bayesian filtering, using Sequential Monte Carlo or Particle Filtering methods.
-He has published extensively in journals, books, and international conference proceedings.
-
-Prof.
-Godsill has given a number of high profile invited and plenary addresses at conferences such as the Valencia conference on Bayesian Statistics and the IEEE Statistical Signal Processing Workshop.
-He was Technical Chair of the successful IEEE NSSPW workshop in 2006 on sequential and nonlinear filtering methods, and has been on the conference panel for numerous other conferences/workshops.
-He has served as an Associate Editor for the IEEE Transactions on Signal Processing and the journal Bayesian Analysis.
-He was Theme Leader in Tracking and Reasoning over Time for the UK’s Data and Information Fusion Defence Technology Centre (DIF-DTC) and Principal Investigator on many grants funded by the EU, EPSRC, QinetiQ, MOD, Microsoft UK, Citibank, and Mastercard.
-\end{IEEEbiography}
-
-% that's all folks
 \end{document}

### 2012-04-03 Paper/complete/submit_single_finaldraft_MCMCsmoother.tex
Submitted paper version

--- 2012-04-03 Paper/complete/submit_double_finaldraft_MCMCsmoother.tex
@@
-\documentclass[10pt,twocolumn,twoside]{IEEEtran}
+\documentclass[peerreview,11pt,draftcls,onecolumn]{IEEEtran}
 
@@
 
-\begin{IEEEkeywords}
-state space model, particle filter, smoothing, MCMC, Bayesian inference
-\end{IEEEkeywords}
+%\begin{IEEEkeywords}
+%state space model, particle filter, smoothing, MCMC, Bayesian inference
+%\end{IEEEkeywords}
+
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+
+\ifCLASSOPTIONpeerreview
+\begin{center} \bfseries EDICS Categories: MLR-BAYL, SSP-NGAU, SSP-TRAC, SSP-FILT \end{center}
+\fi
+
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
 
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_trajectory.pdf}%
+\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_trajectory.pdf}%
 \label{fig_first_case}}
-\\
-\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_trajectory.pdf}%
+\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_trajectory.pdf}%
 \label{fig_second_case}}
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
-\\
-\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
 \caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%

### 2012-04-03 MATLAB/set_parameters.m

@@ -4,8 +4,8 @@ global params;
 params.dt = 1;
 params.N = 100;
 params.proc_var = 1;
-params.bng_var = (pi/720)^2;    % Try (pi/720)^2; (pi/180)^2;
-params.rng_var = 0.1;           % Try 0.1;        10;
+params.bng_var = (pi/36)^2;    % Try (pi/720)^2; (pi/180)^2;
+params.rng_var = 100;           % Try 0.1;        10;
 params.x0 = [-100; 50; 10; 0];
 
 % Prior

### 2012-06-05 Paper/complete/revisions_MCMCsmoother.tex
Post-review revisions and MATLAB

--- 2012-04-03 Paper/complete/submit_double_finaldraft_MCMCsmoother.tex
@@
 
-In a particle filter, the filtering distribution is approximated by a set of weighted samples (or ``particles'') drawn from that distribution.
+In a particle filter, the filtering distribution is approximated by a set of weighted samples (or ``particles'') drawn from that distribution,
 %
 \begin{IEEEeqnarray}{rCl}
-\hat{p}(x_{k}|y_{1:k}) & = & \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)
+\hat{p}(x_{k}|y_{1:k}) & = & \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)     .
 \end{IEEEeqnarray}
@@
 Similar ideas apply to the problem of smoothing.
-In the linear Gaussian case, the entire state sequence, $x_{1:K}$, may be estimated in closed-form using the a Rauch-Tung-Striebel (RTS) smoother \cite{Rauch1965}.
+In the linear Gaussian case, the entire state sequence, $x_{1:K}$, may be estimated in closed-form using the Rauch-Tung-Striebel (RTS) smoother \cite{Rauch1965}.
 This works by modifying the ordinary Kalman filter results in a backward processing pass through the state sequence.
 For nonlinear or non-Gaussian models, it is possible again to resort to numerical methods.
-Now, particles are used to approximate the joint smoothing distribution.
+Now, particles are used to approximate the joint smoothing distribution,
 %
 \begin{IEEEeqnarray}{rCl}
-\hat{p}(x_{1:K}|y_{1:K}) & = & \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})
+\hat{p}(x_{1:K}|y_{1:K}) & = & \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})     .
 \end{IEEEeqnarray}
@@
 The choice of weights and sampling procedure for $x_{k-1}$ determines what from of ``resampling'' is to be used.
-Here we use arbitrary weights for generality.
+Here we use arbitrary weights for generality,
 %
 \begin{IEEEeqnarray}{rCl}
-q(x_{k-1}) & = & \sum_i v_{k-1}^{(i)} \delta_{x_{k-1}^{(i)}}(x_{k-1})
+q(x_{k-1}) & = & \sum_i v_{k-1}^{(i)} \delta_{x_{k-1}^{(i)}}(x_{k-1})     .
 \end{IEEEeqnarray}
@@
 
-Next, the particles are assigned an importance weight according to the ratio of the target and the proposal densities.
+Next, the particles are assigned an importance weight according to the ratio of the target and the proposal densities,
 %
@@
             & \propto & \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) p(x_{k-1}^{(i)}|y_{1:k-1}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) q(x_{k-1}^{(i)}) } \nonumber \\
-            & \approx & \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }
+            & =       & \frac{ p(y_k|x_k^{(i)}) p(x_k^{(i)}|x_{k-1}^{(i)}) }{ q(x_{x}^{(i)}|x_{k-1}^{(i)}) } \times \frac{w_{k-1}^{(i)}}{v_{k-1}^{(i)} }     .
 \end{IEEEeqnarray}
 
-Normalisation is enforced by scaling the weights so that they sum to 1. Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the $x_{k-1}^{(i)}$ values from each particle.
-
+Normalisation is enforced by scaling the weights so that they sum to 1.
+Finally, $x_{k-1}$ is marginalised from the distribution by simply discarding the $x_{k-1}^{(i)}$ values from each particle.
 The particle filter is summarised in Fig.~\ref{alg:PF}.
 
-%\begin{algorithm}
 \begin{figure}
@@
 \end{figure}
-%\end{algorithm}
 
@@
 
-%\begin{algorithm}
 \begin{figure}
@@
 This works by constructing new particles from the smoothing distribution by resampling from the filtering approximations.
-The sampling procedure exploits the following factorisation of the joint distribution.
+The sampling procedure exploits the following factorisation of the joint distribution,
 %
 \begin{IEEEeqnarray}{rCl}
-p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K}) \label{eq:smoothing_factorisation}
+p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K}) \label{eq:smoothing_factorisation}     .
 \end{IEEEeqnarray}
@@
 
-Substituting the filtering approximation into this expression yields a particle distribution which may be sampled easily.
+Substituting the filtering approximation into this expression yields a particle distribution which may be sampled easily,
 %
@@
 \hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) = \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k}) \label{eq:backward_conditional_filter} \\
-\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:DBRS_weights}
+\tilde{w}_k^{(i)} \propto w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)}) \label{eq:DBRS_weights}     .
 \end{IEEEeqnarray}
@@
 The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K})$, by combination of the results from one forward and one backward particle filter.
-Particles from the joint distribution may then be generated by resampling from either the forward or backward filters (or some combination) in the manner of the backward-resampling smoother outlined above.%algorithm~\ref{alg:DBRS}.
+Particles from the joint distribution may then be generated by resampling from either the forward or backward filters (or some combination) in the manner of the backward-resampling smoother outlined above.
 
 The complexity of both the backward-resampling smoother of \cite{Godsill2004} and the two-filter variant of \cite{Briers2010} is $\mathcal{O}(N_F \times N_S \times K)$, where $N_F$ is the number of filter particles, $N_S$ the number of smoother particles, and $K$ the number of time steps.
-In \cite{Douc2009}, a theoretically linear-cost scheme using rejection sampling is outlined, but in practice this appears to be slower than the standard algorithms.
-Methods for approximation of such algorithms with lower complexity are discussed in \cite{Klaas2006}.
-However, computational cost is often still prohibitive, requiring either $N_S$ or $N_F$ to be held very low.
+In \cite{Douc2009}, a theoretically linear-cost scheme using rejection sampling is outlined, but in practice this often appears to be slower than the standard algorithms due to high rejection rates.
+Methods for reducing the computational complexity of smoothing algorithms are also discussed in \cite{Klaas2006}.
+However, these focus on approximating a sum of kernel values over a set of particles, such as those which arise in forward-backward smoothing \cite{Doucet2000a} and two-filter smoothing \cite{Briers2010} for estimation of the marginal smoothing distribution.
+They are not applicable for the joint smoothing procedures.
 
@@
 This limits the support of the smoother to the states that appear in the filtering particles.
-In \cite{Fearnhead2010}, an IS-based algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal, allowing full support.
-In addition, this algorithm achieves $\mathcal{O}( N_S \times K)$ complexity, linear in the number of particles.
-
-%In section~\ref{sec:mcmc_smoother}, we introduce an MCMC approach for drawing samples from the joint smoothing distribution with an improved computational complexity over the direct backward-resampling smoother.
-In section~\ref{sec:new_state_smoother}, we combine this procedure with the new-proposal method used in \cite{Fearnhead2010} to generate joint smoothing distribution particles with full support.
+In \cite{Fearnhead2010}, an IS-based algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal density, thus recovering the full support.
+In addition, this algorithm achieves $\mathcal{O}( N_S \times K)$ complexity, i.e.
+linear in the number of particles.
 
@@
 
-A valid proposal may be constructed using the particles of the filter-smoother approximation with any arbitrary weights.
+A valid proposal may be constructed using the particles of the filter-smoother approximation with any arbitrary weights,
 %
 \begin{IEEEeqnarray}{rCl}
-q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})
+q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})     .
 \end{IEEEeqnarray}
 
-If the current state of the chain is $x_{1:k}^{(m)}$, and a new sample is drawn from the proposal $x_{1:k}^{*} \sim q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability:
+If the current state of the chain is $x_{1:k}^{(m)}$, and a new sample is drawn from the proposal $x_{1:k}^{*} \sim q(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$, then it should be accepted with the following probability,
 %
@@
 Note that $w_k^{*}$ indicates the filtering particle weight of the proposed new state, $x_{1:k}^{*}$.
-
 The final state of each chain is used as a sample from the required conditional distribution.
-
 The complete MCMC backward-resampling smoother is summarised in Fig.~\ref{alg:MCMC-BRS}.
 
-%\begin{algorithm}
 \begin{figure}
@@
 
-%The method described in \cite{Fearnhead2010} uses a forward and a backward particle filter.
-To develop a marginal smoothing approximation at time step $k$, particles are resampled from the forward filter at step $k-1$, and from the backward filter and step $k+1$.
-A new state for step $k$ is then sampled from a continuous proposal distribution conditional on the previous and next states, and an importance weight is assigned.
-It is undesirable to use importance sampling for generating particles from the joint sampling distribution, as the required resampling is liable to lead to a loss of particle diversity similar to that suffered by the filter-smoother.
-
 The smoother described in \cite{Fearnhead2010} uses importance sampling to approximate the marginal smoothing distributions.
@@
 The current state, $x_k$, is then sampled afresh from a new proposal distribution.
-Thus the factorised proposal may be written as follows:
+Thus the factorised proposal may be written as follows,
 %
@@
 
-The target distribution is also factorised.
+The target distribution is also factorised,
 %
@@
 \IEEEeqnarraymulticol{3}{l}{p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K}) = p(x_{1:k}|\tilde{x}_{k+1}, y_{1:k}) } \nonumber \\
-                                    & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k-1}|y_{1:k-1})
+                                    & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_x) p(x_{1:k-1}|y_{1:k-1})     .
 \end{IEEEeqnarray}
 
-Using the filter-smoother approximation for $p(x_{1:k-1}|y_{1:k-1})$, the MH acceptance probabilities may then be calculated.
+Using the filter-smoother approximation for $p(x_{1:k-1}|y_{1:k-1})$, the MH acceptance probabilities may then be calculated,
 %
@@
  &=& \min \Bigg[ 1, \frac{ w_{k-1}^{*} p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) }{ w_{k-1}^{(m)} p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) } \nonumber \\
- & & \times \frac{ \tilde{v}_{k-1}^{(m)} q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ \tilde{v}_{k-1}^{*} q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } \Bigg] \label{eq:MCMC-BSS_ap}
-% \frac{ w_{k-1}^{*} \tilde{v}_{k-1}^{(m)} }{  w_{k-1}^{(m)} \tilde{v}_{k-1}^{*} } \times \nonumber \\
-% \IEEEeqnarraymulticol{3}{l}{\frac{ p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k^{(m)}) p(x_k^{(m)}|x_{k-1}^{(m)}) p(y_k|x_k^{(m)}) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } } .
+ & & \times \frac{ \tilde{v}_{k-1}^{(m)} q(x_k^{(m)}|x_{k-1}^{(m)},\tilde{x}_{k+1},y_k) }{ \tilde{v}_{k-1}^{*} q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) } \Bigg] \label{eq:MCMC-BSS_ap}     .
 \end{IEEEeqnarray}
 
-%As noted in \cite{Fearnhead2010}, the optimal choice of proposal is the conditional posterior state distribution,
+The complete MCMC backward-sampling smoother is summarised in Fig.~\ref{alg:MCMC-BSS}.
+The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
+In contrast with the MCMC resampling method, this smoother targets the exact posterior distribution for the current state at time $t_n$, rather than a particle approximation.
+Consequently, the performance is expected to be improved relative to the resampling-based methods in terms of both error and particle diversity.
 
-%\begin{IEEEeqnarray}{rCl}
-%q(x_k|x_{k-1},x_{k+1},y_k) & = & p(x_k|x_{k-1},x_{k+1},y_k)  .
-%\end{IEEEeqnarray}
+We note that this new scheme bears a resemblance to the improvement strategy for approximations to the joint smoothing distribution published recently in \cite{Dubarry2011}.
+Their algorithm iteratively applies MH steps to a low-diversity particle approximation, such as that from the filter-smoother.
+Our algorithm differs in the fact that it allows simultaneous changes to the particle state history, $x_{1:n-1}$, as well as the current state.
+Furthermore, for each smoothing particle, we run a Markov chain at each time step, using one sweep backwards through time, rather than multiple sweeps with a single MH step at each time point.
 
-The complete MCMC backward-sampling smoother is summarised in Fig.~\ref{alg:MCMC-BSS}.
-
-%\begin{algorithm}
 \begin{figure}
@@
 \end{figure}
-%\end{algorithm}
-
-The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
-In contrast with the MCMC resampling method, this smoother targets the exact posterior distribution for the current state at time $t_n$, rather than a particle approximation.
-Consequently, the performance is expected to be improved relative to the resampling-based methods in terms of both error and particle diversity.
 
@@
                                       & \propto & p(y_{k}, \tilde{x}_{k+1}| x_{1:k-1}) p(x_{1:k-1}|y_{1:k-1}) \nonumber \\
-                                      & \approx & \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})   .
+                                      & \approx & \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})
 \end{IEEEeqnarray}
-
+%
 \begin{IEEEeqnarray}{rCl}
@@
 The performance of the new smoothers was tested on a simple 2D tracking model.
-The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model (see \cite{Bar-Shalom2002} for details).
-%
-%\begin{IEEEeqnarray}{rCl}
-%\underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \underbrace{\begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix}}_{\mathbf{A}} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
-%\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}
-%\end{IEEEeqnarray}
+The trajectory of a single target was simulated using linear-Gaussian dynamics and observed via a nonlinear radar-type measurement model (see \cite{Bar-Shalom2002} for details),
 %
@@
 \underbrace{\begin{bmatrix}x_k \\ y_k \\ \dot{x}_k \\ \dot{y}_k\end{bmatrix}}_{\mathbf{x}_k} &=& \begin{bmatrix}I_2 & \Delta t I_2 \\ 0_2 & I_2\end{bmatrix} \underbrace{\begin{bmatrix}x_{k-1} \\ y_{k-1} \\ \dot{x}_{k-1} \\ \dot{y}_{k-1}\end{bmatrix}}_{\mathbf{x}_{k-1}} + \underbrace{\begin{bmatrix}w_{1,k} \\ w_{2,k} \\ w_{3,k} \\ w_{4,k}\end{bmatrix}}_{\mathbf{w}_k} \\
-\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}
+\underbrace{\begin{bmatrix}b_k \\ r_x\end{bmatrix}}_{\mathbf{y}_k} &=& \begin{bmatrix}\tan^{-1}(y_k / x_k) \\ \sqrt{x_k^2+y_k^2} \end{bmatrix} + \underbrace{\begin{bmatrix}v_{1,k} \\ v_{2,k}\end{bmatrix}}_{\mathbf{v}_k}     .
 \end{IEEEeqnarray}
@@
 $I_2$ and $0_2$ denote the $2 \times 2$ identity and zero matrices respectively.
-The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ are independent and are zero-mean Gaussian distributed with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively.
+The random variables $\mathbf{w}_k$ and $\mathbf{v}_k$ are independent and are zero-mean Gaussian distributed with covariance matrices $\mathbf{Q}$ and $\mathbf{R}$ respectively,
 %
@@
 \mathbf{Q} &=& \sigma^2_P \begin{bmatrix}\frac{\Delta t^3}{3} I_2 & \frac{\Delta t^2}{2} I_2 \\ \frac{\Delta t^2}{2} I_2 & \Delta t I_2 \end{bmatrix} \\
-\mathbf{R} &=& \begin{bmatrix}\sigma^2_B & 0 \\ 0 & \sigma^2_R\end{bmatrix}
+\mathbf{R} &=& \begin{bmatrix}\sigma^2_B & 0 \\ 0 & \sigma^2_R\end{bmatrix}     .
 \end{IEEEeqnarray}
@@
 
-Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms in the two noise scenarios, along with running times (excluding the time for filtering) and the mean number of unique smoother particles at each time instant.
+The performances of the various smoothing algorithms are compared using a number of measures.
+At each time point, a single state estimate is calculated by taking the mean of the values from each particle.
+These point estimates are then used to calculate a root-mean-square error (RMSE).
+In addition, a modified normalised estimation error squared (MNEES) is calculated using,
+%
+\begin{IEEEeqnarray}{rCl}
+ \text{MNEES}_k & = & (\hat{\mathbf{x}}_k - \mathbf{x}_k^*)^T \hat{P}_k^{-1} (\hat{\mathbf{x}}_k - \mathbf{x}_k^*) \nonumber \\
+ \hat{\mathbf{x}}_k & = & \frac{1}{N_S} \sum_i \mathbf{x}_k^{(i)} \nonumber \\
+ \hat{P}_k & = & \frac{1}{N_S} \sum_i (\mathbf{x}^{(i)}_k - \mathbf{x}_k^*) (\mathbf{x}^{(i)}_k - \mathbf{x}_k^*)^T
+\end{IEEEeqnarray}
+
+\noindent and where $\mathbf{x}_k^*$ is the true value of the state.
+This statistic has a value of $1$ when there is only one unique particle in the approximation, and tends to $0$ as the number of independent particles tends to infinity.
+It provides a relative measure of the quality of different particle approximations, taking into account both estimation error and particle diversity.
+Finally, the mean number of unique particles at each time instant and the running times (excluding the time for filtering) are also recorded.
+These performance indicators are all shown in tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} for the two noise scenarios.
+The trade-off between speed and error is shown in Fig.~\ref{fig:rmse_vs_time}.
+
+%Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms in the two noise scenarios, along with running times (excluding the time for filtering) and the mean number of unique smoother particles at each time instant.
 The trade-off between speed and error is shown in Fig.~\ref{fig:rmse_vs_time}.
@@
 \label{tab:case1_performance}
+{\scriptsize
 \centering
-\begin{tabular}{|c||c|c|c|c|}
+\begin{tabular}{|c||c|c|c|c|c|}
 \hline
-Algorithm & \begin{minipage}[c]{1cm} Position RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
+Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  MNEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
 \hline
-FS 						& 0.578 & 0.968 & 0 & 2.25 \\
-DBRS					& 0.475 & 0.762 & 122 & 20.3 \\
+FS              & 0.578 & 0.968 &  & 0 & 2.25 \\
+DBRS            & 0.475 & 0.762 &  & 122 & 20.3 \\
 \hline
-MCMC-BRS (1)    & 0.509 & 0.823 & 3 & 13.7  \\
-MCMC-BRS (3)    & 0.493 & 0.790 & 7 & 17.4  \\
-MCMC-BRS (10)   & 0.482 & 0.768 & 21 & 19.6 \\
-MCMC-BRS (30)   & 0.480 & 0.761 & 62 & 20.2 \\
-MCMC-BRS (100)  & 0.476 & 0.764 & 202 & 20.4  \\
+MCMC-BRS (1)    & 0.509 & 0.823 &  & 3 & 13.7  \\
+MCMC-BRS (3)    & 0.493 & 0.790 &  & 7 & 17.4  \\
+MCMC-BRS (10)   & 0.482 & 0.768 &  & 21 & 19.6 \\
+MCMC-BRS (30)   & 0.480 & 0.761 &  & 62 & 20.2 \\
+MCMC-BRS (100)  & 0.476 & 0.764 &  & 202 & 20.4  \\
 \hline
-MCMC-BSS (1)	& 0.473 & 0.758 & 14 & 44.3 \\
-MCMC-BSS (3)	& 0.461 & 0.734 & 29 & 69.7 \\
-MCMC-BSS (10)	& 0.451 & 0.720 & 82 & 89.8 \\
-MCMC-BSS (30)	& 0.444 & 0.713 & 232 & 96.7 \\
-MCMC-BSS (100)& 0.443 & 0.709 & 755 & 98.6 \\
+MCMC-BSS (1)	& 0.473 & 0.758 &  & 14 & 44.3 \\
+MCMC-BSS (3)	& 0.461 & 0.734 &  & 29 & 69.7 \\
+MCMC-BSS (10)	& 0.451 & 0.720 &  & 82 & 89.8 \\
+MCMC-BSS (30)	& 0.444 & 0.713 &  & 232 & 96.7 \\
+MCMC-BSS (100)  & 0.443 & 0.709 &  & 755 & 98.6 \\
 \hline
 \end{tabular}
+}
 \end{table}
@@
 \label{tab:case2_performance}
+{\scriptsize
 \centering
-\begin{tabular}{|c||c|c|c|c|}
+\begin{tabular}{|c||c|c|c|c|c|}
 \hline
-Algorithm & \begin{minipage}[c]{1cm} Position RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{1cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{1cm}  Unique Particles \end{minipage} \\
+Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  MNEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
 \hline
-FS 						& 6.17 & 2.03 & 0 & 3.96 \\
-DBRS					& 5.79 & 1.89 & 63 & 7.34 \\
+FS              & 6.17 & 2.03 &  & 0 & 3.96 \\
+DBRS            & 5.79 & 1.89 &  & 63 & 7.34 \\
 \hline
-MCMC-BRS (1)	& 6.13 & 2.01 & 2 & 4.77 \\
-MCMC-BRS (3)	& 6.07 & 1.99 & 4 & 5.29 \\
-MCMC-BRS (10)	& 5.94 & 1.94 & 10 & 6.1 \\
-MCMC-BRS (30)	& 5.82 & 1.91 & 29 & 6.73 \\
-MCMC-BRS (100)& 5.80 & 1.89 & 95 & 7.38 \\
+MCMC-BRS (1)	& 6.13 & 2.01 &  & 2 & 4.77 \\
+MCMC-BRS (3)	& 6.07 & 1.99 &  & 4 & 5.29 \\
+MCMC-BRS (10)	& 5.94 & 1.94 &  & 10 & 6.1 \\
+MCMC-BRS (30)	& 5.82 & 1.91 &  & 29 & 6.73 \\
+MCMC-BRS (100)  & 5.80 & 1.89 &  & 95 & 7.38 \\
 \hline
-MCMC-BSS (1)	& 5.90 & 1.91 & 9 & 12.6 \\
-MCMC-BSS (3)	& 5.67 & 1.81 & 18 & 22.7 \\
-MCMC-BSS (10)	& 5.48 & 1.73 & 50 & 44.1 \\
-MCMC-BSS (30)	& 5.43 & 1.69 & 144 & 70.9 \\
-MCMC-BSS (100)& 5.34 & 1.68 & 471 & 91.8 \\
+MCMC-BSS (1)	& 5.90 & 1.91 &  & 9 & 12.6 \\
+MCMC-BSS (3)	& 5.67 & 1.81 &  & 18 & 22.7 \\
+MCMC-BSS (10)	& 5.48 & 1.73 &  & 50 & 44.1 \\
+MCMC-BSS (30)	& 5.43 & 1.69 &  & 144 & 70.9 \\
+MCMC-BSS (100)  & 5.34 & 1.68 &  & 471 & 91.8 \\
 \hline
 \end{tabular}
+}
 \end{table}

### 2012-06-05 MATLAB/NEES.m

@@ -1,0 +1,28 @@
+function [ nees, nees_over_time ] = NEES( x, pts )
+%NEES Calculate the normalised-estimation-error-squared between the mean
+% particle estimate and the true values
+
+[d, K] = size(x);
+
+Np = size(pts, 1);
+
+pts_x = cell2mat(permute(pts, [3,2,1]));
+est_x = mean(pts_x, 3);
+est_P = zeros(d,d,K);
+error = zeros(1,K);
+for kk = 1:K
+%     est_P(:,:,kk) = cov(squeeze(pts_x(:,kk,:))');
+    innov = bsxfun(@minus, squeeze(pts_x(:,kk,:)), x(:,kk));
+    est_P(:,:,kk) = (1/Np)*innov*(innov');
+    if all(all(~diff(innov')))
+        error(kk) = 1;
+    else
+        error(kk) = ((est_x(:,kk)-x(:,kk))'/est_P(:,:,kk))*(est_x(:,kk)-x(:,kk));
+    end
+end
+
+nees = sum(error)/K;
+nees_over_time = error;
+
+end
+

### 2012-06-05 MATLAB/backward_rejectionsampling_smoother.m

@@ -52,7 +52,7 @@ for ii = 1:S
             
         end
         
-        count
+%         count
         
         % Update trajectory
         smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];

### 2012-06-05 MATLAB/backward_sampling_smoother.m

@@ -1,4 +1,4 @@
-function [ smooth_pts ] = backard_sampling_smoother( S, times, pts_array, wts_array, h_trans )
+function [ smooth_pts ] = backward_sampling_smoother( S, times, pts_array, wts_array, h_trans )
 %BACKARD_SAMPLING_SMOOTHER Normal Godsill2004-type backward sampling
 %smoother.
 %

### 2012-06-05 MATLAB/batch_smoother_testing2.m

@@ -3,19 +3,19 @@ dbstop if error
 
 % Get test flag
 % test_flag = str2double(getenv('SGE_TASK_ID'));
-test_flag = 1;
+test_flag = 3;
 
 % Batch test parameters
 num_seed = 10;
-set_batch_parameters;
+set_parameters;
 
 % Set test-specific parameters
 if test_flag == 1
     params.bng_var = (pi/720)^2;
     params.rng_var = 0.1;
 elseif test_flag == 2
-    params.bng_var = (pi/36)^2;
-    params.rng_var = 0.1;
+    params.bng_var = (pi/180)^2;
+    params.rng_var = 10;
 elseif test_flag == 3
     params.bng_var = (pi/36)^2;
     params.rng_var = 100;
@@ -29,26 +29,31 @@ times = cell(NA,1);
 unique_pts = cell(NA,1);
 mean_pos_rmse = cell(NA,1);
 mean_vel_rmse = cell(NA,1);
-mean_nees = cell(NA,1);
 for alg = 1:NA
     times{alg} = zeros(num_seed,1);
     unique_pts{alg} = zeros(num_seed,params.N);
     mean_pos_rmse{alg} = zeros(num_seed,1);
     mean_vel_rmse{alg} = zeros(num_seed,1);
-    mean_nees{alg} = zeros(num_seed,1);
 end
 
-% Loop through random seed
+% Loop through random seend
 for rs = 1:num_seed;
     
     rs
     
-    % Set random seed
-    s = RandStream('mt19937ar', 'seed', rs);
+    % Set random seed so we always generate the same data
+    s = RandStream('mt19937ar', 'seed', 1);
     RandStream.setDefaultStream(s);
+    
+    % Set parameters
+    set_parameters;
     
     %% Generate some Bearings only tracking data
     [ t, x, y ] = generate_radar_data;
+    
+    % Set random seed for running algorithms
+    s = RandStream('mt19937ar', 'seed', rs);
+    RandStream.setDefaultStream(s);
     
     %% Run a PF
     init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);
@@ -63,7 +68,6 @@ for rs = 1:num_seed;
     
     kiti_rmse = RMSE(x, kiti_pts);
     [kiti_Nup, kiti_Nuh] = count_unique_particles(kiti_pts);
-    [kiti_nees, kiti_nees_over_time] = NEES(x, kiti_pts);
     
     %Store
[…]

### 2012-06-05 MATLAB/plot_graphs_for_paper.m

@@ -1,7 +1,7 @@
 clup
 
 % Set test flag
-for test_flag = [1,3];
+for test_flag = [1,2,3];
     
     test_case = test_flag;
     if test_flag == 3
@@ -11,6 +11,7 @@ for test_flag = [1,3];
     % Load data
     load(['smoother_test' num2str(test_flag) '.mat']);
     
+    % RMSEs
     fig = figure; hold on
     inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'ok', 'markersize', 7 )
     inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'r^', 'markersize', 7 )
@@ -25,9 +26,24 @@ for test_flag = [1,3];
     plot([0 0], ylimits, ':k');
     ylim(ylimits);
     
-    wid = 4; hei = 3;
-    format_graph_for_pdf;
-    print(fig, '-dpdf', ['case' num2str(test_case) '_smoother_comparison_posRMSE_time.pdf']);
+    export_pdf(fig, ['case' num2str(test_case) '_smoother_comparison_posRMSE_time.pdf']);
+    
+    % MNEESs
+    fig = figure; hold on
+    inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'ok', 'markersize', 7 )
+    inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'r^', 'markersize', 7 )
+    inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), '*-b', 'markersize', 7 )
+    inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'x-.g', 'markersize', 7 )
+    
+    legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
+    xlabel('Running Time (s)'); ylabel('MNEES');
+    xlimits = get(gca, 'XLim');
+    ylimits = get(gca, 'YLim');
+    xlim([-50, xlimits(2)]);
+    plot([0 0], ylimits, ':k');
+    ylim(ylimits);
+    
[…]

### 2012-06-05 MATLAB/run_mcmcsmoother_demo.m

@@ -2,11 +2,14 @@ clup
 dbstop if error
 
 % DEFINE RANDOM SEED
-rand_seed = 1;
+rand_seed = 8;
 
 % Set random seed
 s = RandStream('mt19937ar', 'seed', rand_seed);
 RandStream.setDefaultStream(s);
+
+% Rejection sampling? Set to false - its too slow
+rej_samp = false;
 
 % Set parameters
 set_parameters;
@@ -52,14 +55,16 @@ plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
 cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), bs_smooth_pts)
 
 %% Run a rejection-sampling smoother
-tic;
-brs_smooth_pts = backward_rejectionsampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
-brs_time = toc;
-
-% Plot it
-figure(6); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
-plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
-cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), brs_smooth_pts)
+if rej_samp
+    tic;
+    brs_smooth_pts = backward_rejectionsampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
+    brs_time = toc;
+    
+    % Plot it
+    figure(6); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+    plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+    cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), brs_smooth_pts)
+end
 
 %% Run the MCMC smoother
 tic;
@@ -84,39 +89,59 @@ cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), mcmc_newstate_smoo
 %% Analysis
 
 filt_rmse = RMSE(x, filter_pts);
+[filt_nees, filt_nees_over_time] = NEES(x, filter_pts);
 
 kiti_rmse = RMSE(x, kiti_pts);
+[kiti_nees, kiti_nees_over_time] = NEES(x, kiti_pts);
[…]

### 2012-06-05 MATLAB/set_batch_parameters.m

@@ -4,9 +4,9 @@ global params;
 params.dt = 1;
 params.N = 100;
 params.proc_var = 1;
-params.bng_var = (pi/36)^2;    % Try (pi/720)^2; (pi/180)^2;
-params.rng_var = 0.1;           % Try 0.1;        10;
 params.x0 = [-100; 50; 10; 0];
+% params.bng_var    Set for each test
+% params.rng_var    Set for each test
 
 % Prior
 params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
@@ -14,4 +14,4 @@ params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
 % Algorithm
 params.Np = 100;
 params.S = 100;
-params.M = 10;                   % Try 5;           50;
+% params.M          Set of each test

### 2012-06-06 Paper/complete/revisions_MCMCsmoother.tex
Paper revisions. Code to rerun.

@@
 \usepackage{url}
+\usepackage{color}
+
+\newenvironment{meta}[0]{\color{red} \em}{}
 
@@
 In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation.
-In this paper, a Markov chain Monte Carlo sampling procedure is used to improve the efficiency of the particle smoother, achieving comparable error performance but with a lower execution time.
+In this paper, a Metropolis-Hastings sampling procedure is used to improve the efficiency of the particle smoother, achieving comparable error performance but with a lower execution time.
 In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error.
@@
 Methods for generating these samples have been described in \cite{Kitagawa1996,Godsill2004,Briers2010}.
-In this paper new algorithms are presented for drawing samples from the joint smoothing distribution, using Markov chain Monte Carlo (MCMC).
+In this paper new algorithms are presented for drawing samples from the joint smoothing distribution, using Metropolis-Hastings (MH).
 These novel procedures have computational advantages and greater flexibility over previous methods.
@@
 In section~\ref{sec:basics}, we briefly review the basic particle filter and existing joint smoother algorithms.
-The new MCMC-based smoothing methods are presented in sections~\ref{sec:mcmc_smoother} and~\ref{sec:new_state_smoother} with supporting simulations in section~\ref{sec:simulations}.
+The new MH-based smoothing methods are presented in sections~\ref{sec:mcmc_smoother} and~\ref{sec:new_state_smoother} with supporting simulations in section~\ref{sec:simulations}.
 
@@
 
-An improved particle smoothing algorithm was presented in \cite{Godsill2004}.
-This works by constructing new particles from the smoothing distribution by resampling from the filtering approximations.
+An improved particle smoother was presented in \cite{Godsill2004}, the forward-filtering-backward-sampling (FFBS) algorithm.
+This works by constructing new particles from the smoothing distribution by sampling from the filtering approximations.
 The sampling procedure exploits the following factorisation of the joint distribution,
@@
 
-The procedure may be repeated to produce as many samples from the joint smoothing distribution as required. This backward-resampling smoother is summarised in Fig.~\ref{alg:DBRS}.
+The procedure may be repeated to produce as many samples from the joint smoothing distribution as required.
+In this paper, we refer to this algorithm as the direct FFBS (D-FFBS) algorithm, in order to distinguish it from the MH variant developed in the next section.
+The D-FFBS algorithm is summarised in Fig.~\ref{alg:DBRS}.
 
@@
 }}
-\caption{Direct backward-resampling smoother algorithm}
+\caption{Direct forward-filtering-backward-sampling algorithm}
 \label{alg:DBRS}
@@
 The principal topic of this paper is a two-filter smoother for estimation of the marginal smoothing distributions, $p(x_{k}|y_{1:K})$, by combination of the results from one forward and one backward particle filter.
-Particles from the joint distribution may then be generated by resampling from either the forward or backward filters (or some combination) in the manner of the backward-resampling smoother outlined above.
+Particles from the joint distribution may then be generated by resampling from either the forward or backward filters (or some combination) in the manner of the D-FFBS algorithm outlined above.
 
-The complexity of both the backward-resampling smoother of \cite{Godsill2004} and the two-filter variant of \cite{Briers2010} is $\mathcal{O}(N_F \times N_S \times K)$, where $N_F$ is the number of filter particles, $N_S$ the number of smoother particles, and $K$ the number of time steps.
-In \cite{Douc2009}, a theoretically linear-cost scheme using rejection sampling is outlined, but in practice this often appears to be slower than the standard algorithms due to high rejection rates.
+The complexity of both the D-FFBS smoother of \cite{Godsill2004} and the two-filter variant of \cite{Briers2010} is $\mathcal{O}(N_F \times N_S \times K)$, where $N_F$ is the number of filter particles, $N_S$ the number of smoother particles, and $K$ the number of time steps.
+In \cite{Douc2009}, a theoretically linear-cost scheme using rejection sampling is outlined.
+In practice this often appears to be slower than the standard algorithms when there are more than a few state dimensions, due to high rejection rates.
 Methods for reducing the computational complexity of smoothing algorithms are also discussed in \cite{Klaas2006}.
-However, these focus on approximating a sum of kernel values over a set of particles, such as those which arise in forward-backward smoothing \cite{Doucet2000a} and two-filter smoothing \cite{Briers2010} for estimation of the marginal smoothing distribution.
+However, these focus on approximating a sum of kernel values over a set of particles, such as those which arise in forward-backward-smoothing \cite{Doucet2000a} and two-filter smoothing \cite{Briers2010} for estimation of the marginal smoothing distributions.
 They are not applicable for the joint smoothing procedures.
 
-In the algorithms described so far, no new states are sampled during the smoothing stage; there is only a resampling of the filtering distributions.
-This limits the support of the smoother to the states that appear in the filtering particles.
+In the algorithms described so far, no new states are sampled during the backwards smoothing pass; state values are only ``recycled'' from the particle filtering approximations.
+This limits the support of the smoother to the states that appear in these filtering particles.
 In \cite{Fearnhead2010}, an IS-based algorithm for estimation of the marginal smoothing distributions is described in which new states are sampled from a proposal density, thus recovering the full support.
@@
 
-\section{New MCMC Particle Smoother} \label{sec:mcmc_smoother}
+\section{An MCMC Particle Smoother} \label{sec:mcmc_smoother}
 
-The bottleneck of the backward-resampling smoother is the calculation of the sampling weights, specified by (\ref{eq:DBRS_weights}).
+The bottleneck of the D-FFBS algorithm is the calculation of the sampling weights, specified by (\ref{eq:DBRS_weights}).
 For each smoother particle and for each time step, a weight must be calculated for each of the $N_F$ filter particles, in order to draw a sample from the conditional distribution $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
-The innovation here is to use Markov chain Monte Carlo (MCMC) to produce these samples instead of sampling directly.
+The innovation here is to use a Markov chain Monte Carlo (MCMC) procedure to produce these particles instead of sampling directly.
 
@@
 
-For the direct backward-resampling smoother, the objective at each time step was to draw a sample from $p(x_{k}|\tilde{x}_{k+1}, y_{1:K})$.
-To simplify the notation and initialisation of chains, the MCMC version will target $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
+For the D-FFBS, the objective at each time step was to draw a sample from $p(x_{k}|\tilde{x}_{k+1}, y_{1:K})$.
+To simplify the notation and initialisation of chains, the MH version will target $p(x_{1:k}|\tilde{x}_{k+1}, y_{1:K})$.
 Theoretically this means using the particles of the filter-smoother, $\{x_{1:k}^{(i)}\}$, rather than just the filter, $\{x_k^{(j)}\}$, which would require $\mathcal{O}(K^2)$ memory.
@@
 The final state of each chain is used as a sample from the required conditional distribution.
-The complete MCMC backward-resampling smoother is summarised in Fig.~\ref{alg:MCMC-BRS}.
+The resulting MH-FFBS algorithm is summarised in Fig.~\ref{alg:MCMC-BRS}.
 
@@
 }}
-\caption{MCMC backward-resampling smoother algorithm}
+\caption{Metropolis-Hastings forward-filtering-backward-sampling algorithm}
 \label{alg:MCMC-BRS}
@@
 
-The computational complexity of the MCMC-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used for each time step and smoothing particle.
+The computational complexity of the MH-based algorithm is $\mathcal{O}(M \times N_S \times K)$, where $M$ is the (mean) number of MH steps used for each time step and smoothing particle.
 The advantage of the new procedure is that $M$ can often be much smaller than $N_F$ and yet achieve almost equal error performance.
 
-With the direct backward-resampling smoother, the only parameter to be chosen is $N_S$, the number of smoother particles.
-With the MCMC variant, we also have the freedom to set $M$, the number of MH steps in each Markov chain.
+With the D-FFBS, the only parameter to be chosen is $N_S$, the number of smoother particles.
+With the MH variant, we also have the freedom to set $M$, the number of MH steps in each Markov chain.
 This controls a trade-off between speed and accuracy/particle diversity.
@@
 
-The new scheme targets the same empirical approximation to the smoothing distribution as used in the direct backward-resampling method.
-However, rather than drawing independent samples from this approximation, MCMC is used, resulting in dependent samples.
+The new scheme targets the same empirical approximation to the smoothing distribution as used in the D-FFBS method.
+However, rather than drawing independent samples from this approximation, MH is used, resulting in dependent samples.
 In the limit as the number of MH steps, $M$, tends to infinity, the Markov chain will produce an independent sample from the target distribution, as the initial state of the chain will have been forgotten.
-At this point, the performance of the MCMC method is expected to match that of the direct method in terms of error and particle diversity.
+At this point, the performance of the MH method is expected to match that of the direct method in terms of error and particle diversity.
 At the other extreme, with $M=0$, the algorithm simply outputs particles from the filter-smoother approximation, which is expected to give worse performance.
@@
 
-This new scheme bears a resemblance to the ``linear cost'' simulation scheme of \cite{Douc2009} which uses rejection sampling instead of MCMC.
-Although theoretically elegant, the latter algorithm has been found to suffer from such high rejection rates as to render it consistently slower than the direct sampling implementation on a practical problem (see section~\ref{sec:simulations}).
+This new scheme bears a resemblance to the ``linear cost'' simulation scheme of \cite{Douc2009} which uses rejection sampling instead of MH.
+Although theoretically elegant, the latter algorithm has been found to suffer from such high rejection rates as to render it consistently slower than the direct sampling implementation on problems with more than 1 state dimension (see section~\ref{sec:simulations}).
 
@@
 
-Smoothers based on backward-resampling, both the direct sampler of \cite{Godsill2004} and the MCMC version outlined in the previous section, are limited by the fact that states are only selected from those which appear in the filtering approximation.
+The FFBS smoothers, both the direct sampling version of \cite{Godsill2004} and the MH version outlined in the previous section, are limited by the fact that states are only selected from those which appear in the filtering approximation.
 If there is a significant difference between the filtering and smoothing densities, then the particles are unlikely to be in the right locations to represent the smoothing distribution well, and the resulting approximation will be poor.
@@
 
-The proposed new method proceeds in a similar manner to the MCMC backward-resampling algorithm.
+The proposed new method proceeds in a similar manner to the MH-FFBS algorithm.
 However, this time only $x_{1:k-1}$ is resampled from the particle filter-smoother approximation from time step $k-1$.
@@
 
-The complete MCMC backward-sampling smoother is summarised in Fig.~\ref{alg:MCMC-BSS}.
-The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MCMC resampling smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
-In contrast with the MCMC resampling method, this smoother targets the exact posterior distribution for the current state at time $t_n$, rather than a particle approximation.
-Consequently, the performance is expected to be improved relative to the resampling-based methods in terms of both error and particle diversity.
+We refer to this algorithm as the Metropolis-Hastings forward-filtering-backward-proposing (MH-FFBP) smoother, emphasising the fact that samples are proposed afresh rather than recycling those from the filtering approximations.
+The complete algorithm is summarised in Fig.~\ref{alg:MCMC-BSS}.
+The computational complexity of this algorithm is $\mathcal{O}(M \times N_S \times K)$, the same as the MH-FFBS smoother, although it is likely take more time to run because of the additional sampling operation and probability calculations.
+In contrast with the MH-FFBS algorithm, this smoother targets the exact posterior distribution for the current state at time $t_n$, rather than a particle approximation.
+Consequently, the performance is expected to be improved relative to the FFBS methods in terms of both error and particle diversity.
 
-We note that this new scheme bears a resemblance to the improvement strategy for approximations to the joint smoothing distribution published recently in \cite{Dubarry2011}.
-Their algorithm iteratively applies MH steps to a low-diversity particle approximation, such as that from the filter-smoother.
-Our algorithm differs in the fact that it allows simultaneous changes to the particle state history, $x_{1:n-1}$, as well as the current state.
-Furthermore, for each smoothing particle, we run a Markov chain at each time step, using one sweep backwards through time, rather than multiple sweeps with a single MH step at each time point.
+We note that the MH-FFBP algorithm shares similar features with the procedure published recently in \cite{Dubarry2011}.
+The latter iteratively applies MH steps to a particle approximation of the joint smoothing algorithm in order to improve particle diversity.
+In this sense, it is akin to a resample-move \cite{Gilks2001} scheme for the entire smoothing distribution.
+In contrast, the MH-FFBP algorithm integrates MH steps into a single backward smoothing pass.
+It allows changes to the entire state history of a particle by proposing $x_{1:k-1}$ from the filtering approximation.
 
@@
 }}
-\caption{MCMC backward-sampling smoother algorithm}
+\caption{Metropolis-Hastings forward-filtering-backward-proposing algorithm}
 \label{alg:MCMC-BSS}
@@
 
-When using MCMC, if samples can be proposed directly from the target distribution then the acceptance probability is always 1 (In combination with a block-sampling scheme, this becomes the Gibbs sampler of \cite{geman1984}).
+When using MH, if samples can be proposed directly from the target distribution then the acceptance probability is always 1 (In combination with a block-sampling scheme, this becomes the Gibbs sampler of \cite{geman1984}).
 This is desirable as it maximises the acceptance rate.
-Clearly, for the basic MCMC backward-resampling smoother of section~\ref{sec:mcmc_smoother}, such a ``target density proposal'' simply yields the direct backward-resampling smoother of \cite{Godsill2004}, as described in section~\ref{sec:existing_smoothers}.
-For the MCMC backward-sampling smoother, a target density proposal can be made by sequentially sampling $x_{1:k-1}$ followed by $x_k$ using the following factorisation,
+Clearly, for the basic MH-FFBS smoother of section~\ref{sec:mcmc_smoother}, such a ``target density proposal'' simply yields the D-FFBS smoother of \cite{Godsill2004}, as described in section~\ref{sec:existing_smoothers}.
+For the MH-FFBP smoother, a target density proposal can be made by sequentially sampling $x_{1:k-1}$ followed by $x_k$ using the following factorisation,
 %
@@
 The weights for the $x_{1:k-1}$ proposal may be approximated using similar methods, but this will often be undesirable.
-If it is necessary to calculate proposal weights for all of the filter-smoother particles, then the computational advantages of using MCMC are lost.
+If it is necessary to calculate proposal weights for all of the filter-smoother particles, then the computational advantages of using MH are lost.
 With this in mind, an appropriate choice for the proposal weights might be simply $\tilde{v}_{k-1}^{(j)} = w_{k-1}^{(j)}$, which avoids calculation of $p(y_{k}, \tilde{x}_{k+1}| x_{1:k-1}^{(j)})$ for all $j$.
@@
 	\item The simple filter-smoother (FS) of \cite{Kitagawa1996}
-	\item The direct backward-resampling smoother (DBRS) of \cite{Godsill2004}
-	\item The new MCMC backward resampling smoother (MCMC-BRS) of section~\ref{sec:mcmc_smoother}, with several values of $M$
-	\item The new MCMC backward sampling smoother (MCMC-BSS) of section~\ref{sec:new_state_smoother}, with several values of $M$
+	\item The direct forward-filtering-backward-sampling smoother (D-FFBS) of \cite{Godsill2004}
+	\item The new Metropolis-Hastings forward-filtering-backward-sampling smoother (MH-FFBS) of section~\ref{sec:mcmc_smoother}, with several values of $M$
+	\item The new Metropolis Hastings forward-filtering-backward-proposing smoother (MH-FFBP) of section~\ref{sec:new_state_smoother}, with several values of $M$
 \end{itemize}
 
-The rejection-sampling version of the DBRS described by \cite{Douc2009} was also implemented, but high rejection rates meant that it was impractically slow so it was omitted from further testing.
+The rejection-sampling version of the FFBS described by \cite{Douc2009} was also implemented, but high rejection rates meant that it was impractically slow so it was omitted from further testing.
 
@@
 $v_k^{(j)} = w_k^{(j)}$.
-The particle filter uses the ``optimal importance density'' and the MCMC-BSS a ``target density proposal'' with linearisations of the observation model where appropriate.
+The particle filter uses the ``optimal importance density'' and the MH-FFBP a ``target density proposal'' with linearisations of the observation model where appropriate.
 
@@
 These point estimates are then used to calculate a root-mean-square error (RMSE).
-In addition, a modified normalised estimation error squared (MNEES) is calculated using,
+In addition, an empirical normalised estimation error squared (ENEES) is calculated using,
 %
 \begin{IEEEeqnarray}{rCl}
- \text{MNEES}_k & = & (\hat{\mathbf{x}}_k - \mathbf{x}_k^*)^T \hat{P}_k^{-1} (\hat{\mathbf{x}}_k - \mathbf{x}_k^*) \nonumber \\
+ \text{ENEES}_k & = & (\hat{\mathbf{x}}_k - \mathbf{x}_k^*)^T \hat{P}_k^{-1} (\hat{\mathbf{x}}_k - \mathbf{x}_k^*) \nonumber \\
  \hat{\mathbf{x}}_k & = & \frac{1}{N_S} \sum_i \mathbf{x}_k^{(i)} \nonumber \\
@@
 \noindent and where $\mathbf{x}_k^*$ is the true value of the state.
-This statistic has a value of $1$ when there is only one unique particle in the approximation, and tends to $0$ as the number of independent particles tends to infinity.
+This statistic has a value of $1$ when there fewer linearly independent state values in the particle approximation than state dimensions, and tends to $0$ as the number of independent particles tends to infinity.
 It provides a relative measure of the quality of different particle approximations, taking into account both estimation error and particle diversity.
@@
 \hline
-Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  MNEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
+Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
 \hline
 FS              & 0.578 & 0.968 &  & 0 & 2.25 \\
-DBRS            & 0.475 & 0.762 &  & 122 & 20.3 \\
+D-FFBS          & 0.475 & 0.762 &  & 122 & 20.3 \\
 \hline
-MCMC-BRS (1)    & 0.509 & 0.823 &  & 3 & 13.7  \\
-MCMC-BRS (3)    & 0.493 & 0.790 &  & 7 & 17.4  \\
-MCMC-BRS (10)   & 0.482 & 0.768 &  & 21 & 19.6 \\
-MCMC-BRS (30)   & 0.480 & 0.761 &  & 62 & 20.2 \\
-MCMC-BRS (100)  & 0.476 & 0.764 &  & 202 & 20.4  \\
+MH-FFBS (1)    & 0.509 & 0.823 &  & 3 & 13.7  \\
+MH-FFBS (3)    & 0.493 & 0.790 &  & 7 & 17.4  \\
+MH-FFBS (10)   & 0.482 & 0.768 &  & 21 & 19.6 \\
+MH-FFBS (30)   & 0.480 & 0.761 &  & 62 & 20.2 \\
+MH-FFBS (100)  & 0.476 & 0.764 &  & 202 & 20.4  \\
 \hline
-MCMC-BSS (1)	& 0.473 & 0.758 &  & 14 & 44.3 \\
-MCMC-BSS (3)	& 0.461 & 0.734 &  & 29 & 69.7 \\
-MCMC-BSS (10)	& 0.451 & 0.720 &  & 82 & 89.8 \\
-MCMC-BSS (30)	& 0.444 & 0.713 &  & 232 & 96.7 \\
-MCMC-BSS (100)  & 0.443 & 0.709 &  & 755 & 98.6 \\
+MH-FFBP (1)    & 0.473 & 0.758 &  & 14 & 44.3 \\
+MH-FFBP (3)    & 0.461 & 0.734 &  & 29 & 69.7 \\
+MH-FFBP (10)   & 0.451 & 0.720 &  & 82 & 89.8 \\
+MH-FFBP (30)   & 0.444 & 0.713 &  & 232 & 96.7 \\
+MH-FFBP (100)  & 0.443 & 0.709 &  & 755 & 98.6 \\
 \hline
@@
 \hline
-Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  MNEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
+Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
 \hline
 FS              & 6.17 & 2.03 &  & 0 & 3.96 \\
-DBRS            & 5.79 & 1.89 &  & 63 & 7.34 \\
+D-FFBS         & 5.79 & 1.89 &  & 63 & 7.34 \\
 \hline
-MCMC-BRS (1)	& 6.13 & 2.01 &  & 2 & 4.77 \\
-MCMC-BRS (3)	& 6.07 & 1.99 &  & 4 & 5.29 \\
-MCMC-BRS (10)	& 5.94 & 1.94 &  & 10 & 6.1 \\
-MCMC-BRS (30)	& 5.82 & 1.91 &  & 29 & 6.73 \\
-MCMC-BRS (100)  & 5.80 & 1.89 &  & 95 & 7.38 \\
+MH-FFBS (1)    & 6.13 & 2.01 &  & 2 & 4.77 \\
+MH-FFBS (3)    & 6.07 & 1.99 &  & 4 & 5.29 \\
+MH-FFBS (10)   & 5.94 & 1.94 &  & 10 & 6.1 \\
+MH-FFBS (30)   & 5.82 & 1.91 &  & 29 & 6.73 \\
+MH-FFBS (100)  & 5.80 & 1.89 &  & 95 & 7.38 \\
 \hline
-MCMC-BSS (1)	& 5.90 & 1.91 &  & 9 & 12.6 \\
-MCMC-BSS (3)	& 5.67 & 1.81 &  & 18 & 22.7 \\
-MCMC-BSS (10)	& 5.48 & 1.73 &  & 50 & 44.1 \\
-MCMC-BSS (30)	& 5.43 & 1.69 &  & 144 & 70.9 \\
-MCMC-BSS (100)  & 5.34 & 1.68 &  & 471 & 91.8 \\
+MH-FFBP (1)    & 5.90 & 1.91 &  & 9 & 12.6 \\
+MH-FFBP (3)    & 5.67 & 1.81 &  & 18 & 22.7 \\
+MH-FFBP (10)   & 5.48 & 1.73 &  & 50 & 44.1 \\
+MH-FFBP (30)   & 5.43 & 1.69 &  & 144 & 70.9 \\
+MH-FFBP (100)  & 5.34 & 1.68 &  & 471 & 91.8 \\
 \hline
@@
 
-The MCMC-BRS smoother error approaches that of the DBRS as the number of MCMC steps, $M$, increases.
-Similarly, the mean number of unique particles in the MCMC-BRS approximation approaches that of the DBRS.
-This behaviour is expected as a longer Markov chain is more likely to deliver an independent sample from the target distribution, whereas independent samples are guaranteed in the DBRS approximation due to the direct sampling procedure.
+{\meta Add comments on ENEES.}
 
-When the number of MCMC steps, $M$, is small, the error of the MCMC-BRS is only very slightly larger than that of the DBRS but with a significantly shorter running time.
+The MH-FFBS smoother error approaches that of the D-FFBS as the number of MH steps, $M$, increases.
+Similarly, the mean number of unique particles in the MH-FFBS approximation approaches that of the D-FFBS.
+This behaviour is expected as a longer Markov chain is more likely to deliver an independent sample from the target distribution, whereas independent samples are guaranteed in the D-FFBS approximation due to the direct sampling procedure.
+
+When the number of MH steps, $M$, is small, the error of the MH-FFBS is only very slightly larger than that of the D-FFBS but with a significantly shorter running time.
 For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed of computation.
-Thus, the MCMC-BRS provides increased flexibility when compared to the DBRS scheme.
+Thus, the MH-FFBS provides increased flexibility when compared to the D-FFBS scheme.
 
-The MCMC-BSS outperforms all of the other smoothers, achieving the lowest RMSE for any given running time.
+The MH-FFBP outperforms all of the other smoothers, achieving the lowest RMSE for any given running time.
 Furthermore, because the support is not limited to using states from the filter distributions, the smoothing particle approximation has many more unique particles, giving a less degenerate representation.
@@
 \section{Conclusion} \label{sec:conclusions}
-Two new algorithms have been presented for drawing particles from the joint smoothing distribution using MCMC.
+Two new algorithms have been presented for drawing particles from the joint smoothing distribution using Metropolis-Hastings.
 These MCMC schemes have a degree of flexibility in the choice of chain length, which controls a trade-off between accuracy and running time.
 
-The first algorithm resamples particles from the filtering distributions to produce particles from the smoothing distribution.
-The error performance of this scheme approaches that of the standard resampling scheme of \cite{Godsill2004} as the chain length increases.
+The first algorithm samples particles from the filtering approximations to produce particles from the smoothing distribution.
+The error performance of this scheme approaches that of the standard forward-filtering-backward-sampling of \cite{Godsill2004} as the chain length increases.
 However, comparable error performance may be achieved with very short chains, leading to a significant saving in processing time.
@@
 The second algorithm samples new values of the state from a continuous proposal density.
-Although slower than the resampling version for a given length of Markov chain, the error performance is reduced relative to the other smoothers.
+Although slower than the previous scheme for a given length of Markov chain, the error performance is reduced relative to the other smoothers.
 

### 2012-06-06 MATLAB/NEES.m

@@ -14,7 +14,7 @@ for kk = 1:K
 %     est_P(:,:,kk) = cov(squeeze(pts_x(:,kk,:))');
     innov = bsxfun(@minus, squeeze(pts_x(:,kk,:)), x(:,kk));
     est_P(:,:,kk) = (1/Np)*innov*(innov');
-    if all(all(~diff(innov')))
+    if det(est_P(:,:,kk))<1E-10  %all(all(~diff(innov')))  %
         error(kk) = 1;
     else
         error(kk) = ((est_x(:,kk)-x(:,kk))'/est_P(:,:,kk))*(est_x(:,kk)-x(:,kk));
@@ -23,6 +23,8 @@ end
 
 nees = sum(error)/K;
 nees_over_time = error;
+
+assert(~isnan(nees));
 
 end
 

### 2012-06-06 MATLAB/batch_smoother_testing.m

@@ -3,10 +3,10 @@ dbstop if error
 
 % Get test flag
 % test_flag = str2double(getenv('SGE_TASK_ID'));
-test_flag = 1;
+test_flag = 3;
 
 % Batch test parameters
-num_seed = 10;
+num_seed = 100;
 set_batch_parameters;
 
 % Set test-specific parameters

### 2012-06-06 MATLAB/run_mcmcsmoother_demo.m

@@ -2,7 +2,7 @@ clup
 dbstop if error
 
 % DEFINE RANDOM SEED
-rand_seed = 8;
+rand_seed = 10;
 
 % Set random seed
 s = RandStream('mt19937ar', 'seed', rand_seed);

### 2012-06-08 Paper/complete/response_to_reviewers.tex
submitted revisions

\documentclass{article}

\usepackage{amsmath}
\usepackage{IEEEtrantools}
\usepackage{color}

\newenvironment{meta}[0]{\color{red} \em}{}

\begin{document}

{\meta Since submitting this paper, we've discovered a related piece of work, ``Particle approximation improvement of the
joint smoothing distribution with on-the-fly variance estimation'' by Dubarry and Douc. This is available as an arXiv preprint (arXiv:1107.5524) http://arxiv.org/abs/1107.5524 . The algorithm it contains shares some characteristics with our backwards proposing smoother. We've added a reference in section IV, along with a brief comparison.  }

Reviewer Comments:

Reviewer: 1

Recommendation: AQ - Publish In Minor, Required Changes

Comments:
An improved particle filtering approach using MCMC is proposed in this paper. The work is novel and the paper is pleasant to read. Some minor points are listed here for consideration.

1) Make sure that paragraphs are properly used. Quite a lot of paragraphs can be merged, e.g., line 42 on page 2.
{\meta I've merged a few, but I like to have a separate paragraph for each point I make. I consider this a matter of taste. }

2) Check all punctuation symbols.
{\meta Done. Yes. Whoops.}

3) Algorithm initialization is not clear in simulations.
{\meta Added a sentence about this.}

4) Reorganize section IV, if necessary.
{\meta I think it is fine.}

Additional Questions:
1. Is the topic appropriate for publication in these transactions?: Yes

2. Is the topic important to colleagues working in the field?: Yes

Explain:

1. Is the paper technically sound?: Yes

why not?:

2. Is the coverage of the topic sufficiently comprehensive and balanced?: Yes

3. How would you describe technical depth of paper?: Appropriate for the Generally Knowledgeable Individual Working in the Field or a Related Field

4. How would you rate the technical novelty of the paper?: Novel

1. How would you rate the overall organization of the paper?: Could be improved

2. Are the title and abstract satisfactory?: Yes

Explain:

3. Is the length of the paper appropriate? If not, recommend how the length of the paper should be amended, including a possible target length for the final manuscript.: Yes

4. Are symbols, terms, and concepts adequately defined?: Yes

5. How do you rate the English usage? : Needs improvement

6. Rate the Bibliography: Satisfactory

null:

1. How would you rate the technical contents of the paper?: Good

2. How would you rate the novelty of the paper?: Sufficiently Novel

3. How would you rate the "literary" presentation of the paper?: Totally Accessible

4. How would you rate the appropriateness of this paper for publication in this IEEE Transactions?: Excellent Match

Reviewer: 2

Recommendation: RQ - Review Again After Major Changes

Comments:
This has the makings of a really nice paper. However, I am bothered by a few things and I think addressing these could transform the paper from being slightly interesiting (as it is) to really great (as it could be).

1) I'm bothered that the authors seem to have a very different view of [13] to me. Certainly, the comments about [13] are probably based on the authors' experience, in which case the text should be softened a little to say so. It would be even better if the techniques in [13] were included in the analysis: the point of [13] (as I understand it) was to change the scaling of the computational complexity, which is clearly a key part of the approach described in the current submission. Exlcuding such a comparison only devalues the submission.

{\meta An interesting point. Our glib dismissal of the this reference is definitely inappropriate given the subject matter, and the reasoning was wrong. However, if I understand correctly, the methods in this paper revolve around fast approximation of summation of function values over a particle population. They are, therefore, tailored towards algorithms for marginal smoothing. For the joint smoothing task considered in our paper, we need to sample from these particle populations, and this is the basis on which we do not consider these fast approximations further. The text has been modified to this effect. }

2) The RMSE results are not very convincing: the differences are small. Indeed, the results could be more convincing: specifically, using NEES (alongside RMSE) would be a good way to downplay the small differences in RMSE and upplay the differences in the number of unique particles in terms of a performance metric that "matters".

{\meta Yes. Very true. A proper NEES score would require us to know the true posterior means and covariances. Since we can't calculate these for a nonlinear model, I've used an empirical form - details in the paper. }

3) The values used in the two cases for the bearing and range accuracy aren't actually that different in their nature. Indeed the cases differ in terms of the absolute value of the noise, but not (much) in its correlation structure. Changing the absolute value of the noise alters the implicit time constant associated with the posterior. I accept that this is interesting since it stresses the ability of the smoother to explore the space. However, the interesting bit about the bearing and range likelihood is when the likleihood function becomes heavily skewed (ie "banana" shaped). Adding a case with a different degree of skewness (eg using $\sigma_B^2 = (\pi/36)^2$ and $\sigma_R^2 = 0.1$) would be really interesting and would strengthen the paper.

{\meta Ok. Done. }

I also spotted a few bits of niff-naff-and-trivia that the authors could sensibly (and easily) address:
1) "using the a Rauch-Tung-Striebel (RTS) smoother" -> "using the Rauch-Tung-Striebel (RTS) smoother" (typo)
2) The first line of (7) is slightly non-standard and results in the approximation in the final line. It would be better if a full path was considered such that equalities alone were used.

{\meta The approximation is a typo. It was actually intentional not to use the full path, to emphasise the difference between the filter and the filter-smoother. By marginalising $x_{n-1}$ after each step, we get the same filter output as if we used the full path. }

3)  "backwards-resampling" is a really misleading name given the use of "resampling" in the (forwards) particle filter. Why not call it "backwards-resimulation"? I realise that this would be a new term, but it might be helpful to distinguish "resampling" and "resampling"!!

{\meta Yes. I was worried about this myself. I've renamed all the algorithms. }

Additional Questions:
1. Is the topic appropriate for publication in these transactions?: Yes

2. Is the topic important to colleagues working in the field?: Yes

Explain:

1. Is the paper technically sound?: Yes

why not?:

2. Is the coverage of the topic sufficiently comprehensive and balanced?: Important Information is missing or superficially treated.

3. How would you describe technical depth of paper?: Appropriate for the Generally Knowledgeable Individual Working in the Field or a Related Field

4. How would you rate the technical novelty of the paper?: Novel

1. How would you rate the overall organization of the paper?: Satisfactory

2. Are the title and abstract satisfactory?: Yes

Explain:

3. Is the length of the paper appropriate? If not, recommend how the length of the paper should be amended, including a possible target length for the final manuscript.: Yes

4. Are symbols, terms, and concepts adequately defined?: Yes

5. How do you rate the English usage? : Satisfactory

6. Rate the Bibliography: Satisfactory

null:

1. How would you rate the technical contents of the paper?: Good

2. How would you rate the novelty of the paper?: Sufficiently Novel

3. How would you rate the "literary" presentation of the paper?: Totally Accessible

4. How would you rate the appropriateness of this paper for publication in this IEEE Transactions?: Excellent Match

\end{document}

### 2012-06-08 Paper/complete/revisions_single_MCMCsmoother.tex
submitted revisions

--- 2012-06-06 Paper/complete/revisions_MCMCsmoother.tex
@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{}
 
@@
 
-\begin{IEEEkeywords}
-state space model, particle filter, smoothing, MCMC, Bayesian inference
-\end{IEEEkeywords}
+%\begin{IEEEkeywords}
+%state space model, particle filter, smoothing, MCMC, Bayesian inference
+%\end{IEEEkeywords}
+
+\ifCLASSOPTIONpeerreview
+\begin{center} \bfseries EDICS Categories: MLR-BAYL, SSP-NGAU, SSP-TRAC, SSP-FILT \end{center}
+\fi
+
+\newpage
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+%\IEEEpeerreviewmaketitle
 
@@
 The proposal density for $x_{k-1}$, $q(x_{k-1})$, is constructed from the particles of the filtering approximation from the previous time step.
-The choice of weights and sampling procedure for $x_{k-1}$ determines what from of ``resampling'' is to be used.
+The choice of weights and sampling procedure for $x_{k-1}$ determine what form of ``resampling'' is to be used.
 Here we use arbitrary weights for generality,
@@
 This, however, leads to degeneracy of the particle weights over time.
-(See \cite{Douc2005,Cappe2007,Doucet2009} for further discussion of resampling.)
+(See \cite{Cappe2007,Doucet2009} for further discussion of resampling.)
 
@@
 
-Following an equivalent derivation to that for (\ref{eq:backward_conditional_filter}), the target distribution may be approximated by a particle distribution,
+Following an equivalent derivation to that for (\ref{eq:backward_conditional_filter}), the target distribution may be approximated by a set of particles,
 %
@@
 This new scheme bears a resemblance to the ``linear cost'' simulation scheme of \cite{Douc2009} which uses rejection sampling instead of MH.
-Although theoretically elegant, the latter algorithm has been found to suffer from such high rejection rates as to render it consistently slower than the direct sampling implementation on problems with more than 1 state dimension (see section~\ref{sec:simulations}).
+Although theoretically elegant, the latter algorithm has been found to suffer from such high rejection rates as to render it consistently slower than the direct sampling implementation on problems with more than one state dimension (see section~\ref{sec:simulations}).
 
@@
 
-10 realisations of 500 time steps were simulated from the model and the algorithms tested on each one.
-Results are shown here for high and low observation noise cases.
+100 realisations of 500 time steps were simulated from the model and the algorithms tested on each one.
+Results are shown here for three noise parameter sets.
 For case 1, the observation variances were set to $\sigma_B^2 = (\pi/720)^2$ and $\sigma_R^2 = 0.1$.
-For case 2, $\sigma_B = (\pi/36)^2$ and $\sigma_R^2 = 100$.
+For case 2, $\sigma_B^2 = (\pi/36)^2$ and $\sigma_R^2 = 0.1$, resulting in a highly skewed likelihood function.
+For case 3, $\sigma_B = (\pi/36)^2$ and $\sigma_R^2 = 100$.
 Other model settings are $\Delta t = 1$, $\sigma_P = 1$, and $x_0 = [-100, 50, 10, 0]^T$.
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_trajectory.pdf}%
-\label{fig_first_case}}
+\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_trajectory.pdf}%
-\label{fig_second_case}}
-\caption{Example trajectory with observations from the two noise covariance cases.}
+\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory.pdf}}
+\\
+\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_trajectory.pdf}}
+\caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
 \label{fig:example_trajectories}
@@
 \noindent and where $\mathbf{x}_k^*$ is the true value of the state.
-This statistic has a value of $1$ when there fewer linearly independent state values in the particle approximation than state dimensions, and tends to $0$ as the number of independent particles tends to infinity.
-It provides a relative measure of the quality of different particle approximations, taking into account both estimation error and particle diversity.
+This statistic has a value of $1$ when there are fewer linearly independent state values in the particle approximation than state dimensions.
+It decreases as the number of independent particles increases (improving the covariance estimate) or the estimation error decreases.
+Thus, it provides a simple relative measure of the quality of different particle approximations, taking into account both estimation error and particle diversity.
 Finally, the mean number of unique particles at each time instant and the running times (excluding the time for filtering) are also recorded.
-These performance indicators are all shown in tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} for the two noise scenarios.
-The trade-off between speed and error is shown in Fig.~\ref{fig:rmse_vs_time}.
-
-%Tables~\ref{tab:case1_performance} and~\ref{tab:case2_performance} show the root-mean-square errors (RMSEs) in position and velocity obtained by the different algorithms in the two noise scenarios, along with running times (excluding the time for filtering) and the mean number of unique smoother particles at each time instant.
+These performance indicators are all shown in tables~\ref{tab:case1_performance}--\ref{tab:case3_performance} for the two noise scenarios.
 The trade-off between speed and error is shown in Fig.~\ref{fig:rmse_vs_time}.
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Smoother performance measures for noise covariance case 1. All results are averaged over 10 runs, with 500 time steps in each run.}
+\caption{Smoother performance measures for noise covariance case 1. All results are averaged over 100 runs, with 500 time steps in each run.}
 \label{tab:case1_performance}
 {\scriptsize
-\centering
+\begin{center}
 \begin{tabular}{|c||c|c|c|c|c|}
@@
 \hline
-FS              & 0.578 & 0.968 &  & 0 & 2.25 \\
-D-FFBS          & 0.475 & 0.762 &  & 122 & 20.3 \\
+FS             & 0.56 & 0.96 & 0.99 & 0.00 & 2.13 \\
+D-FFBS         & 0.45 & 0.76 & 0.84 & 65.88 & 20.54 \\
 \hline
-MH-FFBS (1)    & 0.509 & 0.823 &  & 3 & 13.7  \\
-MH-FFBS (3)    & 0.493 & 0.790 &  & 7 & 17.4  \\
-MH-FFBS (10)   & 0.482 & 0.768 &  & 21 & 19.6 \\
-MH-FFBS (30)   & 0.480 & 0.761 &  & 62 & 20.2 \\
-MH-FFBS (100)  & 0.476 & 0.764 &  & 202 & 20.4  \\
+MH-FFBS (1)    & 0.48 & 0.81 & 0.90 & 1.80 & 13.95 \\
+MH-FFBS (3)    & 0.46 & 0.78 & 0.87 & 3.89 & 17.63 \\
+MH-FFBS (10)   & 0.46 & 0.77 & 0.85 & 11.20 & 19.83 \\
+MH-FFBS (30)   & 0.45 & 0.76 & 0.84 & 32.05 & 20.40 \\
+MH-FFBS (100)  & 0.45 & 0.76 & 0.84 & 105.12 & 20.55 \\
 \hline
-MH-FFBP (1)    & 0.473 & 0.758 &  & 14 & 44.3 \\
-MH-FFBP (3)    & 0.461 & 0.734 &  & 29 & 69.7 \\
-MH-FFBP (10)   & 0.451 & 0.720 &  & 82 & 89.8 \\
-MH-FFBP (30)   & 0.444 & 0.713 &  & 232 & 96.7 \\
-MH-FFBP (100)  & 0.443 & 0.709 &  & 755 & 98.6 \\
+MH-FFBP (1)    & 0.45 & 0.75 & 0.82 & 9.20 & 44.32 \\
+MH-FFBP (3)    & 0.43 & 0.73 & 0.79 & 19.23 & 70.12 \\
+MH-FFBP (10)   & 0.43 & 0.72 & 0.78 & 54.33 & 90.10 \\
+MH-FFBP (30)   & 0.42 & 0.71 & 0.79 & 154.47 & 96.89 \\
+MH-FFBP (100)  & 0.42 & 0.70 & 0.79 & 504.58 & 98.53 \\
 \hline
 \end{tabular}
+\end{center}
 }
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Smoother performance measures for noise covariance case 2. All results are averaged over 10 runs, with 500 time steps in each run.}
+\caption{Smoother performance measures for noise covariance case 2. All results are averaged over 100 runs, with 500 time steps in each run.}
 \label{tab:case2_performance}
 {\scriptsize
-\centering
+\begin{center}
 \begin{tabular}{|c||c|c|c|c|c|}
@@
 \hline
-FS              & 6.17 & 2.03 &  & 0 & 3.96 \\
-D-FFBS         & 5.79 & 1.89 &  & 63 & 7.34 \\
+FS             & 8.01 & 1.90 & 0.99 & 0.00 & 2.75 \\
+D-FFBS         & 7.71 & 1.70 & 0.92 & 66.40 & 14.06 \\
 \hline
-MH-FFBS (1)    & 6.13 & 2.01 &  & 2 & 4.77 \\
-MH-FFBS (3)    & 6.07 & 1.99 &  & 4 & 5.29 \\
-MH-FFBS (10)   & 5.94 & 1.94 &  & 10 & 6.1 \\
-MH-FFBS (30)   & 5.82 & 1.91 &  & 29 & 6.73 \\
-MH-FFBS (100)  & 5.80 & 1.89 &  & 95 & 7.38 \\
+MH-FFBS (1)    & 7.91 & 1.83 & 0.98 & 1.79 & 6.51 \\
+MH-FFBS (3)    & 7.83 & 1.78 & 0.96 & 3.86 & 8.63 \\
+MH-FFBS (10)   & 7.77 & 1.73 & 0.93 & 11.09 & 11.39 \\
+MH-FFBS (30)   & 7.75 & 1.71 & 0.92 & 31.77 & 13.14 \\
+MH-FFBS (100)  & 7.70 & 1.70 & 0.92 & 104.14 & 13.88 \\
 \hline
-MH-FFBP (1)    & 5.90 & 1.91 &  & 9 & 12.6 \\
-MH-FFBP (3)    & 5.67 & 1.81 &  & 18 & 22.7 \\
-MH-FFBP (10)   & 5.48 & 1.73 &  & 50 & 44.1 \\
-MH-FFBP (30)   & 5.43 & 1.69 &  & 144 & 70.9 \\
-MH-FFBP (100)  & 5.34 & 1.68 &  & 471 & 91.8 \\
+MH-FFBP (1)    & 7.75 & 1.72 & 0.92 & 9.25 & 22.97 \\
+MH-FFBP (3)    & 7.67 & 1.67 & 0.88 & 19.34 & 42.22 \\
+MH-FFBP (10)   & 7.62 & 1.63 & 0.85 & 54.61 & 70.06 \\
+MH-FFBP (30)   & 7.60 & 1.61 & 0.84 & 155.38 & 88.81 \\
+MH-FFBP (100)  & 7.59 & 1.60 & 0.84 & 507.58 & 96.82 \\
 \hline
 \end{tabular}
+\end{center}
+}
+\end{table}
+
+\begin{table}[!t]%
+\renewcommand{\arraystretch}{1.3}
+\caption{Smoother performance measures for noise covariance case 3. All results are averaged over 100 runs, with 500 time steps in each run.}
+\label{tab:case3_performance}
+{\scriptsize
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+Algorithm & \begin{minipage}[c]{0.9cm} Position RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  Velocity RMSE \end{minipage} & \begin{minipage}[c]{0.9cm}  ENEES \end{minipage} & \begin{minipage}[c]{0.9cm}  Running Time (s) \end{minipage} & \begin{minipage}[c]{0.9cm}  Unique Particles \end{minipage} \\
+\hline
+FS            & 7.41 & 2.15 & 0.98 & 0.00 & 3.97 \\
+D-FFBS        & 7.19 & 2.03 & 0.97 & 66.27 & 7.48 \\
+\hline
+MH-FFBS (1)   & 7.39 & 2.14 & 0.98 & 1.75 & 4.79 \\
+MH-FFBS (3)   & 7.36 & 2.12 & 0.98 & 3.75 & 5.33 \\
+MH-FFBS (10)  & 7.30 & 2.08 & 0.98 & 10.78 & 6.16 \\
+MH-FFBS (30)  & 7.24 & 2.05 & 0.97 & 30.83 & 6.88 \\
+MH-FFBS (100) & 7.20 & 2.04 & 0.97 & 101.02 & 7.38 \\
+\hline
+MH-FFBP (1)   & 7.22 & 2.05 & 0.97 & 9.12 & 12.57 \\
+MH-FFBP (3)   & 7.08 & 1.98 & 0.94 & 19.01 & 22.78 \\
+MH-FFBP (10)  & 6.95 & 1.92 & 0.92 & 53.64 & 44.15 \\
+MH-FFBP (30)  & 6.87 & 1.88 & 0.90 & 152.42 & 70.71 \\
+MH-FFBP (100) & 6.84 & 1.86 & 0.88 & 498.38 & 91.35 \\
+\hline
+\end{tabular}
+\end{center}
 }
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\\
+\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
 \caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%
@@
 \end{figure}
-
-{\meta Add comments on ENEES.}
 
@@
 
-When the number of MH steps, $M$, is small, the error of the MH-FFBS is only very slightly larger than that of the D-FFBS but with a significantly shorter running time.
+When the number of MH steps, $M$, is small, the error of the MH-FFBS is often still significantly reduced relative to the D-FFBS but with a shorter running time.
 For a particular application, it is possible to adjust $M$ to control the trade-off between error and speed of computation.
@@
 
-The MH-FFBP outperforms all of the other smoothers, achieving the lowest RMSE for any given running time.
+The MH-FFBP outperforms all of the other smoothers, achieving the lowest RMSE and ENEES for any given running time.
 Furthermore, because the support is not limited to using states from the filter distributions, the smoothing particle approximation has many more unique particles, giving a less degenerate representation.
@@
 
+\newpage
 \bibliographystyle{IEEEtran}

### 2012-06-08 MATLAB/batch_smoother_testing.m

@@ -148,7 +148,7 @@ for rs = 1:num_seed;
     
 end
 
-% Average and store
+%% Average and store
 results = cell(NA,1);
 for alg = 1:NA
     results{alg}.mean_pos_rmse = mean(mean_pos_rmse{alg});
@@ -158,4 +158,5 @@ for alg = 1:NA
     results{alg}.times = mean(times{alg});
 end
 
-save(['smoother_test' num2str(test_flag) '.mat'], 'results', 'params');
+%%
+save(['smoother_test' num2str(test_flag) '.mat'], 'results', 'times', 'unique_pts', 'mean_pos_rmse', 'mean_vel_rmse', 'mean_nees', 'params');

### 2012-06-08 MATLAB/plot_graphs_for_paper.m

@@ -4,9 +4,6 @@ clup
 for test_flag = [1,2,3];
     
     test_case = test_flag;
-    if test_flag == 3
-        test_case = 2;
-    end
     
     % Load data
     load(['smoother_test' num2str(test_flag) '.mat']);
@@ -18,7 +15,7 @@ for test_flag = [1,2,3];
     inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-b', 'markersize', 7 )
     inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'x-.g', 'markersize', 7 )
     
-    legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
+    legend('FS', 'D-FFBS', 'MH-FFBS (M=1-100)', 'MH-FFBP (M=1-100)');
     xlabel('Running Time (s)'); ylabel('Position RMSE');
     xlimits = get(gca, 'XLim');
     ylimits = get(gca, 'YLim');
@@ -35,15 +32,15 @@ for test_flag = [1,2,3];
     inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), '*-b', 'markersize', 7 )
     inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'x-.g', 'markersize', 7 )
     
-    legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
-    xlabel('Running Time (s)'); ylabel('MNEES');
+    legend('FS', 'D-FFBS', 'MH-FFBS (M=1-100)', 'MH-FFBP (M=1-100)');
+    xlabel('Running Time (s)'); ylabel('ENEES');
     xlimits = get(gca, 'XLim');
-    ylimits = get(gca, 'YLim');
+    ylimits = [0.75 1];
     xlim([-50, xlimits(2)]);
     plot([0 0], ylimits, ':k');
     ylim(ylimits);
     
-    export_pdf(fig, ['case' num2str(test_case) '_smoother_comparison_MNEES_time.pdf']);
[…]

### 2012-09-13 Poster/MCMC_smoothing_poster.tex
Added poster

\documentclass[landscape]{sciposter}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{IEEEtrantools}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{algorithmic}

\onehalfspacing
%\multicoltolerance = 1000
%\raggedcolumns

\title{Metropolis-Hastings Implementations of the Forward-Filtering-Backward-Sampling \\ Algorithm for Particle Smoothing}

\author{P. Bunch and S. Godsill}
\institute{Department of Engineering, University of Cambridge, Trumpington Street, Cambridge, CB2 1PZ, UK}
\email{\{pb404,sjg\}@eng.cam.ac.uk}

\leftlogo[2]{CUnibig.pdf}

\renewcommand{\titlesize}{\huge}

\begin{document}

\maketitle
\begin{multicols}{4}

\begin{abstract}
The conventional forward-filtering-backward-sampling algorithm for particle smoothing generates samples from the joint smoothing distribution by sequentially sampling backwards from the forward filter approximations. This algorithm has a computational complexity quadratic in the number of particles, which is often restrictive. In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation. Here a Metropolis-Hastings sampling procedure is used to improve the efficiency of the forward-filtering-backward-sampling algorithm, achieving comparable error performance but with a lower execution time. In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error. These algorithms also provide a greater degree of flexibility over existing methods, allowing a trade-off between execution time and estimation error, controlled by the length of the Markov chains.
\end{abstract}

%\columnbreak

\section*{Introduction}

Consider a standard hidden Markov model for Bayesian inference,
%
\begin{IEEEeqnarray*}{rCl}
x_{k} & \sim & p(x_{k}|x_{k-1}) \\
y_{k} & \sim & p(y_{k}|x_{k})      .
\end{IEEEeqnarray*}

We address the task of joint smoothing --- inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps).

%Commonly of interest are the problems of filtering and smoothing. Filtering is the inference of $p(x_k|y_{1:k})$. There are two related smoothing tasks:
%\begin{itemize}
%  \item Joint smoothing --- inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps).
%  \item Marginal smoothing --- inference of $p(x_{k}|y_{1:K})$ for each $k$.
%\end{itemize}
%Here we are concerned with joint smoothing.

For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
\begin{itemize}
  \item Filter: $\hat{p}(x_{k}|y_{1:k}) = \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)$.
  \item Smoother: $\hat{p}(x_{1:K}|y_{1:K}) = \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})$.
\end{itemize}

\section*{Forward-Filtering-Backward-Sampling}

The forward-filtering-backward-sampling (FFBS) algorithm \cite{Godsill2004} draws particles from the joint smoothing distribution by sequentially sampling a set of backwards conditional distributions,
%
\begin{IEEEeqnarray*}{rCl}
p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K})     .
\end{IEEEeqnarray*}

The conditional factors may be expressed as,
%
\begin{IEEEeqnarray*}{rCl}
p(x_k|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_k|\tilde{x}_{k+1}, y_{1:k})   \\
                                & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|y_{1:k})   .
\end{IEEEeqnarray*}

where $\tilde{x}_{k+1}$ is the future state already sampled. Using the filter approximations, the backwards conditional distributions are approximated as,
%
\begin{IEEEeqnarray*}{rCl}
\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k})    \\
\tilde{w}_k^{(i)} & \propto & w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)})      .
\end{IEEEeqnarray*}

This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting multinomial distribution. The whole procedure is repeated $N_S$ times, once for each smoothing particle.

\begin{algorithm}
\begin{algorithmic}[1]
\linespread{1.7} \selectfont
 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$ and store the resulting particles $\{x_k^{(i)}, w_k^{(i)}\}$.
	\FOR{$i = 1$ \TO $N_S$}
		\STATE Sample $\tilde{x}_{K}^{(i)} \sim \hat{p}(x_K|y_{1:K})$.
		\FOR{$k = K-1$ \TO $1$}
            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. This is the critical step. See text for various methods.
%			\FOR{$j = 1$ \TO $N_F$}
%				\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
%			\ENDFOR
%			\STATE Sample $\tilde{x}_{k}^{(i)} \sim \sum_j \tilde{w}_k^{(j)} \delta_{x_{k}^{(j)}}(x_{k})$.
		\ENDFOR
	\ENDFOR
\end{algorithmic}
\caption{Forward-filtering-backward-sampling algorithm}
\end{algorithm}

\subsection*{Deficiencies}
%
The basic FFBS algorithm suffers from two drawbacks:
\begin{itemize}
  \item The algorithmic complexity is $\mathcal{O}(N_F \times N_S)$ which is restrictive.
  \item The support for the smoothing approximation is limited to the values appearing in the filtering particles.
\end{itemize}

\section*{A Metropolis-Hastings implementation}

The FFBS bottleneck is the calculation of the $N_F$ backwards weights for every sampling step, in order to sample from $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$. Instead, we can sample this distribution using Metropolis-Hastings (MH).

\begin{itemize}
  \item The filter weights are used to propose particles from $\hat{p}(x_k | y_{1:k})$.
  \item The new state, $x_k^*$, replaces the old state, $x_k$ with acceptance probability,
\end{itemize}
%
\begin{IEEEeqnarray*}{rCl}
\alpha &=& \min \Bigg[1, \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k) }  \Bigg]     .
\end{IEEEeqnarray*}

\columnbreak

\subsection*{Initialisation}
%
\begin{itemize}
  \item An initial value is required to start the chain.
  \item Instead of sampling the state, $x_k$, at each step, sample the entire history, $x_{1:k}$.
  \item No additional states need be stored, only index of the $(k-1)$-time particle from which each $k$-time particle originates.
  \item No burn-in required --- initial state is drawn from target distribution.
\end{itemize}

\subsection*{Effects}
%
\begin{itemize}
  \item Computational complexity reduced to $\mathcal{O}(M \times N_S)$, where $M$ is the  length of Markov chains.
  \item $M = 0$ reproduces particles from the filter smoother. $M \rightarrow \infty$ draws an independent sample from the distribution.
  \item Performance expected to be bounded by that of the filter smoother and the standard FFBS.
  \item Much faster than FFBS when $M$ can be made small, i.e. when acceptance rate is high.
\end{itemize}

%\begin{algorithm}
%\begin{algorithmic}[1]
%%\linespread{1.7} \selectfont
% 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$ and store the resulting particles $\{x_{1:k}^{(i)}, w_k^{(i)}\}$.
%	\FOR{$i = 1$ \TO $N_S$}
%		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
%		\FOR{$k = K-1$ \TO $1$}
%			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
%			\FOR{$m = 1$ \TO $M$}
%				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
%				\STATE With probability $\alpha$,\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
%			\ENDFOR
%			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
%		\ENDFOR
%	\ENDFOR
%\end{algorithmic}
%\caption{Metropolis-Hastings forward-filtering-backward-sampling (MH-FFBS) algorithm}
%\end{algorithm}

\subsection*{Alternatives}
%
Rejection sampling \cite{Douc2011} can achieve a similar complexity reduction.
\begin{itemize}
  \item Rejection sampling can be very slow if rejection rates are high, e.g. when state dimension high.
  \item MH method more flexible --- chain length can be set to trade-off complexity against running time.
\end{itemize}

\section*{Improving the Support}

Modify the MH proposal to use,
%
\begin{IEEEeqnarray*}{c}
\hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
\end{IEEEeqnarray*}

Support of $x_k$ is no longer restricted to those values appearing in the filter approximation. This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.

Acceptance probability is now,
%
\begin{IEEEeqnarray*}{rCl}
\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }  \Bigg]      .
\end{IEEEeqnarray*}

%\begin{algorithm}
%\begin{algorithmic}[1]
%%\linespread{1.7} \selectfont
% 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
%	\FOR{$i = 1$ \TO $N_S$}
%		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
%		\FOR{$k = K-1$ \TO $1$}
%			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
%			\FOR{$m = 1$ \TO $M$}
%				\STATE Propose a new state history,\\ $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
%				\STATE Propose a new state, $x_{k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
%				\STATE With probability $\alpha$,\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
%			\ENDFOR
%			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
%		\ENDFOR
%	\ENDFOR
%\end{algorithmic}
%\caption{Metropolis-Hastings forward-filtering-backward-proposing (MH-FFBP) algorithm}
%\end{algorithm}

%\columnbreak

\section*{Demonstrations}

New algorithms tested on a simulated tracking problem with near constant velocity dynamic model and range-bearing observation model. Two observation covariance cases are considered.
%
\begin{figure}
\centering
\subfigure[Case 1]{\includegraphics[width=0.45\columnwidth]{POSTER_case1_trajectory.pdf}}
\subfigure[Case 2]{\includegraphics[width=0.45\columnwidth]{POSTER_case2_trajectory.pdf}}
\caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
\end{figure}

\begin{figure}[!t]
\centering
\subfigure[Case 1]{\includegraphics[width=0.45\columnwidth]{POSTER_case1_smoother_comparison_posRMSE_time.pdf}}
\subfigure[Case 2]{\includegraphics[width=0.45\columnwidth]{POSTER_case2_smoother_comparison_posRMSE_time.pdf}}
\caption{Position RMSEs and running times of various smoothing algorithms with two noise covariance cases.}%
\end{figure}

\section*{Conclusions}

\begin{itemize}
  \item New smoothing algorithms achieve lower error for a given running time. Similar results apply to consistency of estimates.
  \item Chain length provides flexibility to trade-off running time against quality of estimate.
\end{itemize}

{\small
\bibliographystyle{IEEEtran}
\bibliography{D:/pb404/Dropbox/PhD/Cleanbib}
}

\end{multicols}

\end{document}

### 2012-09-13 Poster/abstract.tex
Added poster

--- 2012-09-13 Poster/MCMC_smoothing_poster.tex
@@
-\documentclass[landscape]{sciposter}
+Metropolis-Hastings Implementations of the Forward-Filtering-Backward-Sampling Algorithm for Particle Smoothing
 
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{IEEEtrantools}
-\usepackage{multicol}
-\usepackage{setspace}
-\usepackage{algorithmic}
-
-\onehalfspacing
-%\multicoltolerance = 1000
-%\raggedcolumns
-
-\title{Metropolis-Hastings Implementations of the Forward-Filtering-Backward-Sampling \\ Algorithm for Particle Smoothing}
-
-\author{P. Bunch and S. Godsill}
-\institute{Department of Engineering, University of Cambridge, Trumpington Street, Cambridge, CB2 1PZ, UK}
-\email{\{pb404,sjg\}@eng.cam.ac.uk}
-
-\leftlogo[2]{CUnibig.pdf}
-
-\renewcommand{\titlesize}{\huge}
-
-\begin{document}
-
-\maketitle
-\begin{multicols}{4}
-
-\begin{abstract}
 The conventional forward-filtering-backward-sampling algorithm for particle smoothing generates samples from the joint smoothing distribution by sequentially sampling backwards from the forward filter approximations.
@@
 These algorithms also provide a greater degree of flexibility over existing methods, allowing a trade-off between execution time and estimation error, controlled by the length of the Markov chains.
-\end{abstract}
-
-%\columnbreak
-
-\section*{Introduction}
-
-Consider a standard hidden Markov model for Bayesian inference,
-%
-\begin{IEEEeqnarray*}{rCl}
-x_{k} & \sim & p(x_{k}|x_{k-1}) \\
-y_{k} & \sim & p(y_{k}|x_{k})      .
-\end{IEEEeqnarray*}
-
-We address the task of joint smoothing --- inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps).
-
-%Commonly of interest are the problems of filtering and smoothing. Filtering is the inference of $p(x_k|y_{1:k})$. There are two related smoothing tasks:
-%\begin{itemize}
-%  \item Joint smoothing --- inference of $p(x_{1:K}|y_{1:K})$ (where $K$ is the number of time steps).
-%  \item Marginal smoothing --- inference of $p(x_{k}|y_{1:K})$ for each $k$.
-%\end{itemize}
-%Here we are concerned with joint smoothing.
-
-For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
-\begin{itemize}
-  \item Filter: $\hat{p}(x_{k}|y_{1:k}) = \sum_i w_k^{(i)} \delta_{x_k^{(i)}}(x_k)$.
-  \item Smoother: $\hat{p}(x_{1:K}|y_{1:K}) = \sum_i w_K^{(i)} \delta_{x_{1:K}^{(i)}}(x_{1:K})$.
-\end{itemize}
-
-\section*{Forward-Filtering-Backward-Sampling}
-
-The forward-filtering-backward-sampling (FFBS) algorithm \cite{Godsill2004} draws particles from the joint smoothing distribution by sequentially sampling a set of backwards conditional distributions,
-%
-\begin{IEEEeqnarray*}{rCl}
-p(x_{1:K}|y_{1:K}) & = & p(x_K|y_{1:K}) \prod_{k=1}^{K-1} p(x_k|x_{k+1}, y_{1:K})     .
-\end{IEEEeqnarray*}
-
-The conditional factors may be expressed as,
-%
-\begin{IEEEeqnarray*}{rCl}
-p(x_k|\tilde{x}_{k+1}, y_{1:K}) & =       & p(x_k|\tilde{x}_{k+1}, y_{1:k})   \\
-                                & \propto & p(\tilde{x}_{k+1}|x_k) p(x_k|y_{1:k})   .
-\end{IEEEeqnarray*}
-
-where $\tilde{x}_{k+1}$ is the future state already sampled. Using the filter approximations, the backwards conditional distributions are approximated as,
-%
-\begin{IEEEeqnarray*}{rCl}
-\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K}) & = & \sum_i  \tilde{w}_k^{(i)} \delta_{x_{k}^{(i)}}(x_{k})    \\
-\tilde{w}_k^{(i)} & \propto & w_k^{(i)} p(\tilde{x}_{k+1}|x_k^{(i)})      .
-\end{IEEEeqnarray*}
-
-This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting multinomial distribution.
-The whole procedure is repeated $N_S$ times, once for each smoothing particle.
-
-\begin{algorithm}
-\begin{algorithmic}[1]
-\linespread{1.7} \selectfont
- 	\STATE Run a particle filter to approximate $p(x_k|y_{1:k})$ for each $k$ and store the resulting particles $\{x_k^{(i)}, w_k^{(i)}\}$.
-	\FOR{$i = 1$ \TO $N_S$}
-		\STATE Sample $\tilde{x}_{K}^{(i)} \sim \hat{p}(x_K|y_{1:K})$.
-		\FOR{$k = K-1$ \TO $1$}
-            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. This is the critical step. See text for various methods.
-%			\FOR{$j = 1$ \TO $N_F$}
-%				\STATE Calculate weight $\tilde{w}_k^{(j)} = w_k^{(j)} p(\tilde{x}_{k+1}|x_k^{(j)})$.
-%			\ENDFOR
-%			\STATE Sample $\tilde{x}_{k}^{(i)} \sim \sum_j \tilde{w}_k^{(j)} \delta_{x_{k}^{(j)}}(x_{k})$.
-		\ENDFOR
-	\ENDFOR
-\end{algorithmic}
-\caption{Forward-filtering-backward-sampling algorithm}
-\end{algorithm}
-
-\subsection*{Deficiencies}
-%
-The basic FFBS algorithm suffers from two drawbacks:
-\begin{itemize}
-  \item The algorithmic complexity is $\mathcal{O}(N_F \times N_S)$ which is restrictive.
-  \item The support for the smoothing approximation is limited to the values appearing in the filtering particles.
-\end{itemize}
-
-\section*{A Metropolis-Hastings implementation}
-
-The FFBS bottleneck is the calculation of the $N_F$ backwards weights for every sampling step, in order to sample from $\hat{p}(x_k|\tilde{x}_{k+1}, y_{1:K})$.
-Instead, we can sample this distribution using Metropolis-Hastings (MH).
-
-\begin{itemize}
-  \item The filter weights are used to propose particles from $\hat{p}(x_k | y_{1:k})$.
-  \item The new state, $x_k^*$, replaces the old state, $x_k$ with acceptance probability,
-\end{itemize}
-%
-\begin{IEEEeqnarray*}{rCl}
-\alpha &=& \min \Bigg[1, \frac{ p(\tilde{x}_{k+1}|x_k^{*}) }{ p(\tilde{x}_{k+1}|x_k) }  \Bigg]     .
-\end{IEEEeqnarray*}
-
-\columnbreak
-
-\subsection*{Initialisation}
-%
-\begin{itemize}
-  \item An initial value is required to start the chain.
-  \item Instead of sampling the state, $x_k$, at each step, sample the entire history, $x_{1:k}$.
-  \item No additional states need be stored, only index of the $(k-1)$-time particle from which each $k$-time particle originates.
-  \item No burn-in required --- initial state is drawn from target distribution.
-\end{itemize}
-
-\subsection*{Effects}
-%
-\begin{itemize}
-  \item Computational complexity reduced to $\mathcal{O}(M \times N_S)$, where $M$ is the  length of Markov chains.
-  \item $M = 0$ reproduces particles from the filter smoother. $M \rightarrow \infty$ draws an independent sample from the distribution.
-  \item Performance expected to be bounded by that of the filter smoother and the standard FFBS.
-  \item Much faster than FFBS when $M$ can be made small, i.e. when acceptance rate is high.
-\end{itemize}
-
-%\begin{algorithm}
-%\begin{algorithmic}[1]
-%%\linespread{1.7} \selectfont
-% 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$ and store the resulting particles $\{x_{1:k}^{(i)}, w_k^{(i)}\}$.
-%	\FOR{$i = 1$ \TO $N_S$}
-%		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
-%		\FOR{$k = K-1$ \TO $1$}
-%			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
-%			\FOR{$m = 1$ \TO $M$}
-%				\STATE Propose a new state, $x_{1:k}^{(i)*} \sim \sum_j \tilde{v}_k^{(j)} \delta_{x_{1:k}^{(j)}}(x_{1:k})$.
-%				\STATE With probability $\alpha$,\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
-%			\ENDFOR
-%			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
-%		\ENDFOR
-%	\ENDFOR
-%\end{algorithmic}
-%\caption{Metropolis-Hastings forward-filtering-backward-sampling (MH-FFBS) algorithm}
-%\end{algorithm}
-
-\subsection*{Alternatives}
-%
-Rejection sampling \cite{Douc2011} can achieve a similar complexity reduction.
-\begin{itemize}
-  \item Rejection sampling can be very slow if rejection rates are high, e.g. when state dimension high.
-  \item MH method more flexible --- chain length can be set to trade-off complexity against running time.
-\end{itemize}
-
-\section*{Improving the Support}
-
-Modify the MH proposal to use,
-%
-\begin{IEEEeqnarray*}{c}
-\hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
-\end{IEEEeqnarray*}
-
-Support of $x_k$ is no longer restricted to those values appearing in the filter approximation. This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.
-
-Acceptance probability is now,
-%
-\begin{IEEEeqnarray*}{rCl}
-\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }  \Bigg]      .
-\end{IEEEeqnarray*}
-
-%\begin{algorithm}
-%\begin{algorithmic}[1]
-%%\linespread{1.7} \selectfont
-% 	\STATE Run a particle filter to approximate $p(x_{1:k}|y_{1:k})$ for each $k$.
-%	\FOR{$i = 1$ \TO $N_S$}
-%		\STATE Sample $\tilde{x}_{1:K}^{(i)} \sim \sum_j w_K^{(j)} \delta_{x_{1:K}^{(j)}}(x_{1:K})$.
-%		\FOR{$k = K-1$ \TO $1$}
-%			\STATE $x_{1:k}^{(i)(0)} \gets \tilde{x}_{1:k}^{(i)}$.
-%			\FOR{$m = 1$ \TO $M$}
-%				\STATE Propose a new state history,\\ $x_{1:k-1}^{(i)*} \sim \sum_j \tilde{v}_{k-1}^{(j)} \delta_{x_{1:k-1}^{(j)}}(x_{1:k-1})$.
-%				\STATE Propose a new state, $x_{k}^{(i)*} \sim q(x_k|x_{k-1}^{*},\tilde{x}_{k+1},y_k)$.
-%				\STATE With probability $\alpha$,\\ $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)*}$. Otherwise, $x_{1:k}^{(i)(m)} \gets x_{1:k}^{(i)(m-1)}$.
-%			\ENDFOR
-%			\STATE $\tilde{x}_{1:k}^{(i)} \gets x_{1:k}^{(i)(M)}$
-%		\ENDFOR
-%	\ENDFOR
-%\end{algorithmic}
-%\caption{Metropolis-Hastings forward-filtering-backward-proposing (MH-FFBP) algorithm}
-%\end{algorithm}
-
-%\columnbreak
-
-\section*{Demonstrations}
-
-New algorithms tested on a simulated tracking problem with near constant velocity dynamic model and range-bearing observation model. Two observation covariance cases are considered.
-%
-\begin{figure}
-\centering
-\subfigure[Case 1]{\includegraphics[width=0.45\columnwidth]{POSTER_case1_trajectory.pdf}}
-\subfigure[Case 2]{\includegraphics[width=0.45\columnwidth]{POSTER_case2_trajectory.pdf}}
-\caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\subfigure[Case 1]{\includegraphics[width=0.45\columnwidth]{POSTER_case1_smoother_comparison_posRMSE_time.pdf}}
-\subfigure[Case 2]{\includegraphics[width=0.45\columnwidth]{POSTER_case2_smoother_comparison_posRMSE_time.pdf}}
-\caption{Position RMSEs and running times of various smoothing algorithms with two noise covariance cases.}%
-\end{figure}
-
-\section*{Conclusions}
-
-\begin{itemize}
-  \item New smoothing algorithms achieve lower error for a given running time. Similar results apply to consistency of estimates.
-  \item Chain length provides flexibility to trade-off running time against quality of estimate.
-\end{itemize}
-
-{\small
-\bibliographystyle{IEEEtran}
-\bibliography{D:/pb404/Dropbox/PhD/Cleanbib}
-}
-
-\end{multicols}
-
-\end{document}

### 2013-11-28 Paper/complete/12-11-14_final_submission/IEEE-Godsill.tex
Added clean version of code.

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Godsill.jpg}}]{Simon Godsill} is Professor of Statistical Signal Processing in the Engineering Department at Cambridge University and a Professorial Fellow and tutor at Corpus Christi College Cambridge.

He coordinates an active research group in Signal Inference and its Applications within the Signal Processing and Communications Laboratory at Cambridge, specialising in Bayesian computational methodology, multiple object tracking, audio and music processing, and financial time series modeling. A particular methodological theme over recent years has been the development of novel techniques  for optimal Bayesian filtering and smoothing, using Sequential Monte Carlo or Particle Filtering methods.

Prof. Godsill has published extensively in journals, books and international conference proceedings. He was technical chair of the IEEE NSSPW workshop in 2006 on sequential and nonlinear filtering methods, was Technical Chair for Fusion 2010 in Edinburgh, and has been on the conference panel for numerous other conferences and workshops. Prof. Godsill has served as Associate Editor for IEEE Tr. Signal Processing and the journal Bayesian Analysis. He was Theme Leader in Tracking and Reasoning over Time for the UK’s Data and Information Fusion Defence Technology Centre (DIF-DTC) and Principal Investigator on grants funded by the EU, EPSRC, QinetiQ, General Dynamics, MOD, Microsoft UK, Citibank and Mastercard. In 2009-10 he was co-organiser of an 18 month research program in Sequential Monte Carlo Methods at the SAMSI Institute in North Carolina. He is a Director of CEDAR Audio Ltd. (which has received a technical Oscar for its audio processing work), and Input Dynamics Ltd., both companies which use his research work in the audio area.

\end{IEEEbiography}

### 2013-11-28 Paper/complete/12-11-14_final_submission/double_MCMCsmoother.tex
Added clean version of code.

--- 2012-06-08 Paper/complete/revisions_single_MCMCsmoother.tex
@@
-\documentclass[peerreview,11pt,draftcls,onecolumn]{IEEEtran}
+\documentclass[10pt,twocolumn,twoside]{IEEEtran}
 
@@
 \usepackage{cite}
-\usepackage[pdftex]{graphicx} \graphicspath{{images/}}
+\usepackage[pdftex]{graphicx}
 \usepackage[cmex10]{amsmath}
-%\usepackage{algorithm}
 \usepackage{algorithmic}
@@
 \usepackage{fixltx2e}
-%\usepackage{stfloats}
 \usepackage{url}
 \usepackage{color}
-
-\newenvironment{meta}[0]{\color{red} \em}{}
-
-% correct bad hyphenation here
-\hyphenation{}
 
@@
         Simon~Godsill,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
+\thanks{Copyright (c) 2012 IEEE.
+Personal use of this material is permitted.
+However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}
 \thanks{P. Bunch and S. Godsill are with the Department
 of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}% <-this % stops a space
-\thanks{Manuscript received April 03, 2012.}}
-
-%% The paper headers
-%\markboth{IEEE Transactions on Signal Processing,~Vol.~AA, No.~BB, January~YYYY}%
-%{MCMC Particle Smoothing Approximations}
+}
 
@@
 
-%\begin{IEEEkeywords}
-%state space model, particle filter, smoothing, MCMC, Bayesian inference
-%\end{IEEEkeywords}
-
-\ifCLASSOPTIONpeerreview
-\begin{center} \bfseries EDICS Categories: MLR-BAYL, SSP-NGAU, SSP-TRAC, SSP-FILT \end{center}
-\fi
-
-\newpage
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-%\IEEEpeerreviewmaketitle
+\begin{IEEEkeywords}
+state space model, particle filter, smoothing, MCMC, Bayesian inference
+\end{IEEEkeywords}
 
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{fig6-a.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{fig6-b.pdf}}
 \\
-\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_trajectory.pdf}}
+\subfloat[Case 3]{\includegraphics[width=0.9\columnwidth]{fig6-c.pdf}}
 \caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
@@
 {\scriptsize
-\begin{center}
+\centering
 \begin{tabular}{|c||c|c|c|c|c|}
@@
 \hline
+
 \end{tabular}
-\end{center}
 }
@@
 \end{tabular}
-\end{center}
 }
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.9\columnwidth]{fig7-a.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.9\columnwidth]{fig7-b.pdf}}
 \\
-\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 3]{\includegraphics[width=0.9\columnwidth]{fig7-c.pdf}}
 \caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%
@@
 
-%\section{Appendix Title}
-%Appendix one text goes here.
-
-%\section*{Acknowledgment}
-%The authors would like to thank...
-
 \newpage
 \bibliographystyle{IEEEtran}
-\bibliography{D:/pb404/Dropbox/PhD/Cleanbib}
+%\bibliography{Cleanbib}
+\input{MCMCsmoother.bbl}
 
@@
 
-%\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Bunch.tex}
-%\input{D:/pb404/Dropbox/PhD/Biographies/IEEE-Godsill.tex}
+\input{IEEE-Bunch.tex}
+\input{IEEE-Godsill.tex}
 

### 2013-11-28 Paper/complete/preprint_MCMCsmoother.tex
Added clean version of code.

--- 2012-06-08 Paper/complete/revisions_single_MCMCsmoother.tex
@@
-\documentclass[peerreview,11pt,draftcls,onecolumn]{IEEEtran}
+\documentclass{article}
 
 %% Packages
+\usepackage{IEEEtrantools}
 \usepackage{cite}
@@
 
-\author{Pete~Bunch*,~\IEEEmembership{Member,~IEEE,} and
-        Simon~Godsill,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
-\thanks{P. Bunch and S. Godsill are with the Department
-of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}% <-this % stops a space
-\thanks{Manuscript received April 03, 2012.}}
+\author{Pete~Bunch and Simon~Godsill% <-this % stops a space
+\thanks{Department of Engineering, Cambridge University, UK. email: \{pb404,sjg30\}@cam.ac.uk}}
 
@@
 
-%\begin{IEEEkeywords}
-%state space model, particle filter, smoothing, MCMC, Bayesian inference
-%\end{IEEEkeywords}
-
-\ifCLASSOPTIONpeerreview
-\begin{center} \bfseries EDICS Categories: MLR-BAYL, SSP-NGAU, SSP-TRAC, SSP-FILT \end{center}
-\fi
-
-\newpage
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-%\IEEEpeerreviewmaketitle
-
 \section{Introduction} \label{sec:intro}
 
-\IEEEPARstart{T}{he} objective of sequential Bayesian inference is the estimation of an unknown, time-varying quantity from incomplete or inaccurate observations.
+The objective of sequential Bayesian inference is the estimation of an unknown, time-varying quantity from incomplete or inaccurate observations.
 This is commonly achieved through the use of discrete-time probabilistic models for the evolution of the latent state and the measurement process, defined by transition and observation densities,
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_trajectory.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_trajectory.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_trajectory.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_trajectory.pdf}}
 \\
-\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_trajectory.pdf}}
+\subfloat[Case 3]{\includegraphics[width=0.4\columnwidth]{case3_trajectory.pdf}}
 \caption{Example trajectory with observations from the two noise covariance cases. Constant likelihood contours are shown for the first and last position (dashed)}
@@
 \centering
-\subfloat[Case 1]{\includegraphics[width=0.45\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 1]{\includegraphics[width=0.4\columnwidth]{case1_smoother_comparison_posRMSE_time.pdf}}
 \\
-\subfloat[Case 2]{\includegraphics[width=0.45\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 2]{\includegraphics[width=0.4\columnwidth]{case2_smoother_comparison_posRMSE_time.pdf}}
 \\
-\subfloat[Case 3]{\includegraphics[width=0.45\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
+\subfloat[Case 3]{\includegraphics[width=0.4\columnwidth]{case3_smoother_comparison_posRMSE_time.pdf}}
 \caption{Position RMSEs and running times of various smoothing algorithms with different noise covariance cases.}%

### 2013-11-28 Poster/MCMC_smoothing_poster.tex
Added clean version of code.

@@
 In addition, the support of the smoothing distribution is restricted to the values which appear in the filtering approximation.
-Here a Metropolis-Hastings sampling procedure is used to improve the efficiency of the forward-filtering-backward-sampling algorithm, achieving comparable error performance but with a lower execution time.
+Here a Metropolis-Hastings sampling procedure is used to improve the efficiency of the forward-filtering-backward-sampling algorithm.
 In addition, an algorithm for approximating the joint smoothing distribution without limited support is presented, which achieves simultaneous improvements in both execution time and error.
@@
 
+\begingroup
+{\small
+\renewcommand{\section}[2]{}%
+To be published as:
+\begin{thebibliography}
+\newblock P.~Bunch and S.~Godsill.
+\newblock ``Improved particle approximations to the joint smoothing distribution using Markov chain Monte Carlo.''
+\newblock {\em IEEE Transactions on Signal Processing}.
+\end{thebibliography}
+}
+\endgroup
+
 \section*{Introduction}
 
-Consider a standard hidden Markov model for Bayesian inference,
+Consider a hidden Markov model for Bayesian inference,
 %
@@
 
-For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
+%For non-linear, non-Gaussian models, numerical approximations are needed. Particle methods represent posterior distributions with a set of samples \cite{Doucet2009}.
+For non-linear, non-Gaussian models, particle methods may be used to represent posterior distributions with a set of samples \cite{Doucet2009}.
 \begin{itemize}
@@
 
-This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting multinomial distribution.
+This may be sampled easily by calculating $\tilde{w}_k^{(i)}$ for each of the $N_F$ filter particles and sampling the resulting categorical distribution.
 The whole procedure is repeated $N_S$ times, once for each smoothing particle.
@@
 		\FOR{$k = K-1$ \TO $1$}
-            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. This is the critical step. See text for various methods.
+            \STATE Sample $\tilde{x}_{k}^{(i)} \sim p(x_k|\tilde{x}_{k+1}, y_{1:K})$. \\
+            \emph{This is the critical step. See text for various methods.}
 %			\FOR{$j = 1$ \TO $N_F$}
@@
 %
-\begin{IEEEeqnarray*}{c}
-\hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
+\begin{IEEEeqnarray*}{rCl}
+\{ x_{1:k-1}^{(i^*)}, x_k^*  \} &\sim& \hat{p}(x_{1:k-1}|y_{1:k-1}) q(x_k|x_{k-1}, \tilde{x}_{k+1}, y_k)     .
 \end{IEEEeqnarray*}
 
-Support of $x_k$ is no longer restricted to those values appearing in the filter approximation. This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.
+Support of $x_k$ is no longer restricted to those values appearing in the filter approximation (although support of $x_{1:k-1}$ is).
+This is inspired by the method used in \cite{Fearnhead2010} for two-filter marginal smoothing.
 
@@
 \begin{IEEEeqnarray*}{rCl}
-\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{*}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{*},\tilde{x}_{k+1},y_k) }  \Bigg]      .
+\alpha &=& \min \Bigg[ 1, \frac{  p(\tilde{x}_{k+1}|x_k^{*}) p(x_k^{*}|x_{k-1}^{(i^*)}) p(y_k|x_k^{*}) q(x_k|x_{k-1},\tilde{x}_{k+1},y_k) }{ p(\tilde{x}_{k+1}|x_k) p(x_k|x_{k-1}) p(y_k|x_k) q(x_k^{*}|x_{k-1}^{(i^*)},\tilde{x}_{k+1},y_k) }  \Bigg]      .
 \end{IEEEeqnarray*}

### 2013-11-28 MATLAB/Submit_grid_job.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+qsub -m be -t 1-100 ge_script.sh

### 2013-11-28 MATLAB/backward_simulation_smoother.m

@@ -1,0 +1,140 @@
+function [ ps, diagnostics ] = backward_simulation_smoother( display, algo, model, fh, pf, observ, flag_samp_type, true_state )
+%BACKWARD_SIMULATION_SMOOTHER Run backward simulation smoothing with one of
+%a variety of sampling methods
+
+% flag_samp_type
+% 0 = filter
+% 1 = direct (exhaustive evaluation)
+% 2 = rejection sampling with early stopping
+% 3 = mh sampling
+
+if display.text
+    fprintf(1, 'Running backward simulation smoother, type %u.\n', flag_samp_type);
+end
+
+% Local copies
+K = model.K;
+N = algo.N;
+S = algo.S;
+M = algo.M;
+
+% If we're just resampling the filter particles, keep N of them
+if flag_samp_type == 0
+    S = N;
+end
+
+% Set up diagnostics
+switch flag_samp_type  
+    case 2
+        diagnostics.time_out = false(S,K);
+    case 3
+        diagnostics.acceptance_rate = zeros(S,K);
+    otherwise
+        diagnostics = [];
+end
+
+% Resample final frame filter particles
+filt_ancestor = sample_weights(pf(K).weight, S);
+
+% Particle loop
+for ii = 1:S
+    
+    if display.text
+        fprintf(1, '   Simulating trajectory %u.\n', ii);
+    end
+    
+    % Initialise new particle index and state trajectory
+    traj_ind = zeros(1,K);
+    traj_state = zeros(model.ds,K);
+    traj_ind(K) = filt_ancestor(ii);
+    traj_state(:,K) = pf(K).state(:,traj_ind(K));
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        % Get next state
+        next_state = traj_state(:,kk+1);
+        
+        % Sample by appropriate method
+        switch flag_samp_type
+            
+            case 0
+                % Reproduce the filter particle
+                next_ind = traj_ind(kk+1);
+                traj_ind(kk) = pf(kk+1).ancestor(next_ind);
+                traj_state(:,kk) = pf(kk).state(:,traj_ind(kk));
+            
+            case 1
+                % Direct sampling
+                bs_weight = zeros(N,1);
+                for jj = 1:N
+                    [~, trans] = tracking_transition(model, pf(kk).state(:,jj), next_state);
+                    bs_weight(jj) = pf(kk).weight(jj) + trans;
+                end
[…]

### 2013-11-28 MATLAB/batch_bs_processing.m

@@ -1,0 +1,31 @@
+clup;
+
+test_name = 'model_case_2'
+num_tests = 100;
+num_algs = 17;
+
+mean_pos_rmse = zeros(num_algs,num_tests);
+mean_vel_rmse = zeros(num_algs,num_tests);
+mean_tnees = zeros(num_algs,num_tests);
+mean_nus = zeros(num_algs,num_tests);
+mean_rt = zeros(num_algs,num_tests);
+
+for tt = 1:num_tests
+    
+    load(['batch_tests/' test.name '/bs_test_' num2str(test_num) '.mat']);
+    
+    for aa = 1:num_algs
+        mean_pos_rmse(aa,tt) = mean(results.pos_rmse{aa});
+        mean_vel_rmse(aa,tt) = mean(results.vel_rmse{aa});
+        mean_tnees(aa,tt) = mean(results.tnees{aa});
+        mean_nus(aa,tt) = mean(results.nus{aa});
+        mean_rt(aa,tt) = mean(results.rt{aa});
+    end
+    
+end
+
+mean_mean_pos_rmse = mean(mean_pos_rmse,2);
+mean_mean_vel_rmse = mean(mean_vel_rmse,2);
+mean_mean_tnees = mean(mean_tnees,2);
+mean_mean_nus = mean(mean_nus,2);
+mean_mean_rt = mean(mean_rt,2);

### 2013-11-28 MATLAB/batch_bs_testing.m

@@ -1,0 +1,48 @@
+% Batch testing script for BS algorithms
+
+% Add toolboxes to path
+addpath('../toolbox/user');
+addpath('../toolbox/lightspeed');
+addpath('../toolbox/ekfukf');
+
+% Clean up
+clup
+
+% Get environment variable specifying test number
+test_num = str2double(getenv('SGE_TASK_ID'));
+if isnan(test_num)
+    test_num = 0;
+end
+fprintf(1, 'Test number: %u.\n', test_num);
+
+% Set flags
+test.flag_batch = true;
+
+% Define random seed
+rand_seed = test_num;
+
+% Set display options
+display.text = true;
+display.plot_during = false;
+display.plot_after = false;
+
+% Tests to run
+test.name = 'model_case_2'
+test.samp_types = [0 1 2 2 2 2 2 3 3 3 3 3 4 4 4 4 4];
+test.chain_lengths = [NaN NaN 3 10 32 100 320 3 10 32 100 320 3 10 32 100 320];
+  
+% Run the script
+run_bs_testing;
+
+% Get the numbers we want
+results.rt =   rt;
+results.pos_rmse = pos_rmse;
+results.vel_rmse = vel_rmse;
+results.tnees = tnees;
+results.nus = nus;
+
+% Save
+save(['batch_tests/' test.name '/bs_test_' num2str(test_num) '.mat'], 'results', 'model', 'test')
+
+% SHOW WE'VE FINISHED
+disp(['Test:' num2str(test_num) ' - DONE!']);

### 2013-11-28 MATLAB/ge_script.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+#$ -S /bin/bash
+
+/usr/local/apps/matlab/matlabR2012a/bin/matlab -nodisplay -r "batch_bs_testing; exit"

### 2013-11-28 MATLAB/loggausspdf.m

@@ -1,0 +1,59 @@
+function f = loggausspdf(X, Mu, Sigma)
+%MVNPDF Calculates the log of a Gaussian pdf
+%
+%   Y = MVNPDF(X,MU,SIGMA) returns a vector of Gaussian density
+%   values. X is a DxN array of points. Mu is a DxN array of means. Sigma
+%   is a DxDxN array of covariance matrices or a single DxD covariance marix.
+%
+%   Based on mvnpdf from the stats toolbox with all the error checking and
+%   special cases stripped out. Also, the vectors are the right way round.
+
+% Constant
+log2pi = 1.83787706640935;
+
+% Get size of data
+[d,n] = size(X);
+
+if n == 1
+    
+    X0 = X - Mu;
+    R = chol(Sigma);
+    logSqrtDetSigma = sum(log(diag(R)));
+    xRinv = R' \ X0;
+    
+else
+    
+    % Center data
+    X0 = bsxfun(@minus, X, Mu);
+    
+    % Cope with only one covariance matrix
+    if size(Sigma, 3) == 1
+        sameCov = true;
+        R = chol(Sigma);
+        logSqrtDetSigma = sum(log(diag(R)));
+    else
+        sameCov = false;
+        % Create arrays
+        xRinv = zeros(d,n,superiorfloat(X0,Sigma));
+        logSqrtDetSigma = zeros(1,n,class(Sigma));
+    end
+    
+    % Loop through points
+    for i = 1:n
+        
+        if ~sameCov
+            R = chol(Sigma(:,:,i));
+            logSqrtDetSigma(i) = sum(log(diag(R)));
+        end
+        
+        % Transform to standard normal
+        xRinv(:,i) = R' \ X0(:,i);
+        
+    end
+    
+end
+
+% The quadratic form is the inner products of the standardized data
+quadform = sum(xRinv.^2, 1);
+
+f = (-0.5*quadform - logSqrtDetSigma - d*log2pi/2)';

### 2013-11-28 MATLAB/normalise_weights.m

@@ -1,0 +1,8 @@
+function [ weight ] = normalise_weights( weight )
+%NORMALISE_WEIGHTS Normalise log-weights
+
+weight = weight - max(weight);
+weight = weight - logsumexp(weight,2);
+
+end
+

### 2013-11-28 MATLAB/original_version/NEES.m

@@ -1,0 +1,30 @@
+function [ nees, nees_over_time ] = NEES( x, pts )
+%NEES Calculate the normalised-estimation-error-squared between the mean
+% particle estimate and the true values
+
+[d, K] = size(x);
+
+Np = size(pts, 1);
+
+pts_x = cell2mat(permute(pts, [3,2,1]));
+est_x = mean(pts_x, 3);
+est_P = zeros(d,d,K);
+error = zeros(1,K);
+for kk = 1:K
+%     est_P(:,:,kk) = cov(squeeze(pts_x(:,kk,:))');
+    innov = bsxfun(@minus, squeeze(pts_x(:,kk,:)), x(:,kk));
+    est_P(:,:,kk) = (1/Np)*innov*(innov');
+    if det(est_P(:,:,kk))<1E-10  %all(all(~diff(innov')))  %
+        error(kk) = 1;
+    else
+        error(kk) = ((est_x(:,kk)-x(:,kk))'/est_P(:,:,kk))*(est_x(:,kk)-x(:,kk));
+    end
+end
+
+nees = sum(error)/K;
+nees_over_time = error;
+
+assert(~isnan(nees));
+
+end
+

### 2013-11-28 MATLAB/original_version/backward_rejectionsampling_smoother.m

@@ -1,0 +1,88 @@
+function [ smooth_pts ] = backward_rejectionsampling_smoother( S, times, pts_array, wts_array, h_trans )
+%BACKARD_SAMPLING_SMOOTHER Godsill2004-type backward sampling smoother
+%using rejection sampling
+%
+%
+%
+global params; %UGLY
+
+fprintf(1, '\n\n*** Running rejection sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Get weights
+        wts = wts_array{kk};
+        
+        % Rejection sample
+        [~, Q] = build_AQ(dt, params.proc_var);  % This is ugly and should be encapsulated.
+        max_trans_prb = -0.5*log(det(2*pi*Q));
+        u = inf; thresh = 1;
+        count = 0;
+        while (u > thresh)
+            
+            % Sample an index and a uniform variable
+            ind = sample_weights(exp(wts));
+            u = log(rand);
+            
+            % Calculate the acceptance threshold
+            [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{ind}(:,kk), smooth_pts{ii}(:,kk+1) );
+            thresh = trans_prb - max_trans_prb;
+            
+            assert(thresh < 0)
+            
+            count = count + 1;
+            
+            if count == params.M
+                % Do it normally
+%                 fprintf(1,'No sample - direct evaluation.');
+                
+                for jj = 1:length(pts_array{kk})
+                    % Calculate weight
+                    [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{jj}(:,kk), smooth_pts{ii}(:,kk+1) );
+                    wts(jj) = wts(jj) + trans_prb;
+                end
+                
+                % Normalise weights
+                wts = wts - max(wts);
+                lin_wts = exp(wts); lin_wts = lin_wts/sum(lin_wts);
+                wts = log(lin_wts);
+                
[…]

### 2013-11-28 MATLAB/original_version/backward_sampling_smoother.m

@@ -1,0 +1,61 @@
+function [ smooth_pts ] = backward_sampling_smoother( S, times, pts_array, wts_array, h_trans )
+%BACKARD_SAMPLING_SMOOTHER Normal Godsill2004-type backward sampling
+%smoother.
+%
+%
+%
+
+fprintf(1, '\n\n*** Running direct backward sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Get weights
+        wts = wts_array{kk};
+        
+        % Loop through filter particles
+        for jj = 1:length(pts_array{kk})
+            
+            % Calculate weight
+            [~, trans_prb] = feval(h_trans, dt, pts_array{kk}{jj}(:,kk), smooth_pts{ii}(:,kk+1) );
+            wts(jj) = wts(jj) + trans_prb;
+            
+        end
+        
+        % Normalise weights
+        wts = wts - max(wts);
+        lin_wts = exp(wts); lin_wts = lin_wts/sum(lin_wts);
+        wts = log(lin_wts);
+        
+        % Sample
+%         fprintf(1, '     ESS in frame %u: %f\n', kk, ESS(wts));
+        ind = sample_weights(exp(wts));
+        
+        % Update trajectory
+        smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+        
+    end
+    
+end
+
+end
+

### 2013-11-28 MATLAB/original_version/batch_analysis.m

@@ -1,0 +1,59 @@
+%% Preliminaries
+
+close all
+
+% Set test flag
+test_flag = 3;
+
+% Load data
+load(['smoother_test' num2str(test_flag) '.mat']);
+
+% 
+NA = 13;
+t = params.dt:params.dt:params.dt*params.N;
+col = {[1 0 0], [1 .5 .5], [0 1 0], [.5 1 .5], [0 0 1], [.5 .5 1]};
+
+%% Output results
+
+% Unique particles over time
+figure(1), hold on
+to_plot = [2,3,4,6,8,10];
+for ii = 1:length(to_plot)
+    alg = to_plot(ii);
+    plot(t, results{alg}.unique_pts, 'color', col{ii});
+end
+legend('filter-smoother', 'direct resampling smoother', 'MCMC resampling smoother (M=1)', 'MCMC resampling smoother (M=10)', 'MCMC sampling smoother (M=1)', 'MCMC sampling smoother (M=10)');
+xlabel('time'), ylabel('num. of unique particles')
+
+figure(2), hold on
+bar( cellfun(@(x) mean(x.unique_pts), results) );
+
+figure(3), hold on
+bar( cellfun(@(x) x.mean_pos_rmse, results) );
+
+figure(4), hold on
+bar( cellfun(@(x) x.mean_vel_rmse, results) );
+
+figure(5), hold on
+bar( cellfun(@(x) x.times, results) );
+
+figure(6), hold on
+inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*k' )
+inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*r' )
+inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-b' )
+inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-g' )
+legend('FS', 'DBRS', 'MCMC-BRS (M=1-100)', 'MCMC-BSS (M=1-100)');
+xlabel('Running Time'); ylabel('Position RMSE');
[…]

### 2013-11-28 MATLAB/original_version/batch_smoother_testing.m

@@ -1,0 +1,162 @@
+clup
+dbstop if error
+
+% Get test flag
+% test_flag = str2double(getenv('SGE_TASK_ID'));
+test_flag = 3;
+
+% Batch test parameters
+num_seed = 100;
+set_batch_parameters;
+
+% Set test-specific parameters
+if test_flag == 1
+    params.bng_var = (pi/720)^2;
+    params.rng_var = 0.1;
+elseif test_flag == 2
+    params.bng_var = (pi/36)^2;
+    params.rng_var = 0.1;
+elseif test_flag == 3
+    params.bng_var = (pi/36)^2;
+    params.rng_var = 100;
+end
+
+% Number of smoothing algorithms
+NA = 13; % filter, kitigawa, godsill, 5 MCMC-resampling, 5 MCMC-Fearnhead
+
+% Results arrays
+times = cell(NA,1);
+unique_pts = cell(NA,1);
+mean_pos_rmse = cell(NA,1);
+mean_vel_rmse = cell(NA,1);
+mean_nees = cell(NA,1);
+for alg = 1:NA
+    times{alg} = zeros(num_seed,1);
+    unique_pts{alg} = zeros(num_seed,params.N);
+    mean_pos_rmse{alg} = zeros(num_seed,1);
+    mean_vel_rmse{alg} = zeros(num_seed,1);
+    mean_nees{alg} = zeros(num_seed,1);
+end
+
+% Loop through random seed
+for rs = 1:num_seed;
+    
+    rs
+    
+    % Set random seed
+    s = RandStream('mt19937ar', 'seed', rs);
+    RandStream.setDefaultStream(s);
+    
+    %% Generate some Bearings only tracking data
+    [ t, x, y ] = generate_radar_data;
+    
+    %% Run a PF
+    init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);
+    [pts_array, wts_array, filter_pts] = particle_filter( init_pts, t, y, @tracking_ppsl, @tracking_trans, @tracking_obs, 0.5 );
+    
+    % Resample final frame particles
+    parents = systematic_resample(wts_array{end});
+    kiti_pts = pts_array{end}(parents);
+    
+    % Analyse
+    filt_rmse = RMSE(x, filter_pts);
+    
+    kiti_rmse = RMSE(x, kiti_pts);
[…]

### 2013-11-28 MATLAB/original_version/batch_smoother_testing2.m

@@ -1,0 +1,160 @@
+clup
+dbstop if error
+
+% Get test flag
+% test_flag = str2double(getenv('SGE_TASK_ID'));
+test_flag = 3;
+
+% Batch test parameters
+num_seed = 10;
+set_parameters;
+
+% Set test-specific parameters
+if test_flag == 1
+    params.bng_var = (pi/720)^2;
+    params.rng_var = 0.1;
+elseif test_flag == 2
+    params.bng_var = (pi/180)^2;
+    params.rng_var = 10;
+elseif test_flag == 3
+    params.bng_var = (pi/36)^2;
+    params.rng_var = 100;
+end
+
+% Number of smoothing algorithms
+NA = 13; % filter, kitigawa, godsill, 5 MCMC-resampling, 5 MCMC-Fearnhead
+
+% Results arrays
+times = cell(NA,1);
+unique_pts = cell(NA,1);
+mean_pos_rmse = cell(NA,1);
+mean_vel_rmse = cell(NA,1);
+for alg = 1:NA
+    times{alg} = zeros(num_seed,1);
+    unique_pts{alg} = zeros(num_seed,params.N);
+    mean_pos_rmse{alg} = zeros(num_seed,1);
+    mean_vel_rmse{alg} = zeros(num_seed,1);
+end
+
+% Loop through random seend
+for rs = 1:num_seed;
+    
+    rs
+    
+    % Set random seed so we always generate the same data
+    s = RandStream('mt19937ar', 'seed', 1);
+    RandStream.setDefaultStream(s);
+    
+    % Set parameters
+    set_parameters;
+    
+    %% Generate some Bearings only tracking data
+    [ t, x, y ] = generate_radar_data;
+    
+    % Set random seed for running algorithms
+    s = RandStream('mt19937ar', 'seed', rs);
+    RandStream.setDefaultStream(s);
+    
+    %% Run a PF
+    init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);
+    [pts_array, wts_array, filter_pts] = particle_filter( init_pts, t, y, @tracking_ppsl, @tracking_trans, @tracking_obs, 0.5 );
+    
+    % Resample final frame particles
+    parents = systematic_resample(wts_array{end});
+    kiti_pts = pts_array{end}(parents);
+    
+    % Analyse
[…]

### 2013-11-28 MATLAB/original_version/count_unique_particles.m

@@ -1,0 +1,92 @@
+function [ Nup, Nuh ] = count_unique_particles( pts )
+%COUNT_UNIQUE_PARTICLES Counts number of unique particles (at time k) and
+%histories (up to time k)
+
+% Initialise variables
+Np = length(pts);
+[d, K] = size(pts{1});
+
+% Initialise arrays
+Nup = Np*ones(1,K);
+Nuh = Np*ones(1,K);
+
+fprintf(1, '\nCounting unique particles.\n');
+
+% Loop through time
+for k = 1:K
+    
+    if rem(k, 100)==0
+        fprintf(1, '     Reached time step %u.\n', k);
+    end
+    
+    %%%% Particles
+    
+    unique = true(Np,1);
+    
+    % Loop through particles
+    for ii = 1:Np
+        
+        x_ii = pts{ii}(:,k);
+        
+        % Loop through later particles
+        for jj = ii+1:Np
+            
+            % Only look at particles which we still know to be unique
+            if unique(jj)
+                
+                x_jj = pts{jj}(:,k);
+                
+                % Compare number of jumps
+                if all(all(x_ii==x_jj))
+                    
+                    % Particle is not unique, cross it off
+                    unique(jj) = false;
+                    Nup(k) = Nup(k) - 1;
+                    
+                end
+                
+            end
+            
+        end
+        
+    end
+    
+    %%%% Histories
+    
+    unique = true(Np,1);
+    
+    % Loop through particles
+    for ii = 1:Np
+        
+        x_ii = pts{ii}(:,1:k);
+        
+        % Loop through later particles
+        for jj = ii+1:Np
+            
+            % Only look at particles which we still know to be unique
+            if unique(jj)
+                
+                x_jj = pts{jj}(:,1:k);
+                
+                % Compare number of jumps
+                if all(all(x_ii==x_jj))
+                    
+                    % Particle is not unique, cross it off
+                    unique(jj) = false;
+                    Nuh(k) = Nuh(k) - 1;
+                    
+                end
+                
+            end
+            
+        end
+        
+    end
+    
+end
+
+fprintf(1, '     Average number of unique particles: %3.1f.\n', mean(Nup));
+fprintf(1, '     Number of unique trajectories:      %u.\n', Nuh(end));
+
+end
+

### 2013-11-28 MATLAB/original_version/fast_log_mvnpdf.m

@@ -1,0 +1,19 @@
+function y = fast_log_mvnpdf(X, Mu, Sigma)
+%FAST_LOG_MVNPDF Runs fast, no-error-checking version of log(mvnpdf()) for only one data
+%point
+
+% Get size of data.  Column vectors provisionally interpreted as multiple scalar data.
+[n,d] = size(X);
+
+% Set zero mean
+X0 = X - Mu;
+
+% Decompose variance
+R = chol(Sigma);
+xRinv = X0 / R;
+logSqrtDetSigma = sum(log(diag(R)));
+
+% The quadratic form is the inner products of the standardized data
+quadform = sum(xRinv.^2, 2);
+
+y = -0.5*quadform - logSqrtDetSigma - d*log(2*pi)/2;

### 2013-11-28 MATLAB/original_version/generate_radar_data.m

@@ -1,0 +1,33 @@
+function [ t, x, y ] = generate_radar_data
+%GENERATE_RADAR_DATA Generates some radar-style tracking data for inference
+%algorithm testing
+
+global params;
+
+[A, Q] = build_AQ(params.dt, params.proc_var);
+R = [params.bng_var 0; 0 params.rng_var];
+N = params.N;
+x0 = params.x0;
+
+% Initialise arrays
+x = zeros(4, N);
+y = zeros(2, N);
+t = params.dt:params.dt:params.dt*N;
+
+last_x = x0;
+
+% Loop
+for n = 1:N
+    
+    % Simulate new state
+    x(:,n) = mvnrnd((A*last_x)', Q)';
+    last_x = x(:,n);
+    
+    % Simulate new observation
+    [b, r] = cart2pol(x(1,n), x(2,n));
+    y(:,n) = mvnrnd([b, r], R)';
+    
+end
+
+end
+

### 2013-11-28 MATLAB/original_version/mcmc_newstate_smoother.m

@@ -1,0 +1,86 @@
+function [ smooth_pts ] = mcmc_newstate_smoother( S, M, times, pts_array, wts_array, observs, h_trans, h_obs, h_bidirec_ppsl )
+%MCMC_NEWSTATE_SMOOTHER 
+%
+%
+%
+
+fprintf(1, '\n\n*** Running MCMC-based backward new-state sampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+Np = size(pts_array{K},1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:2
+        
+        dt_next = times(kk+1) - times(kk);
+        dt_prev = times(kk) - times(kk-1);
+        
+        old_x = smooth_pts{ii}(:,kk);
+        
+        % Initialise probability
+        [~, old_trans_prb_prev] = feval(h_trans, dt_prev, smooth_pts{ii}(:,kk-1), old_x );
+        [~, old_trans_prb_next] = feval(h_trans, dt_next, old_x, smooth_pts{ii}(:,kk+1) );
+        [~, old_obs_prb] = feval(h_obs, smooth_pts{ii}(:,kk), observs(:,kk));
+        [~, old_ppsl_prb] = feval(h_bidirec_ppsl, smooth_pts{ii}(:,kk-1), smooth_pts{ii}(:,kk+1), observs(:,kk), dt_prev, dt_next, h_trans, h_obs, smooth_pts{ii}(:,kk));
+        
+%         ind = 1;
+%         while smooth_pts{ii}(1,kk-1)~=pts_array{kk-1}{ind}(1,kk-1)
+%             ind = ind + 1;
+%         end
+%         old_weight = wts_array{kk-1}(ind);
+        
+        for jj = 1:M
+            
+            % Propose a new k-1 particle
+%             ind = randsample(Np, 1);
+            ind = sample_weights(exp(wts_array{kk-1}));
+            
+            % Propose a new k state
[…]

### 2013-11-28 MATLAB/original_version/mcmc_smoother.m

@@ -1,0 +1,68 @@
+function [ smooth_pts ] = mcmc_smoother( S, M, times, pts_array, wts_array, h_trans )
+%MCMC_SMOOTHER New MCMC-based backward sampling smoother
+%
+%
+%
+
+fprintf(1, '\n\n*** Running MCMC-based backward resampling smoother. ***\n');
+
+% Initialise variables
+K = size(pts_array,1);
+Np = size(pts_array{K},1);
+
+% Initialise arrays
+smooth_pts = cell(S,1);
+
+% Sample the final filter weights
+parents = systematic_resample(wts_array{K}, S);
+
+% Loop through trajectories
+for ii = 1:S
+        
+    fprintf(1, 'Now simulating trajectory %u.\n', ii);
+    
+    % Copy filtering particle
+    smooth_pts{ii} = pts_array{K}{parents(ii)};
+    
+    % Loop backwards in time
+    for kk = K-1:-1:1
+        
+        dt = times(kk+1) - times(kk);
+        
+        % Initialise probability
+        [~, old_trans_prb] = feval(h_trans, dt, smooth_pts{ii}(:,kk), smooth_pts{ii}(:,kk+1) );
+%         ind = 1;
+%         while smooth_pts{ii}(1,kk)~=pts_array{kk}{ind}(1,kk)
+%             ind = ind + 1;
+%         end
+%         old_weight = wts_array{kk}(ind);
+        
+        for jj = 1:M
+            
+            % Propose a new particle
+%             ind = randsample(Np, 1);
+            ind = sample_weights(exp(wts_array{kk}));
+            
+            % Calculate probability
+            [~, new_trans_prb] = feval(h_trans, dt, pts_array{kk}{ind}(:,kk), smooth_pts{ii}(:,kk+1) );
+            new_weight = wts_array{kk}(ind);
+            
+            % Construct acceptance probability
+            ap = (new_trans_prb - old_trans_prb);%...
+%                 +(new_weight - old_weight);
+            
+            if log(rand)<ap
+                % Update trajectory
+                smooth_pts{ii} = [pts_array{kk}{ind}(:,1:kk) smooth_pts{ii}(:,kk+1:end)];
+                old_weight = new_weight;
+                old_trans_prb = new_trans_prb;
+            end
+            
+        end
+        
+    end
+    
+end
+
+end
+

### 2013-11-28 MATLAB/original_version/plot_banana.m

@@ -1,0 +1,36 @@
+function plot_banana( fig, xy_mean, rng_var, bng_var, c )
+%PLOT_BANANA Plot equal probability contour for a range-bearing observation model
+
+% Number of points to plot (must be divisible by 2)
+Npts = 2000;
+
+% Convert to polar
+[b_mean, r_mean] = cart2pol(xy_mean(1), xy_mean(2));
+
+% Find range of... range
+r_min = r_mean-c*sqrt(rng_var);
+r_max = r_mean+c*sqrt(rng_var);
+
+% Create an array of points
+rb_pts = zeros(2,Npts);
+
+% Grid of range points
+rb_pts(1,1:Npts/2) = (r_min:(r_max-r_min)/(Npts/2-1):r_max);
+rb_pts(1,Npts/2+1:end) = fliplr(r_min:(r_max-r_min)/(Npts/2-1):r_max);
+
+% Calculate corresponding bearing points
+rb_pts(2,1:Npts/2) = b_mean + sqrt(abs(bng_var*(c^2-(rb_pts(1,1:Npts/2)-r_mean).^2/rng_var)));
+rb_pts(2,Npts/2+1:end) = b_mean - sqrt(abs(bng_var*(c^2-(rb_pts(1,Npts/2+1:end)-r_mean).^2/rng_var)));
+
+% Convert back to Cartesian
+xy_pts = zeros(size(rb_pts));
+[xy_pts(1,:), xy_pts(2,:)] = pol2cart(rb_pts(2,:), rb_pts(1,:));
+
+assert(all(all(isreal(xy_pts))));
+
+% Plot
+figure(fig)
+plot(xy_pts(1,:), xy_pts(2,:), ':r')
+
+end
+

### 2013-11-28 MATLAB/original_version/plot_graphs_for_paper.m

@@ -1,0 +1,103 @@
+clup
+
+% Set test flag
+for test_flag = [1,2,3];
+    
+    test_case = test_flag;
+    
+    % Load data
+    load(['smoother_test' num2str(test_flag) '.mat']);
+    
+    % RMSEs
+    fig = figure; hold on
+    inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'ok', 'markersize', 7 )
+    inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'r^', 'markersize', 7 )
+    inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), '*-b', 'markersize', 7 )
+    inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_pos_rmse, results(inds)), 'x-.g', 'markersize', 7 )
+    
+    legend('FS', 'D-FFBS', 'MH-FFBS (M=1-100)', 'MH-FFBP (M=1-100)');
+    xlabel('Running Time (s)'); ylabel('Position RMSE');
+    xlimits = get(gca, 'XLim');
+    ylimits = get(gca, 'YLim');
+    xlim([-50, xlimits(2)]);
+    plot([0 0], ylimits, ':k');
+    ylim(ylimits);
+    
+    export_pdf(fig, ['case' num2str(test_case) '_smoother_comparison_posRMSE_time.pdf']);
+    
+    % MNEESs
+    fig = figure; hold on
+    inds = 2; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'ok', 'markersize', 7 )
+    inds = 3; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'r^', 'markersize', 7 )
+    inds = 4:8; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), '*-b', 'markersize', 7 )
+    inds = 9:13; plot( cellfun(@(x) x.times, results(inds)), cellfun(@(x) x.mean_nees, results(inds)), 'x-.g', 'markersize', 7 )
+    
[…]

### 2013-11-28 MATLAB/original_version/run_mcmcsmoother_demo.m

@@ -1,0 +1,161 @@
+clup
+dbstop if error
+
+% DEFINE RANDOM SEED
+rand_seed = 0;
+
+% Set random seed
+s = RandStream('mt19937ar', 'seed', rand_seed);
+RandStream.setDefaultStream(s);
+
+% Rejection sampling? Set to false - its too slow
+rej_samp = true;
+
+% Set parameters
+set_parameters;
+
+%% Generate some Bearings only tracking data
+[ t, x, y ] = generate_radar_data;
+
+% Plot it
+xmax = max(max(abs(x(1:2,:))))+5;
+figure(1); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3);
+[x_obs, y_obs] = pol2cart(y(1,:), y(2,:)); plot(x_obs, y_obs, 'r');
+plot(0, 0, 'xk', 'markersize', 10)
+legend('Target Position', 'Observations')
+% figure(2); plot(t, y);
+
+%% Run a PF
+init_pts = num2cell(mvnrnd(params.x0', params.prior_var, params.Np)', 1);
+[pts_array, wts_array, filter_pts] = particle_filter( init_pts, t, y, @tracking_ppsl, @tracking_trans, @tracking_obs, 0.5 );
+
+% Resample final frame particles
+parents = systematic_resample(wts_array{end});
+kiti_pts = pts_array{end}(parents);
+
+% Plot filter
+figure(3); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), filter_pts)
+
+% Plot Kitigawa smoother
+figure(4); hold on; xlim([-xmax, xmax]); ylim([-xmax, xmax]);
+plot(x(1,:), x(2,:), 'b', 'linewidth', 3)
+cellfun(@(x) plot(x(1,:), x(2,:), 'color', [rand rand rand]), kiti_pts)
+
+%% Run an ordinary smoother
+tic;
+bs_smooth_pts = backward_sampling_smoother( params.S, t, pts_array, wts_array, @tracking_trans );
+bs_time = toc;
+
[…]

### 2013-11-28 MATLAB/original_version/sample_weights.m

@@ -1,0 +1,14 @@
+function idx = sample_weights(w)
+%SAMPLE_WEGHTS Sample an index from a column vector of weights
+
+% Normalise weights
+sumw = sum(w);
+if sumw ~= 1
+    w = w./sumw;
+end
+
+% Sample
+edges = [0; cumsum(w)];
+[~, idx] = histc(rand,edges);
+
+end

### 2013-11-28 MATLAB/original_version/set_batch_parameters.m

@@ -1,0 +1,17 @@
+global params;
+
+%Model
+params.dt = 1;
+params.N = 100;
+params.proc_var = 1;
+params.x0 = [-100; 50; 10; 0];
+% params.bng_var    Set for each test
+% params.rng_var    Set for each test
+
+% Prior
+params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
+
+% Algorithm
+params.Np = 100;
+params.S = 100;
+% params.M          Set of each test

### 2013-11-28 MATLAB/original_version/set_parameters.m

@@ -1,0 +1,17 @@
+global params;
+
+%Model
+params.dt = 1;
+params.N = 100;
+params.proc_var = 1;
+params.bng_var = (pi/36)^2;    % Try (pi/720)^2; (pi/180)^2;
+params.rng_var = 0.1;           % Try 0.1;        10;
+params.x0 = [-100; 50; 10; 0];
+
+% Prior
+params.prior_var = diag([0.0005; 0.0005; 0.001; 0.001]);
+
+% Algorithm
+params.Np = 1000;
+params.S = 100;
+params.M = 10;                   % Try 5;           50;

### 2013-11-28 MATLAB/original_version/test_functions.m

@@ -1,0 +1,17 @@
+function [ tf ] = test_functions( pts )
+%TEST_FUNCTIONS Calculate the mean of an assortment of test functions using
+%particle filter approximations to the joint distributions
+
+tf = cell(10,1);
+
+for ii = 1:length(pts)
+    
+    tf{1} = sum(pts{ii}(1:2,:).^2, 1);
+    tf{2} = sum(pts{ii}(3:4,:).^2, 1);
+    tf{3} = sum(pts{ii}(1:2,1:end-1).*pts{ii}(1:2,2:end), 1);
+    tf{4} = sum(pts{ii}(3:4,1:end-1).*pts{ii}(3:4,2:end), 1);
+    
+end
+
+end
+

### 2013-11-28 MATLAB/original_version/tracking_bidirec_ppsl.m

@@ -1,0 +1,26 @@
+function [x, ppsl_prb] = tracking_bidirec_ppsl(x_prev, x_next, y, dt_prev, dt_next, h_trans, h_obs, x)
+%TRACKING_BIDIREC_PPSL Propose a new state for a 2D tracking model when the
+% next state is already known (for smoother).
+
+global params;
+
+[A_prev, Q_prev] = build_AQ(dt_prev, params.proc_var);
+[A_next, Q_next] = build_AQ(dt_next, params.proc_var);
+R = [params.bng_var 0; 0 params.rng_var];
+
+% Interpolate between previous and next x
+[m, P] = kf_update(A_prev*x_prev, Q_prev, x_next, A_next, Q_next);
+
+% Now include y
+H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
+     m(1)/sqrt(m(1)^2+m(2)^2), m(2)/sqrt(m(1)^2+m(2)^2), 0, 0];
+[m, P] = ekf_update1(m, P, y, H, R, h_obs);
+
+% Sample
+if (nargin < 8)||isempty(x)
+    x = mvnrnd(m', P)';
+end
+ppsl_prb = fast_log_mvnpdf(x', m', P);
+
+end
+

### 2013-11-28 MATLAB/original_version/tracking_obs.m

@@ -1,0 +1,16 @@
+function [ y_mn, obs_prb ] = tracking_obs( x, y )
+%TRACKING_OBS Observation likelihood for bearing-only 2D radar observations
+
+global params;
+
+R = [params.bng_var 0; 0 params.rng_var];
+
+[b, r] = cart2pol(x(1), x(2));
+y_mn = [b; r];
+
+if ~isempty(y)
+    obs_prb = fast_log_mvnpdf(y', y_mn', R);
+end
+
+end
+

### 2013-11-28 MATLAB/original_version/tracking_ppsl.m

@@ -1,0 +1,22 @@
+function [ x, ppsl_prb ] = tracking_ppsl( last_x, y, dt, h_trans, h_obs)
+%TRACKING_PPSL Propose a new state for a 2D tracking model
+
+global params;
+
+[A, Q] = build_AQ(dt, params.proc_var);
+R = [params.bng_var 0; 0 params.rng_var];
+
+% Optimal importance
+[m, P] = kf_predict(last_x, eps*eye(4), A, Q);
+H = [-m(2)/(m(1)^2+m(2)^2),    m(1)/(m(1)^2+m(2)^2),     0, 0;
+     m(1)/sqrt(m(1)^2+m(2)^2), m(2)/sqrt(m(1)^2+m(2)^2), 0, 0];
+[m, P] = ekf_update1(m, P, y, H, R, h_obs);
+x = mvnrnd(m', P)';
+ppsl_prb = fast_log_mvnpdf(x', m', P);
+
+% % Bootstrap
+% x = mvnrnd((A*last_x), Q)';
+% ppsl_prb = log(mvnpdf(x', (A*last_x)', Q));
+
+end
+

### 2013-11-28 MATLAB/original_version/tracking_trans.m

@@ -1,0 +1,15 @@
+function [ x_mn, trans_prb ] = tracking_trans( dt, last_x, x )
+%TRACKING_TRANS Transition probability for linear 2D tracking dynamics
+
+global params;
+
+[A, Q] = build_AQ(dt, params.proc_var);
+
+x_mn = A*last_x;
+
+if ~isempty(x)
+    trans_prb = fast_log_mvnpdf(x', x_mn', Q);
+end
+
+end
+

### 2013-11-28 MATLAB/particle_filter.m

@@ -1,106 +1,156 @@
-function [ pts_array, wts_array, filter_pts ] = particle_filter( init_pts, times, observs, h_ppsl, h_trans, h_obs, ess_thresh )
-%PARTICLE_FILTER Run a particle filter
-%
-% *** Input ***
-% init_pts      - Cell array of particles defining prior (see below)
-% times         - Row vector of observation times
-% observs       - Matrix of observations, observs(:,k) is the kth observation
-% h_ppsl        - Function handle to the proposal function.
-% h_trans       - Function handle to the transition likelihood function.
-% h_obs         - Function handle to the observation likelihood function.
-%
-% *** Output ***
-% pts_array     - Cell array of particle cell arrays, the filter output at each observation time
-% wts_array     - Cell array of corresponding weight column vectors
-% filter_pts    - Cell array of equal-weight, current filter points (i.e. no Kitigawa smoothing)
-%
-% Particle cell arrays have the following structure: pts{ii}(:,k) is the
-% state of the iith particle at the kth time.
-%
-% All weights and probabilities are log-ed.
-%
+function [ pf, diagnostics ] = particle_filter( display, algo, model, fh, observ, flag_ppsl_type, true_state )
+%particle_filter
 
-fprintf(1, '\n\n*** Running particle filter. ***\n');
+if display.text
+    fprintf(1, 'Running particle filter, type %u.\n', flag_ppsl_type);
+end
 
-% Initialise constants
-Np = length(init_pts);
-K = size(times, 2);
-d = size(init_pts{1}, 1);
+% Initialise diagnostics structure
+diagnostics = repmat(struct('rt', [], 'ess', [], 'se', []), 1, model.K);
 
-% Initialise arrays
-pts_array = cell(K,1);
-wts_array = cell(K,1);
-filter_pts = repmat({zeros(d,K)}, Np, 1);
-wts = log(ones(Np,1)/Np);
-last_pts = init_pts;
+% Initialise particle filter structure
+pf = repmat(struct('state', [], 'ancestor', [], 'weight', []), 1, model.K);
 
-% Loop through time
-for kk = 1:K
+%%% Time loop %%%
+for kk = 1:model.K
     
-    fprintf(1, 'Now processing frame %u.\n', kk);
-    
-    % Create a new particle array
-    pts = repmat({zeros(d,kk)}, Np, 1);
-    
-    % Find time increment
-    t = times(kk);
-    if kk == 1
-        dt = t;
-    else
[…]

### 2013-11-28 MATLAB/particle_smoother_analysis.m

@@ -1,0 +1,50 @@
+function [ pos_rmse, vel_rmse, tnees, nus ] = particle_smoother_analysis( model, true_state, ps )
+%PARTICLE_SMOOTHER_ANALYSIS Calculate RMSE, TNEES and NUS over time for
+%a particle smoother approximation
+
+K = model.K;
+S = length(ps);
+
+% Initialise arrays
+nees = zeros(1,K);
+nus = zeros(1,K);
+
+% Cut up state
+true_pos = true_state(1:model.ds/2,:);
+true_vel = true_state(model.ds/2+1:model.ds,:);
+
+% Make a great big array of all particle states
+state = cat(3,ps.traj_state);
+
+% Take the mean to give us a point estimate
+state_est = mean(state, 3);
+pos_est = state_est(1:model.ds/2,:);
+vel_est = state_est(model.ds/2+1:model.ds,:);
+
+% RMSEs
+pos_rmse = sqrt(sum( (true_pos-pos_est).^2, 1 ));
+vel_rmse = sqrt(sum( (true_vel-vel_est).^2, 1 ));
+
+% NUS
+for kk = 1:K
+    nus(kk) = size(unique(squeeze(state(:,kk,:))', 'rows'),1);
+end
+
+% TNEES
+for kk = 1:K
+    if nus(kk)>model.ds
+        state_diff = squeeze(bsxfun(@minus, state(:,kk,:), state_est(:,kk)));
+        vr_est = state_diff*state_diff'/S;
+        err = state_est(:,kk) - true_state(:,kk);
+        nees(kk) = err'*(vr_est\err);
+    else
+        nees(kk) = Inf;
+    end
+end
+tnees = nees./(1+nees);
+tnees(isnan(tnees))=1;
+
+
+
+end
+

### 2013-11-28 MATLAB/refreshed_backward_simulation_smoother.m

@@ -1,0 +1,150 @@
+function [ ps, diagnostics ] = refreshed_backward_simulation_smoother( display, algo, model, fh, pf, observ, true_state )
+%BACKWARD_SIMULATION_SMOOTHER Run refreshed backward simulation smoothing
+%using MH sampling.
+
+if display.text
+    fprintf(1, 'Running refreshed backward simulation smoother.\n');
+end
+
+% Local copies
+K = model.K;
+N = algo.N;
+S = algo.S;
+M = algo.M;
+
+% Set up diagnostics
+diagnostics.acceptance_rate = zeros(S,K);
+
+% Resample final frame filter particles
+filt_ancestor = sample_weights(pf(K).weight, S);
+
+% Particle loop
+for ii = 1:S
+    
+    if display.text
+        fprintf(1, '   Simulating trajectory %u.\n', ii);
+    end
+    
+    % Initialise new particle index and state trajectory
+    traj_ind = zeros(1,K);
+    traj_state = zeros(model.ds,K);
+    traj_ind(K) = filt_ancestor(ii);
+    
+    % Loop backwards in time
+    for kk = K:-1:1
+        
+        if kk < K
+            next_state = traj_state(:,kk+1);
+        else
+            next_state = [];
+        end
+        
+        % Initialise
+        num_acceptances = 0;
+        if kk > 1
+            prev_ind = pf(kk).ancestor(traj_ind(kk));
+        else
+            prev_ind = 0;
+        end
+        state = pf(kk).state(:,traj_ind(kk));
+        
+        if kk > 1
+            [import_mn, import_vr] = double_kf_update(model, pf(kk-1).state(:,prev_ind), next_state, observ(:,kk));
+        else
+            [import_mn, import_vr] = double_kf_update(model, [], next_state, observ(:,kk));
+        end
+        import = loggausspdf(state, import_mn, import_vr);
+        
+        if kk < K
+            [~, trans1] = tracking_transition(model, state, next_state);
+        else
+            trans1 = 0;
+        end
+        if kk > 1
+            [~, trans2] = tracking_transition(model, pf(kk-1).state(:,prev_ind), state);
+        else
+            [~, trans2] = tracking_stateprior(model, state);
+        end
+        [ ~, lhood ] = tracking_observation( model, state, observ(:,kk) );
+        
+        % Chain
+        for jj = 1:M
+            
[…]

### 2013-11-28 MATLAB/run_bs_testing.m

@@ -1,0 +1,145 @@
+% Base script for backward simulation smoothing comparisons on a tracking model
+
+%% Preliminaries
+
+if ~exist('test', 'var') || ~isfield(test,'flag_batch') || (~test.flag_batch)
+    
+    clup
+    dbstop if error
+    %     dbstop if warning
+    
+    % Set flag to non-batch
+    test.flag_batch = false;
+    
+    %%% SETTINGS %%%
+    
+    % DEFINE RANDOM SEED
+    rand_seed = 0;
+    
+    % Set display options
+    display.text = true;
+    display.plot_during = false;
+    display.plot_after = true;
+    
+    % Tests to run
+    test.samp_types = [0 1 2 3 4];
+    test.chain_lengths = [NaN NaN 10 10 10];
+    
+end
+
+%% Setup
+fprintf('   Random seed: %u.\n', rand_seed);
+
+% Function handles for the model
+addpath('tracking2');
+fh.setmodel = @tracking_setmodel;
+fh.setalgo = @tracking_setalgo;
+fh.generatedata = @tracking_generatedata;
+fh.transition = @tracking_transition;
+fh.observation = @tracking_observation;
+fh.stateprior = @tracking_stateprior;
+fh.ekfproposal = @tracking_ekfproposal;
+fh.ukfproposal = @tracking_ukfproposal;
+
+% Set random seed
+rng(rand_seed);
+
+% Set model parameters
+[model] = feval(fh.setmodel, test);
+
+% Set algorithm parameters
+[algo] = feval(fh.setalgo, test);
+
+% Generate data
+[time, state, observ] = feval(fh.generatedata, model);
+
+%% Filtering
+
+% Run the particle filter
+ppsl_type = 3; % 1,2,3 = bootstrap,EKF,UKF resp.
+[pf, pf_diagnostics] = particle_filter(display, algo, model, fh, observ, ppsl_type, state);
+
+%% Smoothing
+
+num_alg = length(test.samp_types);
+ps = cell(num_alg,1);
+ps_diagnostics = cell(num_alg,1);
+rt = zeros(num_alg,1);
+
+% Run the smoothers
+for aa = 1:num_alg
+    type = test.samp_types(aa);
+    algo.M = test.chain_lengths(aa);
+    tic;
+    if type == 4
[…]

### 2013-11-28 MATLAB/sample_weights.m

@@ -1,14 +1,48 @@
-function idx = sample_weights(w)
-%SAMPLE_WEGHTS Sample an index from a column vector of weights
+function [ ancestor, selected_weight ] = sample_weights( weight, N, resam_type )
+%SAMPLE_WEIGHTS Samples an array of N ancestor indexes from a set of
+%weights. Also (optionally), the weights of the ancesors.
 
-% Normalise weights
-sumw = sum(w);
-if sumw ~= 1
-    w = w./sumw;
+% Default resampling type
+if (nargin<3)||isempty(resam_type)
+    resam_type = 1;
 end
 
-% Sample
-edges = [0; cumsum(w)];
-[~, idx] = histc(rand,edges);
+% Row vectors only here, please
+weight = weight(:)';
+
+% Convert weights to linear domain and normalise
+weight = weight - max(weight);
+weight = exp(weight);
+weight = weight/sum(weight);
+
+% Catch NaNs
+assert(~any(isnan(weight)));
+
+% Create bin boundaries
+edges = min([0 cumsum(weight)],1);
+edges(end) = 1;
+
+% Sample indexes
+if resam_type == 1
+    % Multinomial
+    idx = rand(N,1);
+elseif resam_type == 2
+    % Systematic
+    idx = (1/N)*ones(N, 1);
+    idx(1) = rand/N;
+    idx = cumsum(idx);
+else
+    error('That''s not a valid choice for resampling type');
+end
+
+% Draw particles
+[~, ancestor] = histc(idx,edges);
+ancestor = ancestor';
+
+% Selected weights
+if nargout > 1
+    selected_weight = log(weight(ancestor));
+end
 
 end
+

### 2013-11-28 MATLAB/tracking2/tracking_ekfproposal.m

@@ -1,0 +1,50 @@
+function [state, ppsl_prob] = tracking_ekfproposal( model, prev_state, obs, state )
+%tracking_ekfproposal Sample and/or evaluate EKF approximation of the OID
+% for the tracking model.
+
+% If prev_state is empty, assume that this is the first step, so use the
+% prior density instead of the transition density.
+
+% Prior
+if isempty(prev_state)
+    prior_mn = model.m1;
+    prior_vr = model.P1;
+else
+    prior_mn = model.A*prev_state;
+    prior_vr = model.Q;
+end
+
+% Observation mean
+obs_mn = tracking_h(model, prior_mn);
+
+% Linearise observation model around the prior mean
+R = model.R;
+H = tracking_obsjacobian(model, prior_mn);
+obs_lin = obs - obs_mn + H*prior_mn;
+
+% Resolve angle ambiguity
+% Bearing
+if obs_lin(1) > pi
+    obs_lin(1) = obs_lin(1) - 2*pi;
+elseif obs_lin(1) < -pi
+    obs_lin(1) = obs_lin(1) + 2*pi;
+end
+
+% EKF update
+[ppsl_mn, ppsl_vr] = ekf_update1(prior_mn, prior_vr, obs_lin, H, R, obs_mn);
+ppsl_vr = (ppsl_vr + ppsl_vr')/2;
+
+% Sample state if not provided
+if (nargin<4)||isempty(state)
+    state = mvnrnd(ppsl_mn', ppsl_vr)';
+end
+
+% Calculate probability if required
+if nargout>1
+    ppsl_prob = loggausspdf(state, ppsl_mn, ppsl_vr);
+else
+    ppsl_prob = [];
+end
+
+end
+

### 2013-11-28 MATLAB/tracking2/tracking_generatedata.m

@@ -1,0 +1,24 @@
+function [ time, state, observ ] = tracking_generatedata( model )
+%tracking_generatedata Generate a data set for thye tracking model.
+
+% Initialise arrays
+time = 1:model.K;
+state = zeros(model.ds, model.K);
+observ = zeros(model.do, model.K);
+
+% First state
+[state(:,1), ~] = tracking_stateprior(model);
+
+% Loop through time
+for kk = 1:model.K
+    
+    if kk > 1
+        [state(:,kk), ~] = tracking_transition(model, state(:,kk-1));
+    end
+    
+    [observ(:,kk), ~] = tracking_observation(model, state(:,kk));
+    
+end
+
+end
+

### 2013-11-28 MATLAB/tracking2/tracking_setalgo.m

@@ -1,0 +1,13 @@
+function algo = tracking_setalgo(test)
+% Algorithm parameters
+
+% Number of particles
+algo.N = 1000;
+
+% Smoothing trajectories
+algo.S = 100;
+
+% MH chain length / max rejections
+algo.M = 10;
+
+end

### 2013-11-28 MATLAB/tracking2/tracking_setmodel.m

@@ -1,0 +1,38 @@
+function model = tracking_setmodel(test)
+
+% Model parameters
+
+% Using a 2D near constant velocity model and a
+% bearing-range observation model.
+
+%%%%%%%%%%%%%%%%
+
+% General things
+model.K = 250;                  % Number of time points
+model.ds = 4;                   % Dimension of the state
+model.do = 2;                   % Dimension of the observations
+
+% Parameters
+sigx = 1;
+sigtheta = ( 5*(pi/180) )^2; % Bearing covariance (0.25/5)
+sigr     = 100;                 % Range covariance (0.1/100)
+
+% Matrixes
+T = 1;                          % Sampling period
+a = 0.02;                        % Ensures stability
+model.A = [exp(-a*T) 0 T 0; 
+           0 exp(-a*T) 0 T;
+           0 0 exp(-a*T) 0;
+           0 0 0 exp(-a*T)];
+model.Q = sigx * ...
+    [T^3/3  0      T^2/2  0    ;
+     0      T^3/3  0      T^2/2;
+     T^2/2  0      T      0    ;
+     0      T^2/2  0      T    ];
+model.R = diag([sigtheta sigr]);
+
+% x1 distribution
+model.m1 = [-100; 50; 10; 0];
+model.P1 = diag([0.0005; 0.0005; 0.001; 0.001]);
+
+end

### 2013-11-28 MATLAB/tracking3/tracking_ekfproposal.m

@@ -1,0 +1,56 @@
+function [state, ppsl_prob] = tracking_ekfproposal( model, prev_state, obs, state )
+%tracking_ekfproposal Sample and/or evaluate EKF approximation of the OID
+% for the tracking model.
+
+% If prev_state is empty, assume that this is the first step, so use the
+% prior density instead of the transition density.
+
+% Prior
+if isempty(prev_state)
+    prior_mn = model.m1;
+    prior_vr = model.P1;
+else
+    prior_mn = model.A*prev_state;
+    prior_vr = model.Q;
+end
+
+% Observation mean
+obs_mn = tracking_h(model, prior_mn);
+
+% Linearise observation model around the prior mean
+R = model.R;
+H = tracking_obsjacobian(model, prior_mn);
+obs_lin = obs - obs_mn + H*prior_mn;
+
+% Resolve angle ambiguity
+% Bearing
+if obs_lin(1) > pi
+    obs_lin(1) = obs_lin(1) - 2*pi;
+elseif obs_lin(1) < -pi
+    obs_lin(1) = obs_lin(1) + 2*pi;
+end
+% Elevation
+if obs_lin(2) > pi
+    obs_lin(2) = obs_lin(2) - 2*pi;
+elseif obs_lin(2) < -pi
+    obs_lin(2) = obs_lin(2) + 2*pi;
+end
+
+% EKF update
+[ppsl_mn, ppsl_vr] = ekf_update1(prior_mn, prior_vr, obs_lin, H, R, obs_mn);
+ppsl_vr = (ppsl_vr + ppsl_vr')/2;
+
+% Sample state if not provided
+if (nargin<4)||isempty(state)
+    state = mvnrnd(ppsl_mn', ppsl_vr)';
+end
+
+% Calculate probability if required
+if nargout>1
+    ppsl_prob = loggausspdf(state, ppsl_mn, ppsl_vr);
+else
+    ppsl_prob = [];
+end
+
+end
+

### 2013-11-28 MATLAB/tracking3/tracking_h.m

@@ -1,0 +1,23 @@
+function [ y_mn ] = tracking_h( model, x )
+%nlng_h Deterministic observation function for the nonlinear non-Gaussian
+%benchmark model.
+
+y_mn = zeros(model.do,1);
+
+% Bearing
+y_mn(1) = atan2(x(2), x(1));
+
+% Elevation
+hoz_range = hypot(x(1), x(2));
+y_mn(2) = atan2(x(3), hoz_range);
+
+% Range
+range = sqrt( x(1)^2 + x(2)^2 + x(3)^2 );
+y_mn(3) = range;
+
+% Range Rate
+r = x(1:3);
+v = x(4:6);
+y_mn(4) = dot(r,v)/range;
+
+end

### 2013-11-28 MATLAB/tracking3/tracking_observation.m

@@ -1,0 +1,39 @@
+function [ obs, prob ] = tracking_observation( model, state, obs )
+%tracking_observation Sample and/or evaluate observation density for the
+%tracking model.
+
+% prob is a log-probability.
+
+% Calculate observation mean
+mn = tracking_h(model, state);
+
+% Sample observation if not provided
+if (nargin<3)||isempty(obs)
+    obs = mvnrnd(mn', model.R)';
+end
+
+% Calculate probability if required
+if nargout>1
+    
+    % Resolve angle ambiguity
+    dy = obs - mn;
+    % Bearing
+    if dy(1) > pi
+        dy(1) = dy(1) - 2*pi;
+    elseif dy(1) < -pi
+        dy(1) = dy(1) + 2*pi;
+    end
+    % Elevation
+    if dy(2) > pi
+        dy(2) = dy(2) - 2*pi;
+    elseif dy(2) < -pi
+        dy(2) = dy(2) + 2*pi;
+    end
+    
+    prob = loggausspdf(dy, zeros(size(dy)), model.R);
+else
+    prob = [];
+end
+
+end
+

### 2013-11-28 MATLAB/tracking3/tracking_obsjacobian.m

@@ -1,0 +1,20 @@
+function [ H ] = tracking_obsjacobian( model, x )
+%tracking_obsjacobian Calculate observation function jacobian for the
+%tracking model.
+
+rng = sqrt(x(1)^2 + x(2)^2 + x(3)^2);
+hoz_rng_sq = (x(1)^2+x(2)^2);
+hoz_rng = sqrt(hoz_rng_sq);
+
+dh1_dx = [-x(2), x(1), 0, 0, 0, 0]/hoz_rng_sq;
+dh2_dx = [-x(1)*x(3)/hoz_rng, -x(2)*x(3)/hoz_rng, hoz_rng, 0, 0, 0]/(rng^2);
+dh3_dx = [x(1), x(2), x(3), 0, 0, 0]/rng;
+dh4_dr = [x(4)*(x(2)^2+x(3)^2)-x(1)*(x(5)*x(2) + x(6)*x(3)),...
+          x(5)*(x(1)^2+x(3)^2)-x(2)*(x(4)*x(1) + x(6)*x(3)),...
+          x(6)*(x(1)^2+x(2)^2)-x(3)*(x(4)*x(1) + x(5)*x(2))]/(rng^3);
+dh4_dv = [x(1), x(2), x(3)]/rng;
+
+H = [dh1_dx; dh2_dx; dh3_dx; [dh4_dr dh4_dv]];
+     
+end
+

### 2013-11-28 MATLAB/tracking3/tracking_setmodel.m

@@ -1,0 +1,40 @@
+function model = tracking_setmodel(test)
+
+% Model parameters
+
+% Using a 3D near constant velocity model and a
+% bearing-elevation-range-rangerate observation model.
+
+%%%%%%%%%%%%%%%%
+
+% General things
+model.K = 100;                  % Number of time points
+model.ds = 6;                   % Dimension of the state
+model.do = 4;                   % Dimension of the observations
+
+% Parameters
+sigx = 1;
+sigtheta = ( 10*(pi/180) )^2;   % Bearing covariance
+sigphi   = ( 1*(pi/180) )^2;   % Elevation covariance
+sigr     = 1^2;                % Range covariance
+sigrr    = 10^2;                % Range rate covariance
+
+% Matrixes
+T = 1;                          % Sampling period
+a = 0.01;                        % Ensures stability
+model.A = [exp(-a*T) 0 0 T 0 0; 0 exp(-a*T) 0 0 T 0; 0 0 exp(-a*T) 0 0 T;
+           0 0 0 exp(-a*T) 0 0; 0 0 0 0 exp(-a*T) 0; 0 0 0 0 0 exp(-a*T)];
+model.Q = sigx * ...
+    [T^3/3  0      0      T^2/2  0      0    ;
+     0      T^3/3  0      0      T^2/2  0    ;
+     0      0      T^3/3  0      0      T^2/2;
+     T^2/2  0      0      T      0      0    ;
+     0      T^2/2  0      0      T      0    ;
+     0      0      T^2/2  0      0      T    ];
+model.R = diag([sigtheta sigphi sigr sigrr]);
+
+% x1 distribution
+model.m1 = [-100 100 20 10 0 0]';
+model.P1 = diag([10 10 10 1 1 1]);
+
+end

### 2013-11-28 MATLAB/tracking3/tracking_stateprior.m

@@ -1,0 +1,20 @@
+function [ state, prob ] = tracking_stateprior( model, state )
+%tracking_stateprior Sample and/or evaluate observation density for the
+%tracking model.
+
+% prob is a log-probability.
+
+% Sample state if not provided
+if (nargin<2)||isempty(state)
+    state = mvnrnd(model.m1', model.P1)';
+end
+
+% Calculate probability if required
+if nargout>1
+    prob = loggausspdf(state, model.m1, model.P1);
+else
+    prob = [];
+end
+
+end
+

### 2013-11-28 MATLAB/tracking3/tracking_transition.m

@@ -1,0 +1,23 @@
+function [ new_state, prob ] = tracking_transition( model, state, new_state )
+%tracking_transition Sample and/or evaluate observation density for the
+%tracking model.
+
+% prob is a log-probability.
+
+% Calculate new_state mean
+mn = model.A * state;
+
+% Sample state if not provided
+if (nargin<3)||isempty(new_state)
+    new_state = mvnrnd(mn', model.Q)';
+end
+
+% Calculate probability if required
+if nargout>1
+    prob = loggausspdf(new_state, mn, model.Q);
+else
+    prob = [];
+end
+
+end
+

### 2013-11-28 MATLAB/tracking3/tracking_ukfproposal.m

@@ -1,0 +1,36 @@
+function [state, ppsl_prob] = tracking_ukfproposal( model, prev_state, obs, state )
+%tracking_ukfproposal Sample and/or evaluate UKF approximation of the OID
+% for the tracking model.
+
+% If prev_state is empty, assume that this is the first step, so use the
+% prior density instead of the transition density.
+
+% Prior
+if isempty(prev_state)
+    prior_mn = model.m1;
+    prior_vr = model.P1;
+else
+    prior_mn = model.A*prev_state;
+    prior_vr = model.Q;
+end
+
+% UKF update
+R = model.R;
+h = @(x, par)tracking_h(model, x);
+[ppsl_mn, ppsl_vr] = ukf_update1(prior_mn, prior_vr, obs, h, R, [], [], [], 1);
+ppsl_vr = (ppsl_vr+ppsl_vr')/2;
+
+% Sample state if not provided
+if (nargin<4)||isempty(state)
+    state = mvnrnd(ppsl_mn', ppsl_vr)';
+end
+
+% Calculate probability if required
+if nargout>1
+    ppsl_prob = loggausspdf(state, ppsl_mn, ppsl_vr);
+else
+    ppsl_prob = [];
+end
+
+end
+
